---
title: "me20260129_RCMC_Sox2_Figure_1"
author: "Mathias"
date: "2026-01-29"
output: 
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---

# Aim
I want to plot the 3D configuration of the mouse Sox2 locus from the region capture microC data (GSM6281849) form Goel et al. (Hansen lab). 
The data was in mm39, therefore we had to liftover the coordinates to mm10. 

# Set-up
The newer ggplot version is not compatible with the plotting in the genova package, therefore i loaded an older version of ggplot in order to be compatible with the pyramid function of the GENOVA package. 
```{r}
# install.packages("remotes")  # if needed

# dir.create("/DATA/usr/m.eder/R/ggplot2_3x", showWarnings = FALSE, recursive = TRUE)
# 
# remotes::install_version(
#   "ggplot2",
#   version = "3.5.2",
#   lib = "/DATA/usr/m.eder/R/ggplot2_3x",
#   repos = "https://cloud.r-project.org"
# )
```

Load the libraries and set the parameters.
```{r load_libraries}
# Load older ggplot version that is compatible with the GENOVA package
.libPaths(c("/DATA/usr/m.eder/R/ggplot2_3x", .libPaths()))

library(ggplot2)
packageVersion("ggplot2")  # should now show 3.5.2

# Load dependencies
library(GenomicRanges)
library(rtracklayer)
library(gtools)
library(tidyverse)
library(dtplyr)
library(dplyr)
library(tidyr)
library(ggbeeswarm)
library(ggplot2)
library(Biostrings)
library(stringr)
library(readr)
library(knitr)
library(RColorBrewer)
library(purrr)
library(RcppRoll)
library(viridis)
library(ggh4x)
library(GENOVA)

split_string <- function(vect,sep,N1,N2=N1){
  library(stringr)
  sapply(vect, function(X){
    paste(str_split(X,sep)[[1]][N1:N2],collapse = sep)
  },USE.NAMES = F)
}
```

# Paths & parameters
```{r paths_parameters}
setwd("/DATA/projects/CTCF_Sox2/Scripts/for_figures/Figure_1")

#datatag
datetag = paste0("ME",format(Sys.time(), '%Y%m%d'))

# Prepare output 
output_dir <- paste0("/DATA/projects/CTCF_Sox2/Scripts/for_figures/Figure_1/figure1_",datetag)
dir.create(output_dir, showWarnings = FALSE)

```

```{r knits setup}

opts_chunk$set(fig.width = 20, fig.height = 8, cache = T,
               message = F, 
               warning = F,
               dev=c('png', 'pdf'), 
               dpi = 600,
               fig.path = file.path(output_dir, "/figure_1/")) 
pdf.options(useDingbats = FALSE)
```



# File paths
```{r}
path_CTCF_sites = "/DATA/usr/v.franceschini/Workspaces/2024_01_MATHIAS_CTCF_PEAKS/VF240812_calling_CTCF_motifs_again/02_OUTPUTS/CTCF_sites/6764_2_ME_E2_CGATGT_S2_R1_001_peaks_motifs.bed"
```

# Relevant annotations
```{r element_coordinates}

# Location of enhancer / gene
enhancer <- GRanges(seqnames = "chr3", 
                    ranges = IRanges(start = 34753415,
                                     end = 34766401),
                    strand = "*")

#gene based on mm10, ucsc browser, RefSeq annotation
# Sox2
# chr3:34649995-34652461 (+)
# id = NM_011443.4
Sox2_gene <- GRanges(seqnames = "chr3", 
                    ranges = IRanges(start = 34649995,
                                     end = 34652461),
                    strand = "+")


landingPad_23 <- GRanges(seqnames = "chr3", 
                    ranges = IRanges(start = 34643960,
                                     end = 34643962),
                    strand = "+")

landingPad_8 <- GRanges(seqnames = "chr3", 
                    ranges = IRanges(start = 34721183,
                                     end = 34721184),
                    strand = "+")


LP_tib = tibble(launch_pad_location = c("-116 kb", "-39 kb"), 
                start = c( start(landingPad_23),start(landingPad_8)))

#CTCF sites based on our own mapping
# new analysis (based on mapping to mm10 only)
CTCF_mm10_new <- import.bed('/DATA/usr/v.franceschini/Workspaces/2024_01_MATHIAS_CTCF_PEAKS/VF240812_calling_CTCF_motifs_again/02_OUTPUTS/CTCF_sites/6764_2_ME_E2_CGATGT_S2_R1_001_peaks_motifs.bed')

#filtering:
CTCF_mm10.chr3 <- CTCF_mm10_new[seqnames(CTCF_mm10_new)== 'chr3']
# add the missing site after the SCR
CTCF_mm10.chr3_extra = sort(c(CTCF_mm10.chr3,
                         GRanges(seqnames = "chr3",
                                 IRanges(start = 34772210, end = 34772210),
                                 strand = "+")),
                         ignore.strand = T)


prange_zoom = c(34500000 , 35000000)
prange2 = c(34000000, 35500000)
prange_far = c(32000000, 37000000)
#zoom in on only the SCR +- 50kb
prange_LP = c(start(landingPad_23)-5E4, end(landingPad_23)+5E4)
prange_SCR = c(start(enhancer)-5E4, end(enhancer)+5E4)
prange_plot = c(32643960, 36643960) #+/- 2MB from landing pad
prange_loop = c(34631804, 34778268)
prange_zoom = c(34.5E6, 35E6)
prange_zoom_paper = c(34.59E6, 34.83E6)
prange_SCR = c(34.725E6, 34.785E6)
```

# 3D genome annotation
## LiftOver chains
```{r}
# download.file('https://hgdownload.soe.ucsc.edu/goldenPath/mm9/liftOver/mm9ToMm10.over.chain.gz',
#               "/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/general_analyses/region_capture_microC/mm9ToMm10.over.chain.gz")
# download.file('https://hgdownload.soe.ucsc.edu/goldenPath/mm10/liftOver/mm10ToMm9.over.chain.gz',
#               "/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/general_analyses/region_capture_microC/mm10ToMm9.over.chain.gz")
# download.file('https://hgdownload.soe.ucsc.edu/goldenPath/mm39/liftOver/mm39ToMm10.over.chain.gz',
#               "/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/general_analyses/region_capture_microC/mm39ToMm10.over.chain.gz")
# download.file('https://hgdownload.soe.ucsc.edu/goldenPath/mm10/liftOver/mm10ToMm39.over.chain.gz',
#               "/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/general_analyses/region_capture_microC/mm10ToMm39.over.chain.gz")

# library(R.utils)
# gunzip("/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/general_analyses/region_capture_microC/mm9ToMm10.over.chain.gz")
# gunzip("/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/general_analyses/region_capture_microC/mm10ToMm9.over.chain.gz")
# gunzip("/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/general_analyses/region_capture_microC/mm39ToMm10.over.chain.gz")
# gunzip("/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/general_analyses/region_capture_microC/mm10ToMm39.over.chain.gz")

chain_mm9_to_mm10 = import.chain("/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/general_analyses/region_capture_microC/mm9ToMm10.over.chain")
chain_mm10_to_mm9 = import.chain("/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/general_analyses/region_capture_microC/mm10ToMm9.over.chain")
chain_mm10_to_mm39 = import.chain("/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/general_analyses/region_capture_microC/mm10ToMm39.over.chain")
chaim_mm39_to_mm10 = import.chain("/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/general_analyses/region_capture_microC/mm39ToMm10.over.chain")
#landingPad_23_mm39= liftOver()

```

## Load RCMC data
```{r}
WT_file = "/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/general_analyses/region_capture_microC/CM20230718_GSM6281849_RCMC_BR1_merged_allCap_WT_mm39.merged.50.mcool"

RCMC_1kb = load_contacts(signal_path = WT_file,
                      sample_name = 'WT',
                      resolution = 1000)

RCMC_5kb = load_contacts(signal_path = WT_file,
                      sample_name = 'WT',
                      resolution = 5000)
```

```{r A_RCMC_pyramid_annotated, fig.width = 7.2, fig.height = 3}
#plotting range on mm39
plot_range_hiC = c(34.65E6,
                   34.9E6)

#annotations
annotation_gr_mm10 = c(Sox2_gene, enhancer)
names(annotation_gr_mm10) = c("Sox2", "SCR")

annotation_gr_mm39 = unlist(liftOver(annotation_gr_mm10, chain_mm10_to_mm39))
bed_tib_mm39 = as_tibble(annotation_gr_mm39)

# CTCF annotation
CTCF_ROI_mm10 = subsetByOverlaps(CTCF_mm10.chr3_extra, GRanges(seqnames = "chr3", IRanges(start = 30E6, end = 40E6)))
CTCF_ROI_mm39 = unlist(liftOver(CTCF_ROI_mm10, chain_mm10_to_mm39))
CTCF_ROI_mm39_tib = as_tibble(CTCF_ROI_mm39)
  
p <- GENOVA::pyramid(
  exp = RCMC_1kb,
  chrom = "chr3",
  colour = c(0, 1500),
  start = plot_range_hiC[1],
  end = plot_range_hiC[2]
)

p + 
  geom_rect(data = bed_tib_mm39, aes(xmin = start, xmax = end), ymin = -30E3, ymax = -10E3) +
  geom_segment(data = CTCF_ROI_mm39_tib, aes(x = start, xend = start, col = strand), y = -30E3, yend = -10E3) +
  geom_vline(xintercept = start(CTCF_ROI_mm39), col = 'darkgrey',linetype = 'dotted')

```

