---
title: "me20260303_CBS_hopping_bias"
author: "Mathias Eder"
date: "2026-03-03"
output: 
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---

# Aim
Analyse the CBS induced hopping bias of CBS-Sox2P and CBS only hopping.

#Figures
- fig 6B
- fig 6C
- fig 6D
- fig S6G
- fig 6E
- fig S5F

# Set-up
##Libraries
Load the libraries and set the parameters.

```{r load_libraries}

# Load dependencies
library(GenomicRanges)
library(rtracklayer)
library(gtools)
library(tidyverse)
library(dtplyr)
library(dplyr)
library(tidyr)
library(ggbeeswarm)
library(ggplot2)
#library(broom)
library(Biostrings)
library(stringr)
library(readr)
#library(tagMeppr)
#library(withr)
library(knitr)
library(RColorBrewer)
#library(tagmapAnalyseR)
# library(UpSetR)
# library(ComplexHeatmap)
library(purrr)
# library(ggpubr)
# library(cowplot)
library(RcppRoll)
library(viridis)
library(plotly)
library(GGally)
library(cowplot)
library(ggh4x)
library(Cairo)

split_string <- function(vect,sep,N1,N2=N1){
  library(stringr)
  sapply(vect, function(X){
    paste(str_split(X,sep)[[1]][N1:N2],collapse = sep)
  },USE.NAMES = F)
}
```

# Paths & parameters
```{r paths_parameters}
setwd("/DATA/projects/CTCF_Sox2/Scripts/for_figures/Figure_6")

#datatag
datetag = paste0("ME",format(Sys.time(), '%Y%m%d'))

# Prepare output 
output_dir <- paste0("/DATA/projects/CTCF_Sox2/Scripts/for_figures/Figure_6/",datetag)
dir.create(output_dir, showWarnings = FALSE)
```

```{r knits setup}

library(knitr)
opts_chunk$set(cache = T,
               message = F, 
               warning = F,
               dev=c('png', 'pdf'), 
               dpi = 600,
               fig.path = file.path(output_dir, "figures/"),
               cache.lazy = F) 
pdf.options(useDingbats = FALSE)
```

# File paths
## Load CBS-Sox2P data
```{r}
path_CTCF_sites = "/DATA/usr/v.franceschini/Workspaces/2024_01_MATHIAS_CTCF_PEAKS/VF240812_calling_CTCF_motifs_again/02_OUTPUTS/CTCF_sites/6764_2_ME_E2_CGATGT_S2_R1_001_peaks_motifs.bed"

# Path for integrations
path_CTCF_ints = "/DATA/projects/CTCF_Sox2/Integrations_files/mapped_integrations_-116kb_CBS_pools.txt"
path_tib_23_ints = "/DATA/projects/Sox2/GEO_collection/Tagmentation/still_to_upload/mapped_integrations_-116kb_Sox2P_pools.txt"
path_tib_CTCF_Sox2P = "/DATA/projects/CTCF_Sox2/Integrations_files/mapped_integrations_-116kb_CBS_Sox2P_pools.txt"

# this RDS comes from the following markdown 'me20260302_figure_2_S1_S3"
tib_filtered_curated_strict = readRDS("/DATA/projects/CTCF_Sox2/Scripts/CTCF_Sox2P/tib_filtered_curated_strict.rds")

tib_filtered_curated_strict = tib_filtered_curated_strict %>%
  mutate(cell_line = factor(cell_line, levels = c("23_34A", "23_MK03_C1", "23_MK01_C5", "23_MK02_C2", "8_34", "8_MK03_C3", "8_MK01_C1", "8_MK01_C2", "8_MK02_C5")))
```

## Load CTCF only data
# Load and filter CTCF hopping data
```{r load CTCF data}

# load CTCF data
tib_CTCF = read_tsv(path_CTCF_ints) %>%
  filter(chr %in% paste0("chr", c(1:19, "X"))) %>%
  mutate(chr = factor(chr, levels = paste0("chr", c(1:19, "X")))) %>%
  group_by(chr, start, end, seq, strand, replicate, experiment, launch_pad_location, insert, genotype, sample_type, sorted_gate, pool_nr, sample, cell_line, CBS_motif_orientation) %>%
  summarize(read_count = sum(read_count),
            read_count_1 = sum(read_count_1),
            read_count_2 = sum(read_count_2), .groups = "keep") %>%
  #mark hopped integrations
  mutate(hopped = start != 34643961)

#filter the integrations
tib_filtered_CTCF <- tib_CTCF %>%
  #add variables used in the plotting (old names of cell lines etc) 
  mutate(population = case_when(sample_type == "control pool" ~ "ctrl",
                                .default = sorted_gate),
         #annotate confidence of mapping (one or both ITRs)
         mapped_arms = case_when(read_count_1 == 0 ~ "rv_only",
                                 read_count_2 == 0 ~ "fw_only",
                                 read_count_1 > 0 & read_count_2 > 0 ~ "both_arms"),
         landing_pad = case_when(launch_pad_location == "-116 kb" ~ "23"),
         motif = case_when(CBS_motif_orientation == "FW" ~ "forward",
                           CBS_motif_orientation == "RV" ~ "reverse",
                           CBS_motif_orientation == "-" & strand == "+" ~ "forward",
                           CBS_motif_orientation == "-" & strand == "-" ~ "reverse",
                           .default = CBS_motif_orientation)) %>%
  #filter the data
  filter(population == "ctrl" | read_count >= 2) %>% #for ctrl distribution 1 mapping read is enough
  filter(!(start >= 34721183 & start <= 34721192)) %>% #remove contamination from LP8
  filter(strand %in% c("+", "-")) %>% #remove non annotated strand
  filter(!(population == "mCh_low" & mapped_arms != "both_arms")) %>% #only keep mCh_low integrations with 2 mapped arms
  filter(!(experiment == "E1665" & population == "mCh_mid" & mapped_arms != "both_arms")) %>% #mCh_close is the population of interest in E1665, mid is a bit lower and needs to be filtered for not having deletions
  mutate(population = case_when(experiment == "E1665" & population == "mCh_close" ~ "mCh_mid",
                                .default = population))

# quick check
table(tib_filtered_CTCF$cell_line, tib_filtered_CTCF$population)
table(tib_filtered_CTCF$population, tib_filtered_CTCF$chr)

# Combine the 1xCBS cell lines (CBS16A + 16A)
tib_CBS16_16A_comb = tib_filtered_CTCF %>% filter(population %in% c("ctrl", "mCh_mid", "mCh_high")) %>% 
  mutate(cell_line = case_when(cell_line == "CBS16" ~ "CBS16_16A",
                               cell_line == "16A" ~ "CBS16_16A",
                               .default = cell_line))

tib_CBS16_16A_comb = tib_CBS16_16A_comb %>%
  mutate(cell_line = factor(cell_line, levels = c("37A", "CRE6", "CBS16_16A", "3xCTCF")),
         landing_pad = case_when(landing_pad == "23" ~ "LP23",
                                 .default = landing_pad)) %>%
  ungroup() %>%
  select(chr, start, end, seq, strand, replicate, experiment, launch_pad_location, insert, genotype, sample_type, sorted_gate, pool_nr, read_count, read_count_1, read_count_2, hopped, cell_line, population, landing_pad, mapped_arms)

all_exp = unique(tib_filtered_CTCF$experiment)
```


# Relevant annotations
## Genomic coordinates of features
```{r element_coordinates}
# Location of enhancer / gene
enhancer <- GRanges(seqnames = "chr3", 
                    ranges = IRanges(start = 34753415,
                                     end = 34766401),
                    strand = "*")


#gene based on mm10, ucsc browser, RefSeq annotation
# Sox2
# chr3:34649995-34652461 (+)
# id = NM_011443.4

Sox2_gene <- GRanges(seqnames = "chr3", 
                    ranges = IRanges(start = 34649995,
                                     end = 34652461),
                    strand = "+")


landingPad_23 <- GRanges(seqnames = "chr3", 
                    ranges = IRanges(start = 34643960,
                                     end = 34643962),
                    strand = "+")

landingPad_8 <- GRanges(seqnames = "chr3", 
                    ranges = IRanges(start = 34721183,
                                     end = 34721184),
                    strand = "+")


LP_tib = tibble(landing_pad = c("LP23", "LP8"), 
                start = c( start(landingPad_23),start(landingPad_8)))

#CTCF sites based on our own mapping
# new analysis (based on mapping to mm10 only)
CTCF_mm10_new <- import.bed('/DATA/usr/v.franceschini/Workspaces/2024_01_MATHIAS_CTCF_PEAKS/VF240812_calling_CTCF_motifs_again/02_OUTPUTS/CTCF_sites/6764_2_ME_E2_CGATGT_S2_R1_001_peaks_motifs.bed')

#filtering:
CTCF_mm10.chr3 <- CTCF_mm10_new[seqnames(CTCF_mm10_new)== 'chr3']
# add the missing site after the SCR
CTCF_mm10.chr3_extra = sort(c(CTCF_mm10.chr3,
                         GRanges(seqnames = "chr3",
                                 IRanges(start = 34772210, end = 34772210),
                                 strand = "+")),
                         ignore.strand = T)

# add the missing site at 35.5
CTCF_mm10.chr3_extra = sort(c(CTCF_mm10.chr3_extra,
                         GRanges(seqnames = "chr3",
                                 IRanges(start = 35574646, end = 35574664),
                                 strand = "-")),
                         ignore.strand = T)

```

## Plotting ranges
```{r Plotting ranges}
prange_zoom = c(34500000 , 35000000)
prange2 = c(34000000, 35500000)
prange_far = c(32000000, 37000000)
#zoom in on only the SCR +- 50kb
prange_LP = c(start(landingPad_23)-5E4, end(landingPad_23)+5E4)
prange_SCR = c(start(enhancer)-5E4, end(enhancer)+5E4)
prange_plot = c(32643960, 36643960) #+/- 2MB from landing pad
prange_loop = c(34631804, 34778268)
prange_zoom = c(34.5E6, 35E6)
prange_zoom_paper = c(34.59E6, 34.83E6)
prange_SCR = c(34.725E6, 34.785E6)
range_2Mb = c(33.643960E6, 35.643960E6)
```

## Color schemes
```{r Color schemes}
myCols_insert_cell_line <- c(
  `23_MK01_C5` = '#1b9e77',   # Teal green
  `23_MK02_C2` = '#d95f02',   # Strong orange
  `23_MK03_C1` = '#7570b3',   # Purple
  `23_34A` = 'black', # blue for ctrl cell line
  `8_MK01_C1` = '#1b9e77',   # Teal green
  `8_MK01_C2` = '#1b9e77',   # Teal green
  `8_MK02_C5` = '#d95f02',   # Strong orange
  `8_MK03_C3` = '#7570b3',   # Purple
  `8_34` = 'black' # blue for ctrl cell line
)

col_scale_MK_CL <- scale_colour_manual(
  name = "cell_line", 
  values = myCols_insert_cell_line,
  aesthetics = c("colour", "fill")
)

# Color annotations
## SCR 
brown = "#ffb000"

#color schemes
myCols_blue = c('#525252', "#080E5B",  "#212B71", "#3A4987", "#465893", "#5A719A", "#6C8AA0")
names(myCols_blue) = c("ctrl", paste0("P", 1:6))
colScale7_blue <- scale_colour_manual(name = "population", values = myCols_blue)
fillScale7_blue <- scale_fill_manual(name = "population", values = myCols_blue)

myCols_blue_strand = c( "#080E5B", "#5A719A")
names(myCols_blue_strand) = c("+","-")
colScale2_blue <- scale_colour_manual(name = "strand", values = myCols_blue_strand)
fillScale2_blue <- scale_fill_manual(name = "strand", values = myCols_blue_strand)

myCols_green = c('#525252',"#0A3F30", "#175643", "#246D58", "#30846B", "#3D9B7E", "#49B292")
names(myCols_green) = c("ctrl", paste0("P", 1:6))
colScale7_green <- scale_colour_manual(name = "population", values = myCols_green)
fillScale7_green <- scale_fill_manual(name = "population", values = myCols_green)

myCols_orange = c('#525252', "#572601", "#72380B", "#8F4A15", "#AA5C20", "#C66D2A", "#E07F34")
names(myCols_orange) = c("ctrl", paste0("P", 1:6))
colScale7_orange <- scale_colour_manual(name = "population", values = myCols_orange)
fillScale7_orange <- scale_fill_manual(name = "population", values = myCols_orange)

myCols_purple = c('#525252', "#2E2D47", "#424060", "#565378", "#696791", "#7D7AA9", "#918DC2")
names(myCols_purple) = c("ctrl", paste0("P", 1:6))
colScale7_purple <- scale_colour_manual(name = "population", values = myCols_purple)
fillScale7_purple <- scale_fill_manual(name = "population", values = myCols_purple)
```


# Functions
## Beeswarm function
```{r Beeswarm function}
plot_beeswarm_only = function(TIB, CL = "23_34A", EXP = all_exp, P_RANGE = prange_zoom,
                              PLOT_CTCF = F, PLOT_ANNOT = T,
                              TITLE = NULL, POINT_SIZE = 0.75, RANDOM_METHOD = "quasirandom",
                              NEXT_RANGE = NULL){
  tib_for_plot = filter(TIB, hopped & chr == "chr3" & cell_line %in% CL & experiment %in% EXP & start >= P_RANGE[1] & start <= P_RANGE[2])
  LP_tib_cl = select(tib_for_plot, c(cell_line, landing_pad)) %>% 
    distinct() %>% left_join(LP_tib)
  
  ggplot(filter(tib_for_plot, hopped & chr == "chr3" & cell_line %in% CL & experiment %in% EXP),
         aes(x = start, y = population, col = population)) +
    {if(length(CL)>1) facet_grid(cell_line ~ .)} +  #only facet when necessary

    {if (PLOT_ANNOT)
      list( #add elements in a list, you cannot use + inside an if statement
        # annotate enhancer
        geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown),
        annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown),
        # annotate gene
        annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf, alpha = .2, fill = 'red'),
        geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = 'red') )} +
    
     #annotate landing pad
    {if(PLOT_ANNOT & length(CL)== 1)
      geom_vline(xintercept = pull(
        filter(LP_tib, landing_pad %in% unique(tib_for_plot$landing_pad)), 
        start),
        col = 'darkgrey')} +
    {if(PLOT_ANNOT & length(CL) > 1)
       geom_vline(data = LP_tib_cl, aes(xintercept = start), col = 'darkgrey')} +

    # # annotate CTCFs
    {if(PLOT_CTCF)geom_vline(xintercept = start(CTCF_mm10.chr3_extra), col = 'darkgrey', lty="11")}+
    {if(!is.null(NEXT_RANGE)) geom_rug(data = tibble(range = NEXT_RANGE),
                                       aes(x = range),
                                       fill = 'grey',
                                       inherit.aes = F)}+
    
    
    #datapoints
    geom_quasirandom(alpha = 0.5, size = POINT_SIZE, varwidth = T, bandwidth = 0.8,
                     # shape = 16, #changes to point without border, so alpha applies to the whole point
                     method = RANDOM_METHOD,
                     stroke = NA) +
    geom_text(data = plyr::count(tib_for_plot, vars = c("cell_line", "population")), aes(label = paste("n = ", freq, sep = ""),
                                                                                         x = Inf),
              hjust = "inward", vjust = "top",
              col = 'black') +
    
    #layout:
    labs(x='Genomic coordinate (Mb)',y='sorted gate') +
    theme_classic(base_size = 16) +
    theme(legend.position = "none") +
    scale_x_continuous(breaks = scales::pretty_breaks(n = 6), labels = scales::unit_format(scale = 1E-6, accuracy = 0.01, unit=NULL),
                       limits=P_RANGE, expand=c(0,0)) +
    scale_y_discrete(limits = rev, drop=F) +
    ggtitle(TITLE) +
    ## CHANGED: choose the appropriate colour scale based on CL
    { if (length(CL)==1) {
        switch(CL,
          "23_34A"       = colScale7_blue,
          "23_MK01_C5"   = colScale7_green,
          "23_MK02_C2"   = colScale7_orange,
          "23_MK03_C1"   = colScale7_purple,
          colScale7_blue  # default fallback
        )
      } else {
        colScale7_blue  # when faceting multiple lines, fall back to blue (or choose your own default)
      }
    }
}
```

## Stranded beeswarm function
```{r stranded beeswarm}
plot_beeswarm_only_stranded = function(TIB, CL = "23_34A", EXP = all_exp, P_RANGE = prange_zoom,
                              PLOT_CTCF = FALSE, PLOT_ANNOT = TRUE,
                              SPLIT_STRAND = FALSE,
                              TITLE = NULL, POINT_SIZE = 0.75, RANDOM_METHOD = "quasirandom",
                              NEXT_RANGE = NULL){
  
  # 1) subset once
  tib_for_plot <- TIB %>%
    filter(hopped, chr == "chr3",
           cell_line %in% CL, experiment %in% EXP,
           between(start, P_RANGE[1], P_RANGE[2]))
  
  # 2) landing‐pad join (unchanged)
  LP_tib_cl <- tib_for_plot %>%
    distinct(cell_line, landing_pad) %>%
    left_join(LP_tib, by = "landing_pad")
  
  # 3) compute counts *with* strand if requested
  if (SPLIT_STRAND) {
    count_df <- tib_for_plot %>%
      count(cell_line, strand, population)
  } else {
    count_df <- tib_for_plot %>%
      count(cell_line, population)
  }
  
 
  # 4) now build the plot
  ggplot(tib_for_plot,
         aes(x = start, y = population, col = population)) +
    
    # faceting
    {
      if (SPLIT_STRAND) {
        if (length(CL) > 1) facet_grid(cell_line ~ strand)
        else                facet_wrap(~ strand)
      } else {
        if (length(CL) > 1) facet_grid(cell_line ~ .)
        else                NULL
      }
    } +
    
    # (all your annotations go here, unchanged) … +
    {if (PLOT_ANNOT)
      list( #add elements in a list, you cannot use + inside an if statement
        # annotate enhancer
        geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown),
        annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown),
        # annotate LP
        geom_vline(xintercept = start(landingPad_8), col = "black", linetype = "dotted"), 
        geom_vline(xintercept = start(landingPad_23), col = "black", linetype = "dotted"),
        
        # annotate gene
        annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf, alpha = .2, fill = 'red'),
        geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = 'red') )} +
    
    # # annotate CTCFs
    {if(PLOT_CTCF)geom_vline(xintercept = start(CTCF_mm10.chr3_extra), col = 'darkgrey', lty="11")}+
    {if(!is.null(NEXT_RANGE)) geom_rug(data = tibble(range = NEXT_RANGE),
                                       aes(x = range),
                                       fill = 'grey',
                                       inherit.aes = F)}+
    
    geom_quasirandom(alpha = 0.5, size = POINT_SIZE, varwidth = TRUE,
                     bandwidth = 0.8, method = RANDOM_METHOD, stroke = NA) +
    
    # 5) the *only* geom_text, pointing at count_df
    geom_text(
      data = count_df,
      aes(label = paste0("n = ", n), x = Inf),
      hjust = "inward", vjust = "top",
      col = 'black'
    ) +
    
    # 6) scales / theme / title (unchanged) …
    labs(x = 'Genomic coordinate (Mb)', y = 'sorted gate') +
    theme_classic(base_size = 16) +
    theme(legend.position = "none") +
    scale_x_continuous(breaks = scales::pretty_breaks(n = 6), labels = scales::unit_format(scale = 1E-6, accuracy = 0.01, unit=NULL),
                       limits=P_RANGE, expand=c(0,0)) +
    scale_y_discrete(limits = rev, drop=F) +
    ggtitle(TITLE) +
    ## CHANGED: choose the appropriate colour scale based on CL
    { if (length(CL)==1) {
        switch(CL,
          "23_34A"       = colScale7_blue,
          "23_MK01_C5"   = colScale7_green,
          "23_MK02_C2"   = colScale7_orange,
          "23_MK03_C1"   = colScale7_purple,
          colScale7_blue  # default fallback
        )
      } else {
        colScale7_blue  # when faceting multiple lines, fall back to blue (or choose your own default)
      }
    }
    
}

```

## Enrichment function
How does the enrichment function work
1. makes bins, counts int per bin
2. calculates fraction per bin (fraction over total)
3. summes fractions per window per population
4. divides window fraction of interest over control 

NOTE: I added a section to count integrations of separate populations per window. This should allow us afterwards to filter for windows that have integrations present. 
```{r adapted enrichment function}
# Adapted from Christines function

####################
#Below is the function to calculate enrichment over control per N kb.
# Was built for 6 gate reporter hopping, but I think this one should directly work for CTCF hopping as well.

# It takes the integrations per binsize and number of bins, so 1000 *10 = 10kb window, slided by 1kb
#####################

enrichm_function_fixed_bins_CL <- function(TIB, CL = "CBS16", EXP = all_exp, 
                                           binsize = 1000, N_bins_window = 10, 
                                           calculation_window =  c(34091354, 35744137),
                                           MOTIF = c("forward", "reverse")){
  tib_for_analysis = dplyr::filter(TIB, hopped & cell_line %in% CL & experiment %in% EXP)

  # total integrations per population (no strand filter here)
  total_ints = tib_for_analysis %>% dplyr::group_by(population) %>% dplyr::summarise(total_ints = dplyr::n(), .groups = "drop")

  # bin boundaries
  boundaries_all = seq(from = calculation_window[1], to = calculation_window[2], by = binsize)
  boundaries_all_tib = tibble::tibble(interval_number = seq_along(boundaries_all), start_interval = boundaries_all) %>%
    dplyr::arrange(start_interval) %>%
    dplyr::mutate(mid_interval = (start_interval + dplyr::lead(start_interval, 1))/2)

  # counts per bin per population
  cut_by_intervals = tib_for_analysis %>% 
    dplyr::filter(motif %in% MOTIF) %>% 
    dplyr::filter(chr == "chr3", start >= calculation_window[1], start <= calculation_window[2]) %>%
    dplyr::arrange(start) %>%
    dplyr::mutate(interval_number = findInterval(start, boundaries_all)) %>%
    dplyr::group_by(population, interval_number) %>%
    dplyr::summarise(N_ints = dplyr::n(), .groups = "drop")

  # wide: counts per bin (include empty bins)
  cut_by_intervals_wide = dplyr::left_join(boundaries_all_tib, cut_by_intervals, by = "interval_number") %>%
    dplyr::mutate(N_ints = tidyr::replace_na(N_ints, 0)) %>%
    dplyr::arrange(interval_number) %>%
    tidyr::pivot_wider(id_cols = c('start_interval','interval_number','mid_interval'),
                       names_from = population, values_from = N_ints, values_fill = 0) %>%
    dplyr::select(-tidyselect::any_of('NA'))

  ## ---- NEW PART: rolling raw counts per window -----------------------------
  # compute rolling sum of raw counts per population
  counts_roll_wide <- cut_by_intervals_wide %>%
    dplyr::mutate(dplyr::across(
      -c(start_interval, interval_number, mid_interval),
      ~ RcppRoll::roll_sum(.x, n = N_bins_window, align = "center", fill = NA_real_),
      .names = "{.col}_counts"
    ))

  # long: counts_window per population
  counts_long <- counts_roll_wide %>%
    tidyr::pivot_longer(cols = tidyselect::ends_with("_counts"),
                        names_to = "population", values_to = "counts_window") %>%
    dplyr::mutate(population = sub("_counts$", "", population)) %>%
    dplyr::select(start_interval, interval_number, mid_interval, population, counts_window)

  # grab ctrl counts per window for later filtering
  ctrl_counts <- counts_long %>%
    dplyr::filter(population == "ctrl") %>%
    dplyr::select(interval_number, ctrl_counts_window = counts_window)
  ## -------------------------------------------------------------------------

  # enrichment as before
  enrich_ints_by_bin  = cut_by_intervals_wide %>%
    tidyr::pivot_longer(cols = -c('start_interval','interval_number','mid_interval'),
                        names_to = "population", values_to = 'N_in_window') %>%
    dplyr::left_join(total_ints, by = "population") %>%
    dplyr::mutate(fract_of_ints = N_in_window / total_ints) %>%
    dplyr::select(-c('N_in_window','total_ints')) %>%
    tidyr::pivot_wider(names_from = population, values_from = fract_of_ints) %>%
    dplyr::mutate(dplyr::across(
      -c('start_interval','interval_number','mid_interval'),
      ~ RcppRoll::roll_sum(.x, n = N_bins_window, align = "center", fill = NA_real_)
    )) %>%
    dplyr::mutate(dplyr::across(
      -c('start_interval','interval_number','mid_interval'),
      ~ .x / ctrl
    )) %>%
    tidyr::pivot_longer(-c('start_interval','interval_number','mid_interval'),
                        values_to = 'int_enrich', names_to = 'population')

  # add counts per window for current population + ctrl
  enrich_ints_by_bin %>%
    dplyr::left_join(counts_long, by = c("start_interval","interval_number","mid_interval","population")) %>%
    dplyr::left_join(ctrl_counts, by = "interval_number")
}


enrichm_function_fixed_bins = function(TIB, CL = "CBS16", EXP = all_exp, binsize = 1000, N_bins_window = 10,  calculation_window =  c(34091354, 35744137), MOTIF = c("foward","reverse")){
  enrich_ls = lapply(CL, function(cl){
    enrich_cl = enrichm_function_fixed_bins_CL(TIB, CL = cl, EXP, binsize = binsize, N_bins_window = N_bins_window, calculation_window = calculation_window,  MOTIF)
    enrich_cl
  })
  names(enrich_ls) = CL
  bind_rows(enrich_ls, .id = 'cell_line')
}

enrichm_function_fixed_bins_stranded = function(TIB, CL = "CBS16", EXP = all_exp, binsize = 1000, N_bins_window = 10,  calculation_window =  c(34091354, 35744137)){
  bind_rows("forward" = enrichm_function_fixed_bins(TIB, CL, EXP, binsize = binsize, N_bins_window = N_bins_window, calculation_window = calculation_window, MOTIF = "forward"),
            "reverse" = enrichm_function_fixed_bins(TIB, CL, EXP, binsize = binsize, N_bins_window = N_bins_window, calculation_window = calculation_window, MOTIF = "reverse"), .id = 'motif')}

```

# Plotting data
## Figure 6B - Beeswarm of CBS-Sox2P data
```{r Beeswarm of CBS-Sox2P data Figure 6B}
all_exp = unique(tib_filtered_curated_strict$experiment)

plot_beeswarm_only(TIB = tib_filtered_curated_strict %>% 
                   filter(hopped & chr == "chr3" & experiment %in% all_exp & population == "ctrl") %>% droplevels(),
                   CL = c("23_34A", "23_MK03_C1", "23_MK01_C5", "23_MK02_C2"),
                   PLOT_ANNOT = T,
                   PLOT_CTCF = T,
                   P_RANGE = range_2Mb,
                   POINT_SIZE = 1)

```

## Figure 6C - Beeswarm of CBS only data
```{r Beeswarm of CBS only data - Figure 6C}
all_CL = unique(tib_CBS16_16A_comb$cell_line)
all_exp = unique(tib_CBS16_16A_comb$experiment)

plot_beeswarm_only(TIB = tib_CBS16_16A_comb %>% 
                   filter(hopped & chr == "chr3" & experiment %in% all_exp & population == "ctrl") %>% droplevels(),
                   CL = all_CL,
                   PLOT_ANNOT = T,
                   PLOT_CTCF = T,
                   P_RANGE = range_2Mb,
                   POINT_SIZE = 1)
```

# Calculate hopping bias
## CBS-Sox2P reporters
### Count int. 1MB up/downstream
```{r}
# count 1MB to the right and left of the LP, calculate fraction over total for each

tib_ctrl_fract_tot = tib_filtered_curated_strict %>% filter(population == "ctrl" & launch_pad_location == "-116 kb" & chr == "chr3" & hopped & start >= (start(landingPad_23)-1e6) & start <= start(landingPad_23)+1e6) %>%
  group_by(cell_line) %>% summarise(total_n = n())
  
# coordinate of the launch pad (numeric)
lp_coord <- start(landingPad_23)

# count upstream and downstream separately
tib_ctrl_fract_split <- tib_filtered_curated_strict %>%
  filter(population == "ctrl",
         launch_pad_location == "-116 kb",
         chr == "chr3",
         hopped,
         start >= (lp_coord - 1e6),
         start <= (lp_coord + 1e6)) %>%
  mutate(region = case_when(
    start < lp_coord ~ "upstream",
    start > lp_coord ~ "downstream",
    TRUE ~ "at_LP"
  )) %>%
  group_by(cell_line, region) %>%
  summarise(n = n(), .groups = "drop") %>%
  # join with total counts
  left_join(tib_ctrl_fract_tot, by = "cell_line") %>%
  mutate(frac = n / total_n)

tib_ctrl_fract_split
```

### Function to calculate p-value fisher test
```{r}
# Make a helper function: build a 2x2 table and run Fisher’s test
compare_to_control <- function(df, test_line, control_line = "23_34A") {
  sub <- df %>%
    filter(cell_line %in% c(control_line, test_line)) %>%
    select(cell_line, region, n) %>%
    pivot_wider(names_from = region, values_from = n)
  
  mat <- as.matrix(sub[, c("upstream", "downstream")])
  rownames(mat) <- sub$cell_line
  
  res <- fisher.test(mat)
  tibble(test_line = test_line,
         p_value = res$p.value,
         odds_ratio = res$estimate)
}

# Run for each MK line
results <- map_dfr(c("23_MK01_C5", "23_MK02_C2", "23_MK03_C1"),
                   ~compare_to_control(tib_ctrl_fract_split, .x))

results
```
**Conclusion:**
MK01: odds ratio 2.73 → significantly more downstream integrations than control (p ≈ 1.8e-9).
MK02: odds ratio 0.38 → significantly more upstream integrations than control (p ≈ 1.8e-17).
MK03: odds ratio 2.0 → significantly more downstream integrations than control (p ≈ 9.9e-10).
So compared to 23_34A, the MK lines show skewed distributions, but in opposite directions for MK02

### Figure 6D - CBS_Sox2P Plot results
```{r}
ggplot(tib_ctrl_fract_split,
       aes(x = cell_line, y = frac, fill = region)) +
  geom_col(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = c("upstream" = "skyblue", "downstream" = "orange")) +
  ylab("Fraction of integrations") +
  xlab("Cell line") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r Figure 6D}
library(ggsignif)

# Prep results for plotting (add significance stars)
results <- results %>%
  mutate(sig = case_when(
    p_value < 1e-10 ~ "***",
    p_value < 1e-5  ~ "**",
    p_value < 0.01  ~ "*",
    TRUE            ~ "ns"
  ))

# Fraction barplot
p <- ggplot(tib_ctrl_fract_split,
            aes(x = cell_line, y = frac, fill = region)) +
  geom_col(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = c("upstream" = "skyblue", "downstream" = "orange")) +
  ylab("Fraction of integrations (±1 Mb)") +
  xlab("Cell line") +
  theme_bw() +
  theme(legend.position = "bottom") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Add significance annotations comparing to 23_34A
pCBS_Sox2P_FishTest = 
p + geom_signif(
      comparisons = list(c("23_34A", "23_MK03_C1"),
                         c("23_34A", "23_MK01_C5"),
                         c("23_34A", "23_MK02_C2")),
      annotations = results$p_value ,
      y_position = c(1.05, 1.12, 1.19),  # adjust so stars are nicely spaced above bars
      tip_length = 0.01
    )
pCBS_Sox2P_FishTest
```

### Figure S6G - Fraction of integrations in Sox2-SCR interval
```{r}
tib_ctrl_fract_tot_LP116 = tib_filtered_curated_strict %>% filter(population == "ctrl" & launch_pad_location == "-116 kb" & chr == "chr3" & hopped & start >= (start(landingPad_23)-1e6) & start <= start(landingPad_23)+1e6) %>%
  group_by(cell_line) %>% summarise(total_n = n())

tib_ctrl_fract_Sox2_SCR_LP116 = tib_filtered_curated_strict %>% 
  filter(population == "ctrl" & launch_pad_location == "-116 kb" & chr == "chr3" & hopped & start >= (end(Sox2_gene)) & start <= start(enhancer)) %>%
  group_by(cell_line) %>% summarise(Sox2_SCR_n = n()) %>% left_join(tib_ctrl_fract_tot_LP116) %>%
  mutate(fract_interval = Sox2_SCR_n/total_n)

kable(tib_ctrl_fract_Sox2_SCR_LP116)
```

```{r Figure S6G}
# count 1MB to the left of Sox2 gene, than Sox2-SCR, 1Mb end of SCR, calculate fraction over total
tib_ctrl_fract_tot = tib_filtered_curated_strict %>% filter(population == "ctrl" & launch_pad_location == "-116 kb" & chr == "chr3" & hopped & start >= (start(Sox2_gene)-1e6) & start <= end(enhancer)+1e6) %>%
  group_by(cell_line) %>% summarise(total_n = n())
  
# coordinate of the launch pad (numeric)
Sox2_coord <- start(Sox2_gene)
SCR_coord <- end(enhancer)

# count upstream and downstream separately
tib_ctrl_fract_split <- tib_filtered_curated_strict %>%
  filter(population == "ctrl",
         launch_pad_location == "-116 kb",
         chr == "chr3",
         hopped,
         start >= (Sox2_coord - 1e6),
         start <= (SCR_coord + 1e6)) %>%
  mutate(region = case_when(
    start < Sox2_coord ~ "upstream",
    start > end(Sox2_gene) & start < start(enhancer) ~ "Sox2_SCR",
    start > SCR_coord ~ "downstream",
    TRUE ~ "in_Sox2_or_SCR"
  )) %>%
  group_by(cell_line, region) %>%
  summarise(n = n(), .groups = "drop") %>%
  # join with total counts
  left_join(tib_ctrl_fract_tot, by = "cell_line") %>%
  mutate(frac = n / total_n)

tib_ctrl_fract_split %>% 
  ggplot(aes(x = cell_line, y = frac, fill = region)) +
  geom_bar(stat= "identity") +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = c("upstream" = "skyblue", "downstream" = "orange", "Sox2_SCR" = "purple", "in_Sox2_or_SCR" = "grey")) +
  ylab("Fraction of integrations") +
  xlab("Cell line") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# add % to plot
tib_ctrl_fract_split %>% 
  ggplot(aes(x = cell_line, y = frac, fill = region)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(aes(label = scales::percent(frac, accuracy = 1)),
            position = position_stack(vjust = 0.5),
            color = "black",
            size = 3) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = c(
    "upstream" = "skyblue",
    "downstream" = "orange",
    "Sox2_SCR" = "purple",
    "in_Sox2_or_SCR" = "grey"
  )) +
  ylab("Fraction of integrations (±1 Mb)") +
  xlab("Cell line") +
  theme_bw() +
  theme(legend.position = "bottom") +
  ggtitle("Fraction of integrations from LP-116 kb") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

##CBS only reporters
### Test significance of up/do for ctrl integrations
```{r}
# count 1MB to the right and left of the LP, calculate fraction over total for each
tib_ctrl_fract_tot = tib_CBS16_16A_comb %>% filter(population == "ctrl" & launch_pad_location == "-116 kb" & chr == "chr3" & hopped & start >= (start(landingPad_23)-1e6) & start <= start(landingPad_23)+1e6) %>%
  group_by(cell_line) %>% summarise(total_n = n())
  
# coordinate of the launch pad (numeric)
lp_coord <- start(landingPad_23)

# count upstream and downstream separately
tib_ctrl_fract_split <- tib_CBS16_16A_comb %>%
  filter(population == "ctrl",
         launch_pad_location == "-116 kb",
         chr == "chr3",
         hopped,
         start >= (lp_coord - 1e6),
         start <= (lp_coord + 1e6)) %>%
  mutate(region = case_when(
    start < lp_coord ~ "upstream",
    start > lp_coord ~ "downstream",
    TRUE ~ "at_LP"
  )) %>%
  group_by(cell_line, region) %>%
  summarise(n = n(), .groups = "drop") %>%
  # join with total counts
  left_join(tib_ctrl_fract_tot, by = "cell_line") %>%
  mutate(frac = n / total_n)

tib_ctrl_fract_split
```

### Function to calculate p-value fisher test (counts are quite high so not sure if Fisher test is right choice)
```{r}
# Make a helper function: build a 2x2 table and run Fisher’s test
compare_to_control <- function(df, test_line, control_line = "CRE6") {
  sub <- df %>%
    filter(cell_line %in% c(control_line, test_line)) %>%
    select(cell_line, region, n) %>%
    pivot_wider(names_from = region, values_from = n)
  
  mat <- as.matrix(sub[, c("upstream", "downstream")])
  rownames(mat) <- sub$cell_line
  
  res <- fisher.test(mat)
  tibble(test_line = test_line,
         p_value = res$p.value,
         odds_ratio = res$estimate)
}

# Run for each MK line
results <- map_dfr(c("CBS16_16A", "3xCTCF"),
                   ~compare_to_control(tib_ctrl_fract_split, .x))

results
```
**Conclusion:**
3xCTCF has significantly more downstream integrations than CRE6
CBS16_CRE6 has more upstream than CRE6 

### Figure 6D - CBS - Plot results
```{r}
ggplot(tib_ctrl_fract_split,
       aes(x = cell_line, y = frac, fill = region)) +
  geom_col(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = c("upstream" = "skyblue", "downstream" = "orange")) +
  ylab("Fraction of integrations") +
  xlab("Cell line") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r Figure 6D - CBS}
# Prep results for plotting (add significance stars)
results <- results %>%
  mutate(sig = case_when(
    p_value < 1e-10 ~ "***",
    p_value < 1e-5  ~ "**",
    p_value < 0.01  ~ "*",
    TRUE            ~ "ns"
  ))

# Fraction barplot
p <- ggplot(tib_ctrl_fract_split,
            aes(x = cell_line, y = frac, fill = region)) +
  geom_col(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = c("upstream" = "skyblue", "downstream" = "orange")) +
  ylab("Fraction of integrations (±1 Mb)") +
  xlab("Cell line") +
  theme_bw() +
  theme(legend.position = "bottom") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Add significance annotations comparing to CRE6
pCBS_only_FishTest = 
p + geom_signif(
      comparisons = list(c("CRE6", "3xCTCF"),
                         c("CRE6", "CBS16_16A")),
      annotations = results$p_value,
      y_position = c(1.05, 1.19),  # adjust so stars are nicely spaced above bars
      tip_length = 0.01
    )

pCBS_only_FishTest
```
### Combined figure 6D
```{r combined figure 6D}
plot_grid(pCBS_Sox2P_FishTest, pCBS_only_FishTest, ncol = 2, align = "lb")

```

### Fraction of integrations in Sox2-SCR interval
```{r CBS only}

tib_ctrl_fract_tot_LP116_CBS_only = tib_CBS16_16A_comb %>% filter(population == "ctrl" & launch_pad_location == "-116 kb" & chr == "chr3" & hopped & start >= (start(landingPad_23)-1e6) & start <= start(landingPad_23)+1e6) %>%
  group_by(cell_line) %>% summarise(total_n = n())

tib_ctrl_fract_Sox2_SCR_LP116_CBS_only = tib_CBS16_16A_comb %>% 
  filter(population == "ctrl" & launch_pad_location == "-116 kb" & chr == "chr3" & hopped & start >= (end(Sox2_gene)) & start <= start(enhancer)) %>%
  group_by(cell_line) %>% summarise(Sox2_SCR_n = n()) %>% left_join(tib_ctrl_fract_tot_LP116_CBS_only) %>%
  mutate(fract_interval = Sox2_SCR_n/total_n)

kable(tib_ctrl_fract_Sox2_SCR_LP116_CBS_only)
```

```{r}
# coordinate of the launch pad (numeric)
Sox2_coord <- start(Sox2_gene)
SCR_coord <- end(enhancer)

# count upstream and downstream separately
tib_ctrl_fract_split_CBS_only <- tib_CBS16_16A_comb %>%
  filter(population == "ctrl",
         launch_pad_location == "-116 kb",
         chr == "chr3",
         hopped,
         start >= (Sox2_coord - 1e6),
         start <= (SCR_coord + 1e6)) %>%
  mutate(region = case_when(
    start < Sox2_coord ~ "upstream",
    start > end(Sox2_gene) & start < start(enhancer) ~ "Sox2_SCR",
    start > SCR_coord ~ "downstream",
    TRUE ~ "in_Sox2_or_SCR"
  )) %>%
  group_by(cell_line, region) %>%
  summarise(n = n(), .groups = "drop") %>%
  # join with total counts
  left_join(tib_ctrl_fract_tot, by = "cell_line") %>%
  mutate(frac = n / total_n)

tib_ctrl_fract_split_CBS_only %>% 
  ggplot(aes(x = cell_line, y = frac, fill = region)) +
  geom_bar(stat= "identity") +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = c("upstream" = "skyblue", "downstream" = "orange", "Sox2_SCR" = "purple", "in_Sox2_or_SCR" = "grey")) +
  ylab("Fraction of integrations") +
  xlab("Cell line") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# add % to plot
tib_ctrl_fract_split_CBS_only %>% 
  ggplot(aes(x = cell_line, y = frac, fill = region)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(aes(label = scales::percent(frac, accuracy = 1)),
            position = position_stack(vjust = 0.5),
            color = "black",
            size = 3) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = c(
    "upstream" = "skyblue",
    "downstream" = "orange",
    "Sox2_SCR" = "purple",
    "in_Sox2_or_SCR" = "grey"
  )) +
  ylab("Fraction of integrations (±1 Mb)") +
  xlab("Cell line") +
  theme_bw() +
  theme(legend.position = "bottom") +
  ggtitle("Fraction of integrations from LP-116 kb") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


# Enrichment over non-CBS
## CBS-Sox2P
### Adapted function
```{r}
# Enrichment of cell lines vs a reference cell line
# population is fixed (default "ctrl")

enrich_between_cell_lines_CL <- function(
  TIB,
  REF_CL = "23_34A",
  CL = c("23_34A", "23_MK03_C1", "23_MK01_C5", "23_MK02_C2"),
  POP = "ctrl",
  EXP = all_exp,
  binsize = 1000,
  N_bins_window = 10,
  calculation_window = c(34091354, 35744137)
) {
  # make sure reference is included
  CL_use <- union(REF_CL, CL)

  tib_for_analysis <- TIB %>%
    dplyr::filter(
      hopped,
      !is.na(cell_line), 
      cell_line %in% CL_use,
      experiment %in% EXP,
      population %in% POP
    )

  # total integrations per cell line
  total_ints <- tib_for_analysis %>%
    dplyr::group_by(cell_line) %>%
    dplyr::summarise(total_ints = dplyr::n(), .groups = "drop")

  # bin boundaries
  boundaries_all <- seq(from = calculation_window[1],
                        to   = calculation_window[2],
                        by   = binsize)

  boundaries_all_tib <- tibble::tibble(
    interval_number = seq_along(boundaries_all),
    start_interval  = boundaries_all
  ) %>%
    dplyr::arrange(start_interval) %>%
    dplyr::mutate(
      mid_interval = (start_interval + dplyr::lead(start_interval, 1))/2
    )

  # counts per bin per cell line
  cut_by_intervals <- tib_for_analysis %>%
    dplyr::filter(
      chr == "chr3",
      start >= calculation_window[1],
      start <= calculation_window[2]
    ) %>%
    dplyr::arrange(start) %>%
    dplyr::mutate(interval_number = findInterval(start, boundaries_all)) %>%
    dplyr::group_by(cell_line, interval_number) %>%
    dplyr::summarise(N_ints = dplyr::n(), .groups = "drop")

  # wide: counts per bin (include empty bins)
  cut_by_intervals_wide <- dplyr::left_join(boundaries_all_tib, cut_by_intervals,
                                            by = "interval_number") %>%
    dplyr::mutate(N_ints = tidyr::replace_na(N_ints, 0)) %>%
    dplyr::arrange(interval_number) %>%
    tidyr::pivot_wider(
      id_cols      = c("start_interval", "interval_number", "mid_interval"),
      names_from   = cell_line,
      values_from  = N_ints,
      values_fill  = 0
    ) %>%
    # *** keep only real cell lines, drop any weird `NA` column ***
    dplyr::select(start_interval, interval_number, mid_interval, dplyr::all_of(CL_use))

  ## ---- rolling raw counts per window (for later filtering) ----
  counts_roll_wide <- cut_by_intervals_wide %>%
    dplyr::mutate(dplyr::across(
      -c(start_interval, interval_number, mid_interval),
      ~ RcppRoll::roll_sum(.x, n = N_bins_window,
                           align = "center", fill = NA_real_),
      .names = "{.col}_counts"
    ))

  counts_long <- counts_roll_wide %>%
    tidyr::pivot_longer(
      cols      = tidyselect::ends_with("_counts"),
      names_to  = "cell_line",
      values_to = "counts_window"
    ) %>%
    dplyr::mutate(cell_line = sub("_counts$", "", cell_line))

  ctrl_counts <- counts_long %>%
    dplyr::filter(cell_line == REF_CL) %>%
    dplyr::select(interval_number, ctrl_counts_window = counts_window)

  ## ---- enrichment vs reference cell line -----------------------
  enrich_ints_by_bin <- cut_by_intervals_wide %>%
    tidyr::pivot_longer(
      cols      = -c("start_interval", "interval_number", "mid_interval"),
      names_to  = "cell_line",
      values_to = "N_in_window"
    ) %>%
    dplyr::left_join(total_ints, by = "cell_line") %>%
    dplyr::mutate(fract_of_ints = N_in_window / total_ints) %>%
    dplyr::select(-c("N_in_window", "total_ints")) %>%
    tidyr::pivot_wider(
      names_from  = "cell_line",
      values_from = "fract_of_ints"
    ) %>%
    # smooth: rolling sum of fractions per window
    dplyr::mutate(dplyr::across(
      -c("start_interval", "interval_number", "mid_interval"),
      ~ RcppRoll::roll_sum(.x, n = N_bins_window,
                           align = "center", fill = NA_real_)
    )) %>%
    # divide each cell line by the reference cell line
    dplyr::mutate(dplyr::across(
      -c("start_interval", "interval_number", "mid_interval"),
      ~ .x / .data[[REF_CL]]
    )) %>%
    tidyr::pivot_longer(
      -c("start_interval", "interval_number", "mid_interval"),
      names_to  = "cell_line",
      values_to = "int_enrich"
    )

  # add window counts (per cell line + ref)
  enrich_ints_by_bin %>%
    dplyr::left_join(counts_long,
                     by = c("start_interval", "interval_number",
                            "mid_interval", "cell_line")) %>%
    dplyr::left_join(ctrl_counts, by = "interval_number")
}

```
## Plotting 
### CBS-Sox2P
```{r}
# Calculate enrichment of CBS-Sox2P reporters over Sox2P only. 
all_exp = unique(tib_filtered_curated_strict$experiment)

enrich_23_ctrl_CL <- enrich_between_cell_lines_CL(
  TIB = tib_filtered_curated_strict,
  REF_CL = "23_34A",
  CL = c("23_34A", "23_MK03_C1", "23_MK01_C5", "23_MK02_C2"),
  POP = "ctrl",
  EXP = all_exp,
  binsize = 2000,
  N_bins_window = 10,
  calculation_window = range_2Mb)

# make a plotting column that is NA where we don't trust the value
enrich_23_ctrl_CL_plot <- enrich_23_ctrl_CL %>%
  mutate(
    int_enrich_plot = if_else(
      cell_line != "23_34A" &
        ctrl_counts_window >= 2 &
        counts_window      >= 2 &
        is.finite(int_enrich),
      int_enrich,
      NA_real_
    )
  )

# Quick plot: enrichment vs reference
enrich_23_ctrl_CL_plot %>%
  filter(cell_line != "23_34A") %>%
  ggplot(aes(x = mid_interval / 1e6,
             y = int_enrich_plot,
             colour = cell_line,
             group  = cell_line)) +
  geom_line(na.rm = TRUE) +          # NA -> gaps, not removed rows
  #geom_point() +
  geom_hline(yintercept = 1, linetype = "dashed") +
  xlab("Genomic coordinate (Mb)") +
  ylab("Enrichment over 23_34A (ctrl)") +
  theme_classic()

# with annotations
enrich_23_ctrl_CL_plot %>%
  filter(cell_line != "23_34A") %>%
  ggplot(aes(x = mid_interval,
             y = int_enrich_plot,
             colour = cell_line,
             group  = cell_line)) +
  #geom_point(na.rm = TRUE) +
  geom_line(na.rm = TRUE) +
  facet_wrap(~ cell_line) +
  # enhancer
  annotate("rect", xmin = start(enhancer), xmax = end(enhancer),
           ymin = -Inf, ymax = Inf, alpha = .2, fill = brown) +
  geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown) +
  # gene
  annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene),
           ymin = -Inf, ymax = Inf, alpha = .2, fill = "red") +
  geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = "red") +
  # CTCFs
  geom_vline(xintercept = start(CTCF_mm10.chr3),
             col = "darkgrey", linetype = "dotted") +
  geom_vline(xintercept = 34772210,
             col = "darkgrey", linetype = "dotted") +
  # landing pad
  geom_vline(xintercept = start(landingPad_23), col = "darkgrey") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_x_continuous(
    breaks  = scales::pretty_breaks(n = 6),
    labels  = scales::unit_format(scale = 1e-6, accuracy = 0.1, unit = NULL),
    limits  = range_2Mb,
    expand=c(0,0)
  ) +
  xlab("Genomic coordinate (Mb)") +
  ylab("Enrichment over 23_34A (ctrl)") +
  theme_classic()

```

### CBS-only
```{r}
# Calculate enrichment of CBS-Sox2P reporters over Sox2P only. 
all_exp = unique(tib_CBS16_16A_comb$experiment)

enrich_CBS_ctrl_CL <- enrich_between_cell_lines_CL(
  TIB = tib_CBS16_16A_comb,
  REF_CL = "CRE6",
  CL = c("CRE6", "CBS16_16A", "3xCTCF"),
  POP = "ctrl",
  EXP = all_exp,
  binsize = 2000,
  N_bins_window = 10,
  calculation_window = range_2Mb)

# make a plotting column that is NA where we don't trust the value
enrich_CBS_ctrl_CL_plot <- enrich_CBS_ctrl_CL %>%
  mutate(
    int_enrich_plot = if_else(
      cell_line != "CRE6" &
        ctrl_counts_window >= 3 &
        counts_window      >= 3 &
        is.finite(int_enrich),
      int_enrich,
      NA_real_
    )
  )

# Quick plot: enrichment vs reference
enrich_CBS_ctrl_CL_plot %>%
  filter(cell_line != "CRE6") %>%
  ggplot(aes(x = mid_interval / 1e6,
             y = int_enrich_plot,
             colour = cell_line,
             group  = cell_line)) +
  geom_line(na.rm = TRUE) +          # NA -> gaps, not removed rows
  #geom_point() +
  geom_hline(yintercept = 1, linetype = "dashed") +
  xlab("Genomic coordinate (Mb)") +
  ylab("Enrichment over 23_34A (ctrl)") +
  theme_classic()

# with annotations
enrich_CBS_ctrl_CL_plot %>%
  filter(cell_line != "CRE6") %>%
  ggplot(aes(x = mid_interval,
             y = int_enrich_plot,
             colour = cell_line,
             group  = cell_line)) +
  #geom_point(na.rm = TRUE) +
  geom_line(na.rm = TRUE) +
  facet_wrap(~ cell_line) +
  # enhancer
  annotate("rect", xmin = start(enhancer), xmax = end(enhancer),
           ymin = -Inf, ymax = Inf, alpha = .2, fill = brown) +
  geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown) +
  # gene
  annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene),
           ymin = -Inf, ymax = Inf, alpha = .2, fill = "red") +
  geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = "red") +
  # CTCFs
  geom_vline(xintercept = start(CTCF_mm10.chr3_extra),
             col = "darkgrey", linetype = "dotted") +
  # landing pad
  geom_vline(xintercept = start(landingPad_23), col = "darkgrey") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_x_continuous(
    breaks  = scales::pretty_breaks(n = 6),
    labels  = scales::unit_format(scale = 1e-6, accuracy = 0.1, unit = NULL),
    limits  = range_2Mb
  ) +
  xlab("Genomic coordinate (Mb)") +
  ylab("Enrichment over 23_34A (ctrl)") +
  theme_classic()

```


# Bootstrap function
## CBS-Sox2P
```{r}
cl_23 <- c("23_34A", "23_MK03_C1", "23_MK01_C5", "23_MK02_C2")

tib_boot <- tib_filtered_curated_strict %>%
  filter(
    population == "ctrl",
    chr == "chr3",
    cell_line %in% cl_23
  ) %>%
  ungroup()

```

```{r}
set.seed(456)
n_boot <- 5000
all_exp = unique(tib_filtered_curated_strict$experiment)

boot_enrich_list <- map(seq_len(n_boot), function(bi) {
  cat("▶ Bootstrap", bi, "\n")
  
  # resample integrations with replacement
  boot_dat <- tib_boot %>%
  group_by(cell_line) %>%
  slice_sample(prop = 1, replace = TRUE) %>%
  ungroup()
  
  # enrichment of each cell line vs 23_34A
  enrich_between_cell_lines_CL(
    TIB                = boot_dat,
    REF_CL             = "23_34A",
    CL                 = cl_23,
    POP                = "ctrl",
    EXP                = all_exp,
    binsize            = 2000,
    N_bins_window      = 10,
    calculation_window = range_2Mb
  ) %>%
    mutate(bootstrap = bi)
})

# Save object 
saveRDS(boot_enrich_list, "/DATA/projects/CTCF_Sox2/analysis/Enrichment/ME20260303_enrichment_CL_Sox2P_vs_CBS_Sox2P_5000boot_2000_10.rds")
boot_enrich_list = readRDS("/DATA/projects/CTCF_Sox2/analysis/Enrichment/ME20260303_enrichment_CL_Sox2P_vs_CBS_Sox2P_5000boot_2000_10.rds")



boot_enrich_df <- bind_rows(boot_enrich_list)

# sanity check: same number of rows per bootstrap
boot_enrich_df %>% count(bootstrap)

```

### mask bad windows in the bootsrap, not drop them NAs
```{r}
min_ctrl <- 3   
min_test <- 3

boot_enrich_masked <- boot_enrich_df %>%
  filter(!is.na(cell_line),
    cell_line != "23_34A") %>%
  mutate(
    int_enrich_masked = if_else(
      ctrl_counts_window >= min_ctrl &
      counts_window      >= min_test &
      is.finite(int_enrich),
      int_enrich,
      NA_real_
    )
  )
```

### summarise and keep NA if there is no information in bin
```{r}

ci_df <- boot_enrich_masked %>%
  group_by(cell_line, mid_interval) %>%
  summarize(
    lower  = if (all(is.na(int_enrich_masked))) NA_real_
             else quantile(int_enrich_masked, 0.025, na.rm = TRUE),
    median = if (all(is.na(int_enrich_masked))) NA_real_
             else quantile(int_enrich_masked, 0.5,   na.rm = TRUE),
    upper  = if (all(is.na(int_enrich_masked))) NA_real_
             else quantile(int_enrich_masked, 0.975, na.rm = TRUE),
    .groups = "drop"
  )
```

### make sure all windows exist as rows
```{r}

all_mids <- sort(unique(boot_enrich_df$mid_interval))

#ci_df only contains those mid_intervals where at leasat one boostrap replicate produced something, make sure to have an NA row for every genomic bin
ci_df_full <- ci_df %>% 
  tidyr::complete(cell_line, mid_interval = all_mids)
```

### Figure S5F - Plot
```{r Figure S5F}
#combine with real data
tib_try = enrich_23_ctrl_CL_plot %>% select(mid_interval, cell_line, int_enrich)
tib_comb = ci_df_full %>% left_join(tib_try) %>%
  mutate(cell_line = factor(cell_line, levels = c("23_34A", "23_MK03_C1", "23_MK01_C5", "23_MK02_C2")))

pEnrich_lin_CBS_Sox2P =
ggplot(tib_comb,
       aes(x = mid_interval,
           y = median,
           colour = cell_line,
           fill   = cell_line,
           group  = cell_line)) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  geom_ribbon(aes(ymin = lower, ymax = upper),
              alpha = 0.2, colour = NA) +
  geom_line(linewidth = 0.8, na.rm = TRUE) +
  #geom_line(aes(y = int_enrich), col = "orange") +
  facet_grid(cell_line ~ ., scales = "free_y") +
  geom_vline(xintercept = start(CTCF_mm10.chr3),
             col = "darkgrey", linetype = 'dotted') +
  geom_vline(xintercept = 34772210,
             col = 'darkgrey', linetype = "dotted") +
  # annotate enhancer
  annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown) +
  geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown) +
  # annotate gene
  annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf, alpha = .2, fill = 'red') +
  geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = 'red') +
  scale_x_continuous(
    breaks  = scales::pretty_breaks(n = 6), 
    labels  = scales::unit_format(scale = 1E-6, accuracy = 0.1, unit = NULL),
    limits  = range_2Mb,
    expand=c(0,0)
  ) +
  xlab("Genomic coordinate (Mb)") +
  ylab("enrichment vs 23_34A (ctrl)") +
  theme_classic() +
  theme(legend.position = "bottom") +
  col_scale_MK_CL

pEnrich_lin_CBS_Sox2P
```

## CBS only
```{r}
cl_CBS <-  c("CRE6", "CBS16_16A", "3xCTCF")
  
tib_boot <- tib_CBS16_16A_comb %>%
  filter(
    population == "ctrl",
    chr == "chr3",
    cell_line %in% cl_CBS
  ) %>%
  ungroup()
```

```{r}
set.seed(456)
n_boot <- 5000
all_exp = unique(tib_CBS16_16A_comb$experiment)

boot_enrich_list <- map(seq_len(n_boot), function(bi) {
  cat("▶ Bootstrap", bi, "\n")
  
  # resample integrations with replacement
  boot_dat <- tib_boot %>%
  group_by(cell_line) %>%
  slice_sample(prop = 1, replace = TRUE) %>%
  ungroup()
  
  # enrichment of each cell line vs 23_34A
  enrich_between_cell_lines_CL(
    TIB                = boot_dat,
    REF_CL             = "CRE6",
    CL                 = cl_CBS,
    POP                = "ctrl",
    EXP                = all_exp,
    binsize            = 2000,
    N_bins_window      = 10,
    calculation_window = range_2Mb
  ) %>%
    mutate(bootstrap = bi)
})

# Save object 
saveRDS(boot_enrich_list, "/DATA/projects/CTCF_Sox2/analysis/Enrichment/ME20260303_enrichment_CL_deltaCBS_vs_CBS_5000boot_2000_10.rds")
boot_enrich_list = readRDS("/DATA/projects/CTCF_Sox2/analysis/Enrichment/ME20260303_enrichment_CL_deltaCBS_vs_CBS_5000boot_2000_10.rds")


boot_enrich_df <- bind_rows(boot_enrich_list)

# sanity check: same number of rows per bootstrap
boot_enrich_df %>% count(bootstrap)

```

### mask bad windows in the bootsrap, not drop them NAs
```{r}
min_ctrl <- 3   
min_test <- 3

boot_enrich_masked <- boot_enrich_df %>%
  filter(!is.na(cell_line),
    cell_line != "CRE6") %>%
  mutate(
    int_enrich_masked = if_else(
      ctrl_counts_window >= min_ctrl &
      counts_window      >= min_test &
      is.finite(int_enrich),
      int_enrich,
      NA_real_
    )
  )
```

### summarise and keep NA if there is no information in bin
```{r}

ci_df <- boot_enrich_masked %>%
  group_by(cell_line, mid_interval) %>%
  summarize(
    lower  = if (all(is.na(int_enrich_masked))) NA_real_
             else quantile(int_enrich_masked, 0.025, na.rm = TRUE),
    median = if (all(is.na(int_enrich_masked))) NA_real_
             else quantile(int_enrich_masked, 0.5,   na.rm = TRUE),
    upper  = if (all(is.na(int_enrich_masked))) NA_real_
             else quantile(int_enrich_masked, 0.975, na.rm = TRUE),
    .groups = "drop"
  )
```

### make sure all windows exist as rows
```{r}

all_mids <- sort(unique(boot_enrich_df$mid_interval))

#ci_df only contains those mid_intervals where at leasat one boostrap replicate produced something, make sure to have an NA row for every genomic bin
ci_df_full <- ci_df %>% 
  tidyr::complete(cell_line, mid_interval = all_mids)

```

### Figure 6E Plot
```{r Figure 6E}
#combine with real data
tib_try_CBS = enrich_CBS_ctrl_CL_plot %>% select(mid_interval, cell_line, int_enrich)
tib_comb_CBS = ci_df_full %>% left_join(tib_try_CBS) %>%
  mutate(cell_line = factor(cell_line, levels = c("CBS16_16A", "3xCTCF")))

pEnrich_lin_CBS_only =
ggplot(tib_comb_CBS,
       aes(x = mid_interval,
           y = median,
           colour = cell_line,
           fill   = cell_line,
           group  = cell_line)) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  geom_ribbon(aes(ymin = lower, ymax = upper),
              alpha = 0.2, colour = NA) +
  geom_line(linewidth = 0.8, na.rm = TRUE) +
  #geom_line(aes(y = int_enrich), col = "orange") +
  facet_grid(cell_line ~ ., scales = "free_y") +
  #Annotate CTCF sites
  # geom_vline(xintercept = start(CTCF_mm10.chr3),
  #            col = 'darkgrey', linetype = 'dotted') +
  # geom_vline(xintercept = 34772210,
  #            col = 'darkgrey', linetype = "dotted") +
  #annotate deWit CTCF sites
  geom_vline(xintercept = start(CTCF_mm10.chr3_extra),
             col = 'darkgrey', linetype = 'dotted') +
  
  # annotate enhancer
  annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown) +
  geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown) +
  # annotate gene
  annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf, alpha = .2, fill = 'red') +
  geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = 'red') +
  scale_x_continuous(
    breaks  = scales::pretty_breaks(n = 6), 
    labels  = scales::unit_format(scale = 1E-6, accuracy = 0.1, unit = NULL),
    limits  = range_2Mb,
    expand=c(0,0)
  ) +
  xlab("Genomic coordinate (Mb)") +
  ylab("enrichment vs deltaCBS (ctrl)") +
  theme_classic() +
  theme(legend.position = "bottom")

pEnrich_lin_CBS_only
```

#### Add CBS as triangles to plot
```{r}
# CTCF positions as a tibble
ctcf_tri_df <- tibble(
  mid    = (start(CTCF_mm10.chr3) + end(CTCF_mm10.chr3))/2,             # start of motif; or (start+end)/2
  strand = as.character(strand(CTCF_mm10.chr3))
) %>%
  filter(
    mid >= range_2Mb[1],
    mid <= range_2Mb[2]
  )

# CTCF panel data: put everything into the "CTCF" facet, y = 1 (arbitrary)
ctcf_tri_df <- ctcf_tri_df %>%
  mutate(
    cell_line = factor("CTCF", levels = levels(ci_df_full$cell_line)),
    y         = 1
  )

#Plot
pEnrich_lin_CBS_only_CBS_annot =
ggplot(ci_df_full,
       aes(x = mid_interval,
           y = median,
           colour = cell_line,
           fill   = cell_line,
           group  = cell_line)) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  geom_ribbon(aes(ymin = lower, ymax = upper),
              alpha = 0.2, colour = NA) +
  geom_line(linewidth = 0.8, na.rm = TRUE) +

  ## --- NEW: CTCF orientation track on top ------------------------
  geom_point(
    data = ctcf_tri_df,
    aes(x = mid, y = y, shape = strand),
    inherit.aes = FALSE,
    size = 4,
    colour = "black"
  ) +
  # "+" = triangle pointing right, "-" = triangle pointing left
  scale_shape_manual(values = c("+" = 62, "-" = 60)) +
  guides(shape = "none") +
  ## ---------------------------------------------------------------

  facet_grid(cell_line ~ ., scales = "free_y") +
  # keep your gene / enhancer / extra CTCF vlines if you like,
  # but you probably don't need the old CTCF vlines anymore:
   geom_vline(xintercept = start(CTCF_mm10.chr3),
              col = 'darkgrey', linetype = 'dotted') +
  geom_vline(xintercept = 34772210,
             col = 'darkgrey', linetype = "dotted") +
  # annotate enhancer
  annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown) +
  geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown) +
  # annotate gene
  annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf, alpha = .2, fill = 'red') +
  geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = 'red') +
  scale_x_continuous(
    breaks  = scales::pretty_breaks(n = 6), 
    labels  = scales::unit_format(scale = 1E-6, accuracy = 0.1, unit = NULL),
    limits  = range_2Mb,
    expand=c(0,0)
  ) +
  xlab("Genomic coordinate (Mb)") +
  ylab("enrichment vs deltaCBS (ctrl)") +
  theme_classic() +
  theme(legend.position = "bottom")

pEnrich_lin_CBS_only_CBS_annot


# Arrow annotation
pEnrich_lin_CBS_only_CBS_annot <-
  ggplot(ci_df_full,
         aes(x = mid_interval,
             y = median,
             colour = cell_line,
             fill   = cell_line,
             group  = cell_line)) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  geom_ribbon(aes(ymin = lower, ymax = upper),
              alpha = 0.2, colour = NA) +
  geom_line(linewidth = 0.8, na.rm = TRUE) +

  ## --- NEW: CTCF orientation track with unicode triangles ------
  geom_text(
    data = ctcf_tri_df,
    aes(
      x = mid,
      y = y,
      label = ifelse(strand == "+", "\u25B6", "\u25C0")  # ▶ or ◀
    ),
    inherit.aes = FALSE,
    colour = "black",
    size = 4
  ) +
  ## -------------------------------------------------------------

  facet_grid(cell_line ~ ., scales = "free_y") +
  geom_vline(xintercept = start(CTCF_mm10.chr3),
             col = "darkgrey", linetype = "dotted") +
  geom_vline(xintercept = 34772210,
             col = "darkgrey", linetype = "dotted") +
  annotate("rect", xmin = start(enhancer), xmax = end(enhancer),
           ymin = -Inf, ymax = Inf, alpha = .2, fill = brown) +
  geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown) +
  annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene),
           ymin = -Inf, ymax = Inf, alpha = .2, fill = "red") +
  geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = "red") +
  scale_x_continuous(
    breaks  = scales::pretty_breaks(n = 6),
    labels  = scales::unit_format(scale = 1e-6, accuracy = 0.1, unit = NULL),
    limits  = range_2Mb,
    expand=c(0,0)
  ) +
  xlab("Genomic coordinate (Mb)") +
  ylab("enrichment vs deltaCBS (ctrl)") +
  theme_classic() +
  theme(legend.position = "bottom")

pEnrich_lin_CBS_only_CBS_annot

#ggsave(filename='Enrich_lin_CBS_only_CBS_annot.pdf', plot = pEnrich_lin_CBS_only_CBS_annot,  device=cairo_pdf)
```

#Plot all together
```{r}
#Figure S5F
pEnrich_lin_CBS_Sox2P

#Figure 6E
pEnrich_lin_CBS_only
pEnrich_lin_CBS_only_CBS_annot
```
