---
title: "me20250708_FACS_analysis_MK_competition"
author: "Mathias"
date: "2025-07-08"
output: 
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---


# Aim 
In E2252 we flipped MK01, MK02, MK03 into LP-39kb (between Sox2 and SCR) and LP-116kb (5kb upstream of Sox2::mCherry) to see the effect on reporter and mCherry expression.
We transfected, selected, picked colonies, genotyped them by PCR and measured them by FACS. 

Here i want to display the Fluo values of the samples.

# Method
- gated fcs files in flowjo for single cells
- import data
- correct for autofluorescnece 
- normalize to 23_34A
- plot reporter and Sox2::mCherry


#Compatibility problems
The newer ggplot version is not compatible with the plotting in the ggcyto package, therefore i loaded an older version of ggplot in order to be compatible with the functions of ggcyto.
```{r}
# Load older ggplot version that is compatible with the ggcyto package
.libPaths(c("/DATA/usr/m.eder/R/ggplot2_3x", .libPaths()))

library(ggplot2)
packageVersion("ggplot2")  # should now show 3.5.1
```

# Set-up
##Libraries
Load the libraries and set the parameters.
```{r}

# Load dependencies
library(GenomicRanges)
library(rtracklayer)
library(tidyverse)
# library(dtplyr)
library(dplyr)
# library(tidyr)
library(ggbeeswarm)
library(ggplot2)
library(Biostrings)
library(stringr)
library(readr)
library(knitr)
library(RColorBrewer)
library(RcppRoll)
# library(plotly)
library(readxl)
library(ggpubr)
library(ggpmisc)
library(ggridges)
library(import)
library(ggh4x)
library(rstatix)
# library(flowCore)
import::from(flowCore, .except=c("filter")) #I don't load the whole flowCore and ggcyto libraries, because they overwrite dplyr's filter function :0
import::from(ggcyto, 'fortify_fs')
library(flowDensity)
library(cowplot)
library(Signac)

split_string <- function(vect,sep,N1,N2=N1){
  library(stringr)
  sapply(vect, function(X){
    paste(str_split(X,sep)[[1]][N1:N2],collapse = sep)
  },USE.NAMES = F)
}
```


```{r paths_parameters}
#datatag
datetag = paste0("ME",format(Sys.time(), '%Y%m%d'))

# Prepare output 
output_dir <- paste0("/DATA/projects/CTCF_Sox2/Scripts/for_figures/Figure_3/analysis_FACS_CBS_Sox2P_", datetag)
dir.create(output_dir, showWarnings = FALSE)

library(knitr)
opts_chunk$set(
               cache = T,
               message = F, 
               warning = F,
               dev=c('png', 'pdf'), 
               dpi = 600,
               fig.path = file.path(output_dir, "figures/")) 
pdf.options(useDingbats = FALSE)
```

## Color scheme
```{r color_settings}

myCols_insert_cell_lines <- c(
  `MK01` = '#1b9e77',   # Teal green
  `MK02` = '#d95f02',   # Strong orange
  `MK03` = '#7570b3',   # Purple
  `23_34A` = 'black', # 
  `8_34_8` = 'black', # 
  `WT` = 'black'        # Black for wild type
)

col_scale_MK_cl <- scale_colour_manual(
  name = "cell_line", 
  values = myCols_insert_cell_lines,
  aesthetics = c("colour", "fill")
)


## SCR
brown = "#ffb000"
```

## Load fcs files
```{r}
path_fcs_files_E2252 = "/DATA/projects/CTCF_Sox2/FACS_files/export_fcs_R_E2252"

fcs_files_E2252 = dir(path_fcs_files_E2252, full.names = T)
fcs_files_E2252_nms = str_remove(fcs_files_E2252, "^.*/")

fcs_files_E2252_nms2 = split_string(fcs_files_E2252_nms, "_", 6, 8)
names(fcs_files_E2252) = fcs_files_E2252_nms2

flowset_E2252 = read.flowSet(files = fcs_files_E2252, alter.names = T, truncate_max_range = F)
sampleNames(flowset_E2252) = names(fcs_files_E2252)
pData(flowset_E2252)$name = names(fcs_files_E2252)

autoplot(flowset_E2252[1:10], 'FSC.A', 'SSC.A')
autoplot(flowset_E2252[1:10], 'BL.B..530_30.A', 'YG.D..610_20.A')
autoplot(flowset_E2252[1:10], 'BL.B..530_30.A', 'V.F..450_50.A')
# 
flowset_fluo_E2252 = flowset_E2252[,c('BL.B..530_30.A', 'YG.D..610_20.A', 'V.F..450_50.A', 'FSC.A', 'SSC.A')]
colnames(flowset_fluo_E2252) = c("GFP", "mCherry", "mTurq", 'FSC', 'SSC')
```

## Annotation
```{r}

tib_annot_E2252 = tibble(sample_name = names(fcs_files_E2252)) %>%
  mutate(cell_line = case_when(grepl("_MK1", sample_name) ~ "MK01",
                               grepl("_MK2", sample_name) ~ "MK02",
                               grepl("_MK3", sample_name) ~ "MK03",
                               grepl("23_34A", sample_name) ~ "23_34A",
                               grepl("8_34_8", sample_name) ~ "8_34_8",
                               grepl("F121_9_002", sample_name) ~ "F121",
                               grepl("WT", sample_name) ~ "WT"),
         landing_pad_location = case_when(grepl("23_", sample_name) ~ "LP -116 kb",
                                          grepl("8_", sample_name) ~ "LP -39 kb",
                                          .default = "none"),
         LP_CL = paste0(cell_line, "_", landing_pad_location),
         clone_tmp = split_string(sample_name, "_", 3),
         clone = case_when(clone_tmp == "003" ~ "34A",
                           clone_tmp == "002" ~ "ctrl",
                           clone_tmp == "Single Cells.fcs" ~ "ctrl",
                           .default = clone_tmp),
         insert_name = case_when(cell_line == "MK01" ~ "CBS3x_Fw_Sox2P",
                                 cell_line == "MK02" ~ "CBS3x_Rv_Sox2P",
                                 cell_line == "MK03" ~ "CBS1x_Fw_Sox2P",
                                 cell_line == "23_34A" ~ "Sox2P",
                                 cell_line == "8_34_8" ~ "Sox2P",
                                 cell_line == "F121" ~ "none",
                                 .default = cell_line),
          experiment = "E2252") %>%
  select(-clone_tmp)


pheno_data_tib_E2252 = left_join(pData(flowset_fluo_E2252), tib_annot_E2252, join_by(name == sample_name) ) 
rownames(pheno_data_tib_E2252) = pheno_data_tib_E2252$name

pData(flowset_fluo_E2252) = pheno_data_tib_E2252
```

## calculate expression
For now I calculate the median expression without subtracting any autofluorescence
```{r median_E2252}

total_cell_count_tmp = flowCore::keyword(flowset_fluo_E2252, "$TOT")[,'$TOT']
total_cell_count = as.numeric(total_cell_count_tmp)
names(total_cell_count) = names(total_cell_count_tmp)

median_mTurq = sapply(1:length(flowset_fluo_E2252), function(i){
  median(flowCore::exprs(flowset_fluo_E2252[[i]])[,'mTurq'])
})
names(median_mTurq) = sampleNames(flowset_fluo_E2252)

median_mCherry = sapply(1:length(flowset_fluo_E2252), function(i){
  median(flowCore::exprs(flowset_fluo_E2252[[i]])[,'mCherry'])
})
names(median_mCherry) = sampleNames(flowset_fluo_E2252)


median_GFP = sapply(1:length(flowset_fluo_E2252), function(i){
  median(flowCore::exprs(flowset_fluo_E2252[[i]])[,'GFP'])
})
names(median_GFP) = sampleNames(flowset_fluo_E2252)

flowCore::pData(flowset_fluo_E2252)$median_mTurq = median_mTurq
flowCore::pData(flowset_fluo_E2252)$median_mCherry = median_mCherry
flowCore::pData(flowset_fluo_E2252)$median_GFP = median_GFP

flowCore::pData(flowset_fluo_E2252)$total_cell_count = total_cell_count

fcs_tib = tibble(flowCore::pData(flowset_fluo_E2252))
```

# Normalize expression
```{r normalize_expression_values}
#subtract autofluorescence, normalize to GFP
autofluo_tib_comb = fcs_tib  %>%
  filter(cell_line == "WT") %>%
  summarize(mTurq_WT = mean(median_mTurq),
           mCherry_WT = mean(median_mCherry),
            GFP_WT = mean(median_GFP) )

expression_tib_comb = fcs_tib %>%
  bind_cols(autofluo_tib_comb) %>%
  mutate(median_mTurq_cor = median_mTurq - mTurq_WT,
         median_mCherry_cor = median_mCherry - mCherry_WT,
         median_GFP_cor = median_GFP - GFP_WT) %>%
  mutate(mTurq_norm = median_mTurq_cor / median_GFP_cor,
         mCherry_norm = median_mCherry_cor / median_GFP_cor)

#normalize to 23_34A (to compare between measurement days)
standard_tib = expression_tib_comb %>%
  filter(cell_line == "23_34A") %>%
  summarize(mTurq_norm_standard = mean(mTurq_norm),
           mCherry_norm_standard = mean(mCherry_norm),
            GFP_standard = mean(median_GFP_cor) )

expression_tib_comb = expression_tib_comb %>%
  bind_cols(standard_tib) %>%
  mutate(rel_mTurq = mTurq_norm / mTurq_norm_standard,
         rel_mCherry = mCherry_norm / mCherry_norm_standard,
         rel_GFP_cor = median_GFP_cor / GFP_standard)

stat_table = 
expression_tib_comb %>% 
  group_by(cell_line, landing_pad_location) %>% 
  summarise(mean_mCh = mean(rel_mCherry),
            mean_mTurq = mean(rel_mTurq))
``` 
**Conclusion:**
maximum insulation in LP-39kb with MK01, rel_mean_mCherry (normalized to GFP and 23_34A) is 0.65, that means ~ 34% reduction in Sox2::mCherry expression. Sox2P only does not lead to such a reduction, actually has a bit more than 23_34A (1.05).


## Plot clones
```{r}
filter(expression_tib_comb, total_cell_count >= 500) %>%
  #plotting
  ggplot(aes(x = LP_CL, y = rel_mTurq, col = cell_line)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(position = position_jitterdodge(jitter.width = 0.8), size = 2) +
  geom_hline(yintercept = 1, linetype = "dashed") + 
  geom_hline(yintercept = 0) + 
  facet_nested(. ~ landing_pad_location  , scales = "free_x", space = "free_x") +
  #layout
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none",
        axis.title.x = element_blank()) +
  ggtitle("mTurq, flipping clones E2252") +
  ylab("relative reporter expression") +
  col_scale_MK_cl
  

# Relative mCherry
filter(expression_tib_comb, total_cell_count >= 500) %>%
  #plotting
  ggplot(aes(x = cell_line, y = rel_mCherry, col = cell_line)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(position = position_jitterdodge(jitter.width = 0.8), size = 2) +
  geom_hline(yintercept = 1, linetype = "dashed") + 
  geom_hline(yintercept = 0) + 
  facet_nested(. ~ landing_pad_location  , scales = "free_x", space = "free_x") +
  #layout
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none",
        axis.title.x = element_blank()) +
  ggtitle("mCherry, flipping clones E2252") +
  ylab("relative Sox2::mCherry expression") + 
  col_scale_MK_cl
  
```

## Exclude WT and F1212 ctrls from plot
```{r}
## Reorder samples
order <- c("WT","F121","23_34A","8_34_8","MK03","MK01","MK02")

pClones_mT_LP = 
filter(expression_tib_comb, total_cell_count >= 500, landing_pad_location != "none") %>%
  mutate(cell_line = fct_relevel(cell_line, order[order %in% cell_line])) %>%  # avoid warnings
  ggplot(aes(x = cell_line, y = rel_mTurq, col = cell_line)) +
  geom_jitter(size = 2, width = 0.2) +
  stat_summary(fun = "mean", geom = "crossbar", width = 0.5, color = "black", fatten = 2) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  geom_hline(yintercept = 0) +
  facet_wrap(~ landing_pad_location, scales = "free") +   # <— free x *and* y
  scale_x_discrete(drop = TRUE) +                         # (default, but explicit)
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        legend.position = "none",
        axis.title.x = element_blank()) +
  ggtitle("mTurq, flipping clones E2252") +
  ylab("relative reporter expression") +
  col_scale_MK_cl


pClones_mCh_LP = 
filter(expression_tib_comb, total_cell_count >= 500 & landing_pad_location != "none") %>%
   mutate(cell_line = fct_relevel(cell_line, order[order %in% cell_line])) %>%  # avoid warnings
  #plotting
  ggplot(aes(x = cell_line, y = rel_mCherry, col = cell_line)) +
  geom_jitter(size = 2, width = 0.2) +
  # add mean line 
  stat_summary(fun = "mean", geom = "crossbar", 
               width = 0.5, color = "black",  
               aes(y = rel_mCherry), fatten = 2) +
  
  geom_hline(yintercept = 1, linetype = "dashed") + 
  geom_hline(yintercept = 0) + 
  facet_wrap(~ landing_pad_location, scales = "free") +   # <— free x *and* y
  scale_x_discrete(drop = TRUE) +                         # (default, but explicit)
  #layout
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none",
        axis.title.x = element_blank()) +
  ggtitle("mCherry, flipping clones E2252") +
  ylab("relative Sox2::mCherry expression") + 
  col_scale_MK_cl


plot_grid(pClones_mT_LP, pClones_mCh_LP, ncol = 1, align = "hv", axis = "lr")
```

## Plot FACS density
```{r}
# Filter for landing_pad_location LP-116kb
LP116kb = expression_tib_comb %>% 
  filter(landing_pad_location == "LP -116 kb") %>% 
  pull(name)

ggplot(flowset_fluo_E2252[LP116kb,], aes(x = mTurq, col = insert_name, linetype = name)) + #, col = as_factor(sort2))) + 
  facet_grid(cell_line ~ .) +
  geom_density() +
   ggcyto::scale_x_flowjo_biexp(widthBasis = -100, breaks = c(-10^3, 0, 10^3, 10^4, 10^5),
                               pos = 4.42) +
  theme_bw()
  
ggplot(flowset_fluo_E2252[LP116kb,], aes(x = mTurq, col = insert_name, linetype = clone)) + #, col = as_factor(sort2))) + 
  #facet_grid(cell_line ~ .) +
  geom_density() +
   ggcyto::scale_x_flowjo_biexp(widthBasis = -100, breaks = c(-10^3, 0, 10^3, 10^4, 10^5),
                               pos = 4.42) +
  theme_bw()


# Filter for landing_pad_location LP-39 kb
LP39kb = expression_tib_comb %>% 
  filter(landing_pad_location == "LP -39 kb") %>% 
  pull(name)

ggplot(flowset_fluo_E2252[LP39kb,], aes(x = mTurq, col = insert_name, linetype = name)) + #, col = as_factor(sort2))) + 
  facet_grid(cell_line ~ .) +
  geom_density() +
   ggcyto::scale_x_flowjo_biexp(widthBasis = -100, breaks = c(-10^3, 0, 10^3, 10^4, 10^5),
                               pos = 4.42) +
  theme_bw()


# Plot both LPs & ctrl (F121)
LP_both_ctrl = expression_tib_comb %>% 
  filter(landing_pad_location %in% c("LP -39 kb", "LP -116 kb")) %>% 
  pull(name)


ggplot(flowset_fluo_E2252, aes(x = mTurq, col = cell_line, linetype = clone)) + #, col = as_factor(sort2))) + 
  facet_grid(landing_pad_location ~ .) +
  geom_density() +
   ggcyto::scale_x_flowjo_biexp(widthBasis = -100, breaks = c(-10^3, 0, 10^3, 10^4, 10^5),
                               pos = 4.42) +
  theme_bw() +
  col_scale_MK_cl
```


