---
title: "me20250709_E2296_FACS_timecourse"
author: "Mathias"
date: "2025-07-08"
output: 
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---
# Aim 
In a previous experiment we flipped MK constructs into LP8 and LP23 (E2224). During the first FACS measurements we saw some clonal variation between the same constructs, this can only be epigenetic since the clones have been genotyped. We had some indication that after we brought the same clone in culture again, it chaged expression. 
Therefore we were wondering if the reporter expression of clones changes over time. 
We got the following cells in culture
8.34.8 - 2 x vials frozen on 09/05 *these contained only 1 mill per vial so I used two)
8_MK1_C1 - vial frozen on 19/6 (cryobase) (new)
8_MK1_C2 - vial frozen on 19/6 (cryobase) (new)
8_MK2_C1 - vial frozen on 24/04 (old)
8_MK2_C2 - vial frozen on 24/04 (old)
8_MK2_C4 - vial frozen on 24/04 (old)
8_MK2_C5 - vial frozen on 03/04 (only one vial left now) (old)
8_MK2_C6 - vial frozen on 24/04 (old)
8_MK2_C5 - vial frozen on 19/6 (cryobase) (new)

Compare the measurements over time and see if they change

# Method
- load data into R (gate for single cells in FlowJo)
- curate dataset, annotate
- merge with FACS metadata
- substract autofluorescence from each WT measurement day
- normalize to GFP and to 8.38.8 control measurements per day

# Set-up
##Libraries
Load the libraries and set the parameters.
```{r}

# Load dependencies
library(GenomicRanges)
library(rtracklayer)
library(tidyverse)
# library(dtplyr)
library(dplyr)
# library(tidyr)
library(ggbeeswarm)
library(ggplot2)
library(Biostrings)
library(stringr)
library(readr)
library(knitr)
library(RColorBrewer)
library(RcppRoll)
# library(plotly)
library(readxl)
library(ggpubr)
library(ggpmisc)
library(ggridges)
library(import)
library(ggh4x)
library(rstatix)
# library(flowCore)
import::from(flowCore, .except=c("filter")) #I don't load the whole flowCore and ggcyto libraries, because they overwrite dplyr's filter function :0
import::from(ggcyto, 'fortify_fs')
library(flowDensity)
library(cowplot)
library(GENOVA)
library(Signac)

split_string <- function(vect,sep,N1,N2=N1){
  library(stringr)
  sapply(vect, function(X){
    paste(str_split(X,sep)[[1]][N1:N2],collapse = sep)
  },USE.NAMES = F)
}
```


```{r paths_parameters}
#datatag
datetag = paste0("ME",format(Sys.time(), '%Y%m%d'))

# Prepare output 
output_dir <- paste0("/DATA/projects/CTCF_Sox2/Scripts/for_figures/Figure_3/analysis_FACS_timecourse_", datetag)
dir.create(output_dir, showWarnings = FALSE)

library(knitr)
opts_chunk$set(
               cache = T,
               message = F, 
               warning = F,
               dev=c('png', 'pdf'), 
               dpi = 600,
               fig.path = file.path(output_dir, "figures/")) 
pdf.options(useDingbats = FALSE)
```


## Load fcs files
```{r}
path_fcs_files_E2296_T1 = "/DATA/projects/CTCF_Sox2/FACS_files/export_fcs_R_E2296/T1"
path_fcs_files_E2296_T3 = "/DATA/projects/CTCF_Sox2/FACS_files/export_fcs_R_E2296/T3/"
path_fcs_files_E2296_T5 = "/DATA/projects/CTCF_Sox2/FACS_files/export_fcs_R_E2296/T5/"
path_fcs_files_E2296_T7 = "/DATA/projects/CTCF_Sox2/FACS_files/export_fcs_R_E2296/T7/"
path_fcs_files_E2296_T9 = "/DATA/projects/CTCF_Sox2/FACS_files/export_fcs_R_E2296/T9/"
path_fcs_files_E2296_T11 = "/DATA/projects/CTCF_Sox2/FACS_files/export_fcs_R_E2296/T11/"

path_fcs_files_E2296 = "/DATA/projects/CTCF_Sox2/FACS_files/export_fcs_R_E2296/all_combined/"

fcs_files_E2296 = dir(path_fcs_files_E2296, full.names = T)
fcs_files_E2296_nms = str_remove(fcs_files_E2296, "^.*/")

fcs_files_E2296_nms2 = fcs_files_E2296_nms
```
#chat gpt
```{r}

# your raw filenames
nms <- fcs_files_E2296_nms

# 1) pull out the three-digit index before “_Single Cells.fcs”
index <- str_extract(nms, "(?<=_)(\\d{3})(?=_Single Cells\\.fcs)")

# 2) pull out any “t=N” timepoint, default to 1 if missing
timepoint <- str_extract(nms, "(?<=t=)\\d+")
timepoint[is.na(timepoint)] <- "1"

# 3) strip off everything up to the clone name
clone_raw <- str_replace(nms, "^.*#8_flip_", "")
clone_raw <- str_replace(clone_raw, "^.*timeline_[^_]+_", "")

# 4) remove the “ t=N” bit
clone_raw <- str_replace(clone_raw, " t=\\d+", "")

# 5) remove the “_<index>_Single Cells.fcs” suffix
clone <- str_replace(clone_raw,
                     paste0("_", index, "_Single Cells\\.fcs$"),
                     "")

# 6) assemble “clone_index” as your sample IDs
sample_id <- paste0(clone, "_", index)

# 7) ensure uniqueness (in case of exact duplicates)
sample_id <- make.unique(sample_id)

# view
sample_id

```


```{r}
# assign our cleaned sample IDs to the file vector and read into a flowSet
names(fcs_files_E2296) <- sample_id
flowset_E2296 <- read.flowSet(files = fcs_files_E2296,
                              alter.names = TRUE,
                              truncate_max_range = FALSE)
sampleNames(flowset_E2296) <- sample_id
pData(flowset_E2296)$name <- sample_id

# subset only the fluorescence + scatter channels and rename
flowset_fluo_E2296 <- flowset_E2296[, c(
  "BL.B..530_30.A",  # GFP
  "YG.D..610_20.A",  # mCherry
  "V.F..450_50.A",   # mTurq
  "FSC.A", "SSC.A"
)]
colnames(flowset_fluo_E2296) <- c("GFP", "mCherry", "mTurq", "FSC", "SSC")

# build annotation tibble
tib_annot_E2296 <- tibble(
  sample_name = sample_id,
  clone       = clone,
  timepoint   = factor(timepoint, levels = c("1", "3", "5", "7", "9", "11"),
                       labels = c("T1", "T3", "T5", "T7", "T9", "T11"))) %>%
  mutate(cryo = case_when(grepl("CTRL", sample_name) ~ "cryo",
                                grepl("old", sample_name) ~ "old",
                                grepl("new_cryo", sample_name) ~ "new_cryo",
                                .default = "other"),
         population = case_when(grepl("pMK1", sample_name) ~ "MK01",
                                grepl("pMK2", sample_name) ~ "MK02",
                                grepl("pMK3", sample_name) ~ "MK03",
                                grepl("WT", sample_name) ~ "ctrl",
                                grepl("8_34_8", sample_name) ~ "ctrl"),
         cell_line = case_when(str_detect(clone, "^8_34_8")      ~ "8_34_8",
                               str_detect(clone, "#8_pMK1_C1")     ~ "8_MK01_C1",
                               str_detect(clone, "#8_pMK1_C2")     ~ "8_MK01_C2",
                               str_detect(clone, "#8_pMK2_C1")     ~ "8_MK02_C1",
                               str_detect(clone, "#8_pMK2_C2")     ~ "8_MK02_C2",
                               str_detect(clone, "#8_pMK2_C4")     ~ "8_MK02_C4",
                               str_detect(clone, "#8_pMK2_C5")     ~ "8_MK02_C5",
                               str_detect(clone, "#8_pMK2_C6")     ~ "8_MK02_C6",
                               str_detect(clone, "^WT")          ~ "WT",
                               TRUE ~ NA_character_),
         cell_line_TP = paste0(cell_line, "_", timepoint),
         experiment = "E2296")


# merge annotation into the flowSet phenotype data
pheno_data_tib_E2296 = left_join(pData(flowset_fluo_E2296), tib_annot_E2296, join_by(name == sample_name)) 
rownames(pheno_data_tib_E2296) = pheno_data_tib_E2296$name
pData(flowset_fluo_E2296) = pheno_data_tib_E2296
```

## calculate expression
For now I calculate the median expression without subtracting any autofluorescence
```{r median_E2296}

total_cell_count_tmp = flowCore::keyword(flowset_fluo_E2296, "$TOT")[,'$TOT']
total_cell_count = as.numeric(total_cell_count_tmp)
names(total_cell_count) = names(total_cell_count_tmp)

median_mTurq = sapply(1:length(flowset_fluo_E2296), function(i){
  median(flowCore::exprs(flowset_fluo_E2296[[i]])[,'mTurq'])
})
names(median_mTurq) = sampleNames(flowset_fluo_E2296)

median_mCherry = sapply(1:length(flowset_fluo_E2296), function(i){
  median(flowCore::exprs(flowset_fluo_E2296[[i]])[,'mCherry'])
})
names(median_mCherry) = sampleNames(flowset_fluo_E2296)


median_GFP = sapply(1:length(flowset_fluo_E2296), function(i){
  median(flowCore::exprs(flowset_fluo_E2296[[i]])[,'GFP'])
})
names(median_GFP) = sampleNames(flowset_fluo_E2296)

flowCore::pData(flowset_fluo_E2296)$median_mTurq = median_mTurq
flowCore::pData(flowset_fluo_E2296)$median_mCherry = median_mCherry
flowCore::pData(flowset_fluo_E2296)$median_GFP = median_GFP

flowCore::pData(flowset_fluo_E2296)$total_cell_count = total_cell_count

fcs_tib = tibble(flowCore::pData(flowset_fluo_E2296))
```

# Normalize expression
```{r normalize_expression_values}
#subtract autofluorescence, normalize to GFP
autofluo_tib_comb = fcs_tib  %>%
  group_by(cell_line, timepoint) %>%
  filter(cell_line == "WT") %>%
  summarize(mTurq_WT = mean(median_mTurq),
           mCherry_WT = mean(median_mCherry),
            GFP_WT = mean(median_GFP)) %>% 
  ungroup() %>% 
  select(-(cell_line))

expression_tib_comb = fcs_tib %>%
  left_join(autofluo_tib_comb, join_by(timepoint)) %>%
  mutate(median_mTurq_cor = median_mTurq - mTurq_WT,
         median_mCherry_cor = median_mCherry - mCherry_WT,
         median_GFP_cor = median_GFP - GFP_WT) %>%
  mutate(mTurq_norm = median_mTurq_cor / median_GFP_cor,
         mCherry_norm = median_mCherry_cor / median_GFP_cor)

#normalize to 8_34_8_ctrl per timepoint (another alternative could be to normalize over the median of the 8_34_8 measurements)
standard_tib = expression_tib_comb %>%
  group_by(cell_line, timepoint) %>%
  filter(cell_line == "8_34_8") %>%
  summarize(mTurq_norm_standard = mean(mTurq_norm),
           mCherry_norm_standard = mean(mCherry_norm),
            GFP_standard = mean(median_GFP_cor)) %>%
  ungroup() %>% 
  select(-(cell_line))

expression_tib_comb = expression_tib_comb %>%
  left_join(standard_tib, join_by(timepoint)) %>%
  mutate(rel_mTurq = mTurq_norm / mTurq_norm_standard,
         rel_mCherry = mCherry_norm / mCherry_norm_standard,
         rel_GFP_cor = median_GFP_cor / GFP_standard)
``` 

## Color setup
```{r}

myCols_insert_population <- c(
  `MK01` = '#1b9e77',   # Teal green
  `MK02` = '#d95f02',   # Strong orange
  `MK03` = '#7570b3',   # Purple
  `WT` = 'black', # Black for wild type
  "ctrl" = "black" # Black for ctrl
)

col_scale_MK_pop <- scale_colour_manual(
  name = "cell_line", 
  values = myCols_insert_population,
  aesthetics = c("colour", "fill")
)

## SCR
brown = "#ffb000"
```


## Plot clones
```{r expression of clones in timecourse}

pTimecourse_mTurq = 
filter(expression_tib_comb, total_cell_count >= 500 & cell_line != "WT") %>%
  #plotting
  ggplot(aes(x = timepoint, y = rel_mTurq, col = timepoint)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(position = position_jitterdodge(jitter.width = 0.8), size = 2) +
  geom_hline(yintercept = 1, linetype = "dashed") + 
  geom_hline(yintercept = 0) + 
  facet_nested(. ~ cell_line  , scales = "free_x", space = "free_x") +
  #layout
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        axis.title.x = element_blank()) +
  ggtitle("mTurq, flipping clones E2296") +
  ylab("relative reporter expression")
  

# Relative mCherry
pTimecourse_mCherry = 
filter(expression_tib_comb, total_cell_count >= 500 & cell_line != "WT") %>%
  #plotting
  ggplot(aes(x = timepoint, y = rel_mCherry, col = timepoint)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(position = position_jitterdodge(jitter.width = 0.8), size = 2) +
  geom_hline(yintercept = 1, linetype = "dashed") + 
  geom_hline(yintercept = 0) + 
  facet_nested(. ~ cell_line  , scales = "free_x", space = "free_x") +
  #layout
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        axis.title.x = element_blank()) +
  ggtitle("mCherry, flipping clones E2296") +
  ylab("relative Sox2::mCherry expression")

plot_grid(pTimecourse_mTurq, pTimecourse_mCherry, ncol = 1)
```
**Conclusion** There is clearly something going on in T3, so discard T3 WT sample. I checked and the mCherry of WT is lower than usual, so if i normlize the MKs to that, everything will be higher. 

## Filter out T3
```{r expression of clones in timecourse without T3}
pTimecourse_mTurq_noT3 = 
filter(expression_tib_comb, total_cell_count >= 500 & cell_line != "WT" & timepoint != "T3") %>%
  #plotting
  ggplot(aes(x = timepoint, y = rel_mTurq, col = population)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(position = position_jitterdodge(jitter.width = 0.8), size = 1) +
  geom_hline(yintercept = 1, linetype = "dashed") + 
  geom_hline(yintercept = 0) + 
  facet_nested(. ~ cell_line  , scales = "free_x", space = "free_x") +
  #layout
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        axis.title.x = element_blank()) +
  ggtitle("mTurq, flipping clones E2296") +
  ylab("relative reporter expression") +
  col_scale_MK_pop

# Relative mCherry
pTimecourse_mCherry_noT3 = 
filter(expression_tib_comb, total_cell_count >= 500 & cell_line != "WT" & timepoint != "T3") %>%
  #plotting
  ggplot(aes(x = timepoint, y = rel_mCherry, col = population)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(position = position_jitterdodge(jitter.width = 0.8), size = 1) +
  geom_hline(yintercept = 1, linetype = "dashed") + 
  geom_hline(yintercept = 0) + 
  facet_nested(. ~ cell_line  , scales = "free_x", space = "free_x") +
  #layout
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        axis.title.x = element_blank()) +
  ggtitle("mCherry, Timecourse E2296") +
  ylab("relative Sox2::mCherry expression") +
  col_scale_MK_pop

plot_grid(pTimecourse_mTurq_noT3, pTimecourse_mCherry_noT3, ncol = 1, align = "hv")

#facet by timepoints
filter(expression_tib_comb, total_cell_count >= 500 & cell_line != "WT" & timepoint != "T3") %>%
  #plotting
  ggplot(aes(x = cell_line, y = rel_mTurq, col = cell_line)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(position = position_jitterdodge(jitter.width = 0.8), size = 2) +
  geom_hline(yintercept = 1, linetype = "dashed") + 
  geom_hline(yintercept = 0) + 
  facet_nested(. ~ timepoint  , scales = "free_x", space = "free_x") +
  #layout
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none",
        axis.title.x = element_blank()) +
  ggtitle("mTurq, Timecourse E2296") +
  ylab("relative mTurq expression")


filter(expression_tib_comb, total_cell_count >= 500 & cell_line != "WT" & timepoint != "T3") %>%
  #plotting
  ggplot(aes(x = cell_line, y = rel_mCherry, col = cell_line)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(position = position_jitterdodge(jitter.width = 0.8), size = 2) +
  geom_hline(yintercept = 1, linetype = "dashed") + 
  geom_hline(yintercept = 0) + 
  facet_nested(. ~ timepoint  , scales = "free_x", space = "free_x") +
  #layout
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none",
        axis.title.x = element_blank()) +
  ggtitle("mCherry, Timecourse E2296") +
  ylab("relative Sox2::mCherry expression")


#Facet for insert
filter(expression_tib_comb, total_cell_count >= 500 & cell_line != "WT" & timepoint != "T3") %>%
  #plotting
  ggplot(aes(x = population, y = rel_mTurq, col = population)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(position = position_jitterdodge(jitter.width = 0.8), size = 2) +
  geom_hline(yintercept = 1, linetype = "dashed") + 
  geom_hline(yintercept = 0) + 
  facet_nested(. ~ timepoint  , scales = "free_x", space = "free_x") +
  #layout
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none",
        axis.title.x = element_blank()) +
  ggtitle("mTurq, Timecourse E2296") +
  ylab("relative mTurq expression")

filter(expression_tib_comb, total_cell_count >= 500 & cell_line != "WT" & timepoint != "T3") %>%
  #plotting
  ggplot(aes(x = population, y = rel_mCherry, col = population)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(position = position_jitterdodge(jitter.width = 0.8), size = 2) +
  geom_hline(yintercept = 1, linetype = "dashed") + 
  geom_hline(yintercept = 0) + 
  facet_nested(. ~ timepoint  , scales = "free_x", space = "free_x") +
  #layout
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none",
        axis.title.x = element_blank()) +
  ggtitle("mCherry, Timecourse E2296") +
  ylab("relative Sox2::mCherry expression")

```

## Paper plot
```{r}
pTimecourse_mTurq_noT3_paper =
expression_tib_comb %>% 
  filter(total_cell_count >= 500 & cell_line != "WT" & timepoint != "T3") %>%
  group_by(cell_line, timepoint, population) %>% 
  summarise(mean_rel_mTurq = mean(rel_mTurq),
            mean_rel_mCh = mean (rel_mCherry)) %>%
  #plotting
  ggplot(aes(x = timepoint, y = mean_rel_mTurq, col = population)) +
  geom_point(position = position_jitterdodge(jitter.width = 0.2), size = 2) +
  geom_hline(yintercept = 1, linetype = "dashed") + 
  geom_hline(yintercept = 0) + 
  facet_nested(. ~ population+cell_line , scales = "free_x", space = "free_x") +
  #layout
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        axis.title.x = element_blank()) +
  ggtitle("Timecourse of reporter expresion in LP-39kb", subtitle = "rel. Sox2P reporter expression") +
  ylab("relative reporter expression") +
  col_scale_MK_pop



pTimecourse_mCh_noT3_paper =
expression_tib_comb %>% 
  filter(total_cell_count >= 500 & cell_line != "WT" & timepoint != "T3") %>%
  group_by(cell_line, timepoint, population) %>% 
  summarise(mean_rel_mTurq = mean(rel_mTurq),
            mean_rel_mCh = mean (rel_mCherry)) %>%
  #plotting
  ggplot(aes(x = timepoint, y = mean_rel_mCh, col = population)) +
  geom_point(position = position_jitterdodge(jitter.width = 0.2), size = 2) +
  geom_hline(yintercept = 1, linetype = "dashed") + 
  geom_hline(yintercept = 0) + 
  facet_nested(. ~ population+cell_line , scales = "free_x", space = "free_x") +
  #layout
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        axis.title.x = element_blank()) +
  ggtitle("Timecourse of reporter expresion in LP-39kb", subtitle = "rel. Sox2::mCherry expression") +
  ylab("relative Sox2::mCherry expression") +
  col_scale_MK_pop

plot_grid(pTimecourse_mTurq_noT3_paper, pTimecourse_mCh_noT3_paper, ncol = 1, align = "hv")
```

## display as violin plot
```{r violin plots of clones timecourse}

without_WT_T3 = fcs_tib %>% filter(cell_line != "WT" & timepoint != "T3") %>% pull(name)
flowset_oi = flowset_fluo_E2296[without_WT_T3]

ggplot(flowset_oi, aes(x = cell_line, y = mTurq, fill = population)) +
  geom_violin(trim = TRUE, alpha = 0.6) +
  geom_boxplot(width = 0.1, outlier.shape = NA) +
  facet_wrap(~ timepoint, scales = "free_y") +
  scale_y_log10() +
  theme_classic() +
  labs(
    x = "Clone",
    y = "mTurq (log10)",
    title = "Distributions of mTurq by clone at each timepoint"
  ) +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
    legend.position = "none"
  ) +
  col_scale_MK_pop
```

