---
title: "me20251125_Comp_clonal_vs_ExpScore_Figure_S3G"
author: "Mathias"
date: "2025-11-25"
output: 
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---

# To-do
- load stranded expression scores
- load clonal data

# Setup 
## Libraries
```{r}
library(GenomicRanges)
library(rtracklayer)
library(tidyverse)
library(dplyr)
library(ggbeeswarm)
library(ggplot2)
library(Biostrings)
library(stringr)
library(readr)
library(knitr)
library(RColorBrewer)
library(RcppRoll)
# library(plotly)
library(readxl)
library(ggpubr)
library(ggpmisc)
library(ggridges)
library(import)
library(ggh4x)
# library(flowCore)
import::from(flowCore, .except=c("filter")) #I don't load the whole flowCore and ggcyto libraries, because they overwrite dplyr's filter function :0
import::from(ggcyto, 'fortify_fs')
library(flowDensity)
library(cowplot)


split_string <- function(vect,sep,N1,N2=N1){
  library(stringr)
  sapply(vect, function(X){
    paste(str_split(X,sep)[[1]][N1:N2],collapse = sep)
  },USE.NAMES = F)
}
```

## Paths & parameters
```{r paths_parameters}
#datatag
datetag = paste0("ME",format(Sys.time(), '%Y%m%d'))

# Prepare output 
output_dir <- paste0("analysis_",datetag)
dir.create(output_dir, showWarnings = FALSE)

```

```{r knits setup}

library(knitr)
opts_chunk$set(fig.width = 20, fig.height = 8, cache = T,
               message = F, warning = F,
               dev=c('png', 'pdf'), 
               fig.path = file.path(output_dir, "figures/")) 
pdf.options(useDingbats = FALSE)
```


## Color scheme
```{r}

myCols_insert_cell_lines <- c(
  `MK01` = '#1b9e77',   # Teal green
  `MK02` = '#d95f02',   # Strong orange
  `MK03` = '#7570b3',   # Purple
  `23_34A` = 'black', # 
  `8_34_8` = 'black', # 
  `WT` = 'black'        # Black for wild type
)

col_scale_MK_cl <- scale_colour_manual(
  name = "cell_line", 
  values = myCols_insert_cell_lines,
  aesthetics = c("colour", "fill")
)


## SCR
brown = "#ffb000"
```

## Load functions
```{r load functions}
source("/DATA/projects/Sox2/General_functions/CM20240621_general_plotting_functions.R")
source("/DATA/projects/Sox2/General_functions/CM20250318_new_expression_score_functions.R")
source("/DATA/projects/CTCF_Sox2/functions/ME20250707_mCh_new_expression_score_functions.R")

```

## Relevant annotations
```{r element_coordinates}

# Location of enhancer / gene
enhancer <- GRanges(seqnames = "chr3", 
                    ranges = IRanges(start = 34753415,
                                     end = 34766401),
                    strand = "*")

# SCR sequence used by Luca Giorgetti study
enhancer_giorgetti <- GRanges(seqnames = "chr3", 
                    ranges = IRanges(start = 34757649,
                                     end = 34762535),
                    strand = "*")

#gene based on mm10, ucsc browser, RefSeq annotation
# Sox2
# chr3:34649995-34652461 (+)
# id = NM_011443.4

Sox2_gene <- GRanges(seqnames = "chr3", 
                    ranges = IRanges(start = 34649995,
                                     end = 34652461),
                    strand = "+")

#the old gene annotation, did not include the 5' UTR (so from start codon to end of 3' UTR)
gene_old <- GRanges(seqnames = "chr3", 
                    ranges = IRanges(start = 34650405,
                                     end = 34652461),
                    strand = "+")


landingPad_23 <- GRanges(seqnames = "chr3", 
                    ranges = IRanges(start = 34643960,
                                     end = 34643962),
                    strand = "+")

landingPad_8 <- GRanges(seqnames = "chr3", 
                    ranges = IRanges(start = 34721183,
                                     end = 34721184),
                    strand = "+")


LP_tib = tibble(landing_pad = c("LP23", "LP8"), 
                start = c( start(landingPad_23),start(landingPad_8)))

#CTCF sites based on our own mapping
# new analysis (based on mapping to mm10 only)
CTCF_mm10_new <- import.bed('/DATA/usr/v.franceschini/Workspaces/2024_01_MATHIAS_CTCF_PEAKS/VF240812_calling_CTCF_motifs_again/02_OUTPUTS/CTCF_sites/6764_2_ME_E2_CGATGT_S2_R1_001_peaks_motifs.bed')

#filtering:
CTCF_mm10.chr3 <- CTCF_mm10_new[seqnames(CTCF_mm10_new)== 'chr3']
# add the missing site after the SCR
CTCF_mm10.chr3_extra = sort(c(CTCF_mm10.chr3,
                         GRanges(seqnames = "chr3",
                                 IRanges(start = 34772210, end = 34772210),
                                 strand = "+")),
                         ignore.strand = T)


prange_zoom = c(34500000 , 35000000)
prange2 = c(34000000, 35500000)
prange_far = c(32000000, 37000000)
#zoom in on only the SCR +- 50kb
prange_LP = c(start(landingPad_23)-5E4, end(landingPad_23)+5E4)
prange_SCR = c(start(enhancer)-5E4, end(enhancer)+5E4)
prange_plot = c(32643960, 36643960) #+/- 2MB from landing pad
prange_loop = c(34631804, 34778268)
prange_zoom = c(34.5E6, 35E6)
prange_zoom_paper = c(34.59E6, 34.83E6)
prange_SCR = c(34.725E6, 34.785E6)
```

# Load stranded expression score 
Stranded expression score was calculated with a binzise of 2000bp and 10 bins per window = 20kb with 1000 boostraps and a minimum of 3 integrations per window. 

## Cell line 23_34A
```{r}
# Read the integration files
tib_filtered_FACS_vals_comb = readRDS("/DATA/projects/CTCF_Sox2/analysis/RDS_files/tib_filtered_FACS_vals_comb.rds")

#all_exp
all_exp = unique(tib_filtered_FACS_vals_comb$experiment)
set.seed(20240331)
min_ints_window = 3
N_bootstrap = 1000
BinSize = 2000

expr_score_tib_23_34A_str = readRDS("/DATA/projects/CTCF_Sox2/analysis/expression_score_new/ME20250708_new_expression_score_23_34A_str.rds")

expr_score_sum_tib_23_34A_str = bootstrap_summary_fun_str(expr_score_tib_23_34A_str,
                                           score_column = "expr_score_window",
                                           conf_int = 0.95)

#calculate actual expr score (for filtering # of ints per window)
expr_score_real_23_34A_str =
  expr_score_function_new_weigh_stranded(tib_filtered_FACS_vals_comb,
                                binsize = BinSize,
                                N_bins_window = 10,
                                calculation_window =  prange_zoom,
                                CL = "23_34A",
                                EXP = all_exp)
  
expr_score_sum_tib_filt_comb_23_34A_str = left_join(expr_score_sum_tib_23_34A_str,  expr_score_real_23_34A_str) %>%
  mutate(expr_score_window_median_sel = case_when(
    total_ints_window >= min_ints_window ~ expr_score_window_median,
    .default = NA),
    expr_score_window_lower_sel = case_when(
      total_ints_window >= min_ints_window ~ expr_score_window_lower,
      .default = NA),
    expr_score_window_upper_sel = case_when(
      total_ints_window >= min_ints_window ~ expr_score_window_upper,
      .default = NA))

```
## Cell line MK01
```{r}
expr_score_tib_23_MK01_C5_str = readRDS("/DATA/projects/CTCF_Sox2/analysis/expression_score_new/ME20250708_new_expression_score_23_MK01_C5_str.rds")

expr_score_sum_tib_23_MK01_C5 = bootstrap_summary_fun_str(expr_score_tib_23_MK01_C5_str,
                                           score_column = "expr_score_window",
                                           conf_int = 0.95)

#calculate actual expr score (for filtering # of ints per window)
expr_score_real_23_MK01_C5 =
  expr_score_function_new_weigh_stranded(tib_filtered_FACS_vals_comb,
                                binsize = BinSize,
                                N_bins_window = 10,
                                calculation_window =  prange_zoom,
                                CL = "23_MK01_C5",
                                EXP = all_exp)
  
expr_score_sum_tib_filt_comb_23_MK01_C5_str = left_join(expr_score_sum_tib_23_MK01_C5,  expr_score_real_23_MK01_C5) %>%
  mutate(expr_score_window_median_sel = case_when(
    total_ints_window >= min_ints_window ~ expr_score_window_median,
    .default = NA),
    expr_score_window_lower_sel = case_when(
      total_ints_window >= min_ints_window ~ expr_score_window_lower,
      .default = NA),
    expr_score_window_upper_sel = case_when(
      total_ints_window >= min_ints_window ~ expr_score_window_upper,
      .default = NA))
```
## Cell line MK02
```{r}
expr_score_tib_23_MK02_C2_str = readRDS("/DATA/projects/CTCF_Sox2/analysis/expression_score_new/ME20250708_new_expression_score_23_MK02_C2_str.rds")



expr_score_sum_tib_23_MK02_C2 = bootstrap_summary_fun_str(expr_score_tib_23_MK02_C2_str,
                                           score_column = "expr_score_window",
                                           conf_int = 0.95)

#calculate actual expr score (for filtering # of ints per window)
expr_score_real_23_MK02_C2_str =
  expr_score_function_new_weigh_stranded(tib_filtered_FACS_vals_comb,
                                binsize = BinSize,
                                N_bins_window = 10,
                                calculation_window =  prange_zoom,
                                CL = "23_MK02_C2",
                                EXP = all_exp)
  
expr_score_sum_tib_filt_comb_23_MK02_C2_str = left_join(expr_score_sum_tib_23_MK02_C2,  expr_score_real_23_MK02_C2_str) %>%
  mutate(expr_score_window_median_sel = case_when(
    total_ints_window >= min_ints_window ~ expr_score_window_median,
    .default = NA),
    expr_score_window_lower_sel = case_when(
      total_ints_window >= min_ints_window ~ expr_score_window_lower,
      .default = NA),
    expr_score_window_upper_sel = case_when(
      total_ints_window >= min_ints_window ~ expr_score_window_upper,
      .default = NA))
```
##Cell line MK03
```{r}

expr_score_tib_23_MK03_C1 = readRDS("/DATA/projects/CTCF_Sox2/analysis/expression_score_new/ME20250708_new_expression_score_23_MK03_C1_stranded.rds")

expr_score_sum_tib_23_MK03_C1 = bootstrap_summary_fun_str(expr_score_tib_23_MK03_C1,
                                           score_column = "expr_score_window",
                                           conf_int = 0.95)

#calculate actual expr score (for filtering # of ints per window)
expr_score_real_23_MK03_C1_str =
  expr_score_function_new_weigh_stranded(tib_filtered_FACS_vals_comb,
                                binsize = BinSize,
                                N_bins_window = 10,
                                calculation_window =  prange_zoom,
                                CL = "23_MK03_C1",
                                EXP = all_exp)
  
expr_score_sum_tib_filt_comb_23_MK03_C1_str = left_join(expr_score_sum_tib_23_MK03_C1,  expr_score_real_23_MK03_C1_str) %>%
  mutate(expr_score_window_median_sel = case_when(
    total_ints_window >= min_ints_window ~ expr_score_window_median,
    .default = NA),
    expr_score_window_lower_sel = case_when(
      total_ints_window >= min_ints_window ~ expr_score_window_lower,
      .default = NA),
    expr_score_window_upper_sel = case_when(
      total_ints_window >= min_ints_window ~ expr_score_window_upper,
      .default = NA))
```


# Load clonal data
## Load FCS files
```{r}
path_fcs_files_E2252 = "/DATA/projects/CTCF_Sox2/FACS_files/export_fcs_R_E2252"

fcs_files_E2252 = dir(path_fcs_files_E2252, full.names = T)
fcs_files_E2252_nms = str_remove(fcs_files_E2252, "^.*/")

fcs_files_E2252_nms2 = split_string(fcs_files_E2252_nms, "_", 6, 8)
names(fcs_files_E2252) = fcs_files_E2252_nms2

flowset_E2252 = read.flowSet(files = fcs_files_E2252, alter.names = T, truncate_max_range = F)
sampleNames(flowset_E2252) = names(fcs_files_E2252)
pData(flowset_E2252)$name = names(fcs_files_E2252)


flowset_fluo_E2252 = flowset_E2252[,c('BL.B..530_30.A', 'YG.D..610_20.A', 'V.F..450_50.A', 'FSC.A', 'SSC.A')]
colnames(flowset_fluo_E2252) = c("GFP", "mCherry", "mTurq", 'FSC', 'SSC')
```

## Annotation
```{r}
tib_annot_E2252 = tibble(sample_name = names(fcs_files_E2252)) %>%
  mutate(cell_line = case_when(grepl("_MK1", sample_name) ~ "MK01",
                               grepl("_MK2", sample_name) ~ "MK02",
                               grepl("_MK3", sample_name) ~ "MK03",
                               grepl("23_34A", sample_name) ~ "23_34A",
                               grepl("8_34_8", sample_name) ~ "8_34_8",
                               grepl("F121_9_002", sample_name) ~ "F121",
                               grepl("WT", sample_name) ~ "WT"),
         landing_pad_location = case_when(grepl("23_", sample_name) ~ "LP -116 kb",
                                          grepl("8_", sample_name) ~ "LP -39 kb",
                                          .default = "none"),
         LP_CL = paste0(cell_line, "_", landing_pad_location),
         clone_tmp = split_string(sample_name, "_", 3),
         clone = case_when(clone_tmp == "003" ~ "34A",
                           clone_tmp == "002" ~ "ctrl",
                           clone_tmp == "Single Cells.fcs" ~ "ctrl",
                           .default = clone_tmp),
         insert_name = case_when(cell_line == "MK01" ~ "CBS3x_Fw_Sox2P",
                                 cell_line == "MK02" ~ "CBS3x_Rv_Sox2P",
                                 cell_line == "MK03" ~ "CBS1x_Fw_Sox2P",
                                 cell_line == "23_34A" ~ "Sox2P",
                                 cell_line == "8_34_8" ~ "Sox2P",
                                 cell_line == "F121" ~ "none",
                                 .default = cell_line),
          experiment = "E2252") %>%
  select(-clone_tmp)


pheno_data_tib_E2252 = left_join(pData(flowset_fluo_E2252), tib_annot_E2252, join_by(name == sample_name) ) 
rownames(pheno_data_tib_E2252) = pheno_data_tib_E2252$name

pData(flowset_fluo_E2252) = pheno_data_tib_E2252
```

## Calculate expression
```{r}
total_cell_count_tmp = flowCore::keyword(flowset_fluo_E2252, "$TOT")[,'$TOT']
total_cell_count = as.numeric(total_cell_count_tmp)
names(total_cell_count) = names(total_cell_count_tmp)

median_mTurq = sapply(1:length(flowset_fluo_E2252), function(i){
  median(flowCore::exprs(flowset_fluo_E2252[[i]])[,'mTurq'])
})
names(median_mTurq) = sampleNames(flowset_fluo_E2252)

median_mCherry = sapply(1:length(flowset_fluo_E2252), function(i){
  median(flowCore::exprs(flowset_fluo_E2252[[i]])[,'mCherry'])
})
names(median_mCherry) = sampleNames(flowset_fluo_E2252)


median_GFP = sapply(1:length(flowset_fluo_E2252), function(i){
  median(flowCore::exprs(flowset_fluo_E2252[[i]])[,'GFP'])
})
names(median_GFP) = sampleNames(flowset_fluo_E2252)

flowCore::pData(flowset_fluo_E2252)$median_mTurq = median_mTurq
flowCore::pData(flowset_fluo_E2252)$median_mCherry = median_mCherry
flowCore::pData(flowset_fluo_E2252)$median_GFP = median_GFP

flowCore::pData(flowset_fluo_E2252)$total_cell_count = total_cell_count

fcs_tib = tibble(flowCore::pData(flowset_fluo_E2252))
```

## Normalize expression
```{r}
#subtract autofluorescence, normalize to GFP
autofluo_tib_comb = fcs_tib  %>%
  filter(cell_line == "WT") %>%
  summarize(mTurq_WT = mean(median_mTurq),
           mCherry_WT = mean(median_mCherry),
            GFP_WT = mean(median_GFP) )

expression_tib_comb = fcs_tib %>%
  bind_cols(autofluo_tib_comb) %>%
  mutate(median_mTurq_cor = median_mTurq - mTurq_WT,
         median_mCherry_cor = median_mCherry - mCherry_WT,
         median_GFP_cor = median_GFP - GFP_WT) %>%
  mutate(mTurq_norm = median_mTurq_cor / median_GFP_cor,
         mCherry_norm = median_mCherry_cor / median_GFP_cor)

#normalize to 23_34A (to compare between measurement days)
standard_tib = expression_tib_comb %>%
  filter(cell_line == "23_34A") %>%
  summarize(mTurq_norm_standard = mean(mTurq_norm),
           mCherry_norm_standard = mean(mCherry_norm),
            GFP_standard = mean(median_GFP_cor) )

expression_tib_comb = expression_tib_comb %>%
  bind_cols(standard_tib) %>%
  mutate(rel_mTurq = mTurq_norm / mTurq_norm_standard,
         rel_mCherry = mCherry_norm / mCherry_norm_standard,
         rel_GFP_cor = median_GFP_cor / GFP_standard)

stat_table = 
expression_tib_comb %>% 
  group_by(cell_line, landing_pad_location) %>% 
  summarise(mean_mCh = mean(rel_mCherry),
            mean_mTurq = mean(rel_mTurq))
```

## Plot clones
```{r}
# clones have been FACS measured and insert genotyped 

filter(expression_tib_comb, total_cell_count >= 500 & landing_pad_location != "none") %>%
  #plotting
  ggplot(aes(x = cell_line, y = rel_mTurq, col = cell_line)) +
  #geom_boxplot(outlier.shape = NA) +
  geom_jitter(size = 2, width = 0.2) +
  # add mean line
  stat_summary(fun = "mean", geom = "crossbar", 
               width = 0.5, color = "black", 
               aes(y = rel_mTurq), fatten = 2) +
  
  geom_hline(yintercept = 1, linetype = "dashed") + 
  geom_hline(yintercept = 0) + 
  facet_nested(. ~ landing_pad_location  , scales = "free_x", space = "free_x") +
  #layout
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none",
        axis.title.x = element_blank()) +
  ggtitle("mTurq, flipping clones E2252") +
  ylab("relative reporter expression") +
  col_scale_MK_cl
```

```{r}
# Relative mCherry
filter(expression_tib_comb, total_cell_count >= 500 & landing_pad_location != "none") %>%
  #plotting
  ggplot(aes(x = LP_CL, y = rel_mCherry, col = cell_line)) +
  geom_jitter(size = 2, width = 0.2) +
  # add mean line 
  stat_summary(fun = "mean", geom = "crossbar", 
               width = 0.5, color = "black",  
               aes(y = rel_mCherry), fatten = 2) +
  
  geom_hline(yintercept = 1, linetype = "dashed") + 
  geom_hline(yintercept = 0) + 
  facet_nested(. ~ landing_pad_location  , scales = "free_x", space = "free_x") +
  #layout
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none",
        axis.title.x = element_blank()) +
  ggtitle("mCherry, flipping clones E2252") +
  ylab("relative Sox2::mCherry expression") + 
  col_scale_MK_cl
```

# Compare Expression score with clonal data
```{r}
# filter expression score for bins overlapping LP39
lp_pos <- start(landingPad_8)[1]   # 34721183

bin_row <- expr_score_tib_23_34A_str %>%
  filter(cell_line == "23_34A", strand == "+") %>%      # optional filters
  mutate(bin_end = start_interval + 2 * (mid_interval - start_interval) - 1) %>%
  filter(start_interval <= lp_pos, bin_end >= lp_pos)

bin_row
```

## Comparison of calculated expression score with clonal data
Overlap is plus strand specific, matching the orientation of the inserts in the clones
```{r}
#combine all the data
expr_score_real_23_34A_str
expr_score_real_23_MK01_C5
expr_score_real_23_MK02_C2_str
expr_score_real_23_MK03_C1_str

tib_comb_str = bind_rows(expr_score_real_23_34A_str, expr_score_real_23_MK01_C5, expr_score_real_23_MK02_C2_str, expr_score_real_23_MK03_C1_str)

# 1) Add bin_end and drop rows with NA coordinates / strand
expr_with_end <- tib_comb_str %>%
  mutate(
    bin_end = start_interval + 2 * (mid_interval - start_interval) - 1
  ) %>%
  # keep only rows where the bin is defined
  tidyr::drop_na(start_interval, bin_end, strand) %>%
  ungroup()

# (Optional) check you cleaned it
colSums(is.na(expr_with_end[c("start_interval", "bin_end", "strand")]))

# 2) Turn bins into GRanges
bins_gr <- GRanges(
  seqnames = Rle("chr3"),  # or expr_with_end$seqnames if you have that column
  ranges   = IRanges(
    start = expr_with_end$start_interval,
    end   = expr_with_end$bin_end
  ),
  strand   = Rle(expr_with_end$strand)
)

all_lps <- c(landingPad_8, landingPad_23)
names(all_lps) <- c("landingPad_8", "landingPad_23")

hits <- findOverlaps(all_lps, bins_gr)

bins_for_lps <- expr_with_end[subjectHits(hits), ] %>%
  mutate(landing_pad = names(all_lps)[queryHits(hits)])

bins_for_lps


# plot real calculated exp score
bins_for_lps %>% 
  ggplot(aes(x = interval_number, y = expr_score, col = cell_line)) +
  geom_point()

bins_for_lps %>% 
  ggplot(aes(x = interval_number, y = expr_score_window, col = cell_line)) +
  geom_point()




bins_for_lps_ctrl = 
bins_for_lps %>% select(cell_line, landing_pad, expr_score_window) %>% 
  filter(cell_line == "23_34A") %>%
  group_by(landing_pad) %>%
  mutate(expr_score_ctrl = mean(expr_score_window)) %>%
  select(expr_score_ctrl)

tib_expr_score_real_enrich = 
bins_for_lps %>% 
  select(cell_line, landing_pad, expr_score_window) %>% 
  left_join(bins_for_lps_ctrl, join_by(landing_pad)) %>% 
  mutate(enrich = expr_score_window /expr_score_ctrl)

pExpr_score_enrich = 
  tib_expr_score_real_enrich %>% 
  ggplot(aes(x = landing_pad, y = enrich, col = cell_line)) +
  geom_point(size = 2) +
  geom_hline(yintercept = 1, col = "black", linetype = "dotted") +
  theme_bw()

pExpr_score_enrich

 
 
# plot together
#--> first normalize clones to each LP
#normalize to 23_34A per LP
expression_tib_comb = fcs_tib %>%
  bind_cols(autofluo_tib_comb) %>%
  mutate(median_mTurq_cor = median_mTurq - mTurq_WT,
         median_mCherry_cor = median_mCherry - mCherry_WT,
         median_GFP_cor = median_GFP - GFP_WT) %>%
  mutate(mTurq_norm = median_mTurq_cor / median_GFP_cor,
         mCherry_norm = median_mCherry_cor / median_GFP_cor)


standard_tib_LP = expression_tib_comb %>%
  filter(insert_name == "Sox2P") %>%
  group_by(landing_pad_location) %>%
  summarize(mTurq_norm_standard = mean(mTurq_norm),
           mCherry_norm_standard = mean(mCherry_norm),
            GFP_standard = mean(median_GFP_cor) )

expression_tib_comb_2 = expression_tib_comb %>%
  left_join(standard_tib_LP, join_by(landing_pad_location)) %>%
  mutate(rel_mTurq = mTurq_norm / mTurq_norm_standard,
         rel_mCherry = mCherry_norm / mCherry_norm_standard,
         rel_GFP_cor = median_GFP_cor / GFP_standard)


tib_clones = expression_tib_comb_2 %>% select(cell_line, insert_name, clone, landing_pad_location, rel_mTurq) %>%
  mutate(landing_pad = case_when(landing_pad_location == "LP -116 kb" ~ "landingPad_23",
                                 landing_pad_location == "LP -39 kb" ~ "landingPad_8"),
         exp_type = "clones") %>% select(landing_pad, insert_name, rel_mTurq, exp_type)
  

tib_exp_score = tib_expr_score_real_enrich %>% mutate(insert_name = case_when(cell_line == "23_34A" ~ "Sox2P",
                                                                         cell_line == "23_MK01_C5" ~ "CBS3x_Fw_Sox2P",
                                                                         cell_line == "23_MK02_C2" ~ "CBS3x_Rv_Sox2P",
                                                                         cell_line == "23_MK03_C1" ~ "CBS1x_Fw_Sox2P"),
                                                      rel_mTurq = enrich,
                                                      exp_type = "expr_score") %>% select(landing_pad, insert_name, rel_mTurq, exp_type)

# Plot combined data
tib_comb = rbind(tib_clones, tib_exp_score) %>% mutate(insert_name = factor(insert_name, levels = c("none", "WT", "Sox2P", "CBS1x_Fw_Sox2P", "CBS3x_Fw_Sox2P", "CBS3x_Rv_Sox2P")))



tib_comb %>% drop_na() %>%
  ggplot(aes(x = exp_type, y = rel_mTurq, col = exp_type)) +
  facet_nested(~ landing_pad + insert_name) +
  geom_point(width = 0.2, size = 2) +
  # add mean line
  stat_summary(data = tib_comb %>% filter(exp_type == "clones") %>% drop_na(),
    fun = "mean", geom = "crossbar", 
               width = 0.5, color = "black", 
               aes(y = rel_mTurq), fatten = 1) +
  geom_hline(yintercept = 1, col = "black", linetype = "dotted") + 
  theme_bw()


tib_comb %>% drop_na() %>%
  ggplot(aes(x = exp_type, y = rel_mTurq, col = exp_type)) +
  facet_nested(landing_pad ~ insert_name, scales = "free_y") +
  geom_point(width = 0.2, size = 2) +
  # add mean line
  stat_summary(data = tib_comb %>% filter(exp_type == "clones") %>% drop_na(),
    fun = "mean", geom = "crossbar", 
               width = 0.5, color = "black", 
               aes(y = rel_mTurq), fatten = 1) +
  geom_hline(yintercept = 1, col = "black", linetype = "dotted") + 
  theme_bw()


# Plot for paper with real enrichment
pLP23_real = 
tib_comb %>% drop_na() %>%
  filter(landing_pad == "landingPad_23") %>%
  ggplot(aes(x = exp_type, y = rel_mTurq, col = exp_type)) +
  facet_nested(~ landing_pad + insert_name) +
  geom_point(width = 0.2, size = 3, alpha = 0.6) +
  # add mean line
  stat_summary(data = tib_comb %>% filter(exp_type == "clones" & landing_pad == "landingPad_23") %>% drop_na(),
    fun = "mean", geom = "crossbar", 
               width = 0.5, color = "black", 
               aes(y = rel_mTurq), fatten = 1) +
  geom_hline(yintercept = 1, col = "black", linetype = "dotted") + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(legend.position = "none") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

pLP23_real

pLP8_real = 
tib_comb %>% drop_na() %>%
  filter(landing_pad == "landingPad_8") %>%
  ggplot(aes(x = exp_type, y = rel_mTurq, col = exp_type)) +
  facet_nested(~ landing_pad + insert_name) +
  geom_point(width = 0.2, size = 3, alpha = 0.6) +
  # add mean line
  stat_summary(data = tib_comb %>% filter(exp_type == "clones" & landing_pad == "landingPad_8") %>% drop_na(),
    fun = "mean", geom = "crossbar", 
               width = 0.5, color = "black", 
               aes(y = rel_mTurq), fatten = 1) +
  geom_hline(yintercept = 1, col = "black", linetype = "dotted") + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(legend.position = "bottom")

pLP8_real

plot_grid(pLP23_real, pLP8_real, ncol = 1, align = "v", rel_heights = c(0.55, 1))
```

## Comparison of bootstrapped expression score with clonal data
```{r}
# This are bootsrapped tibbles with upper and lower CI
expr_score_sum_tib_filt_comb_23_34A_str
expr_score_sum_tib_filt_comb_23_MK01_C5_str
expr_score_sum_tib_filt_comb_23_MK02_C2_str
expr_score_sum_tib_filt_comb_23_MK03_C1_str

tib_comb_str_boot = bind_rows(expr_score_sum_tib_filt_comb_23_34A_str, expr_score_sum_tib_filt_comb_23_MK01_C5_str, expr_score_sum_tib_filt_comb_23_MK02_C2_str, expr_score_sum_tib_filt_comb_23_MK03_C1_str)

# 1) Add bin_end and drop rows with NA coordinates / strand
expr_with_end <- tib_comb_str_boot %>%
  mutate(
    bin_end = start_interval + 2 * (mid_interval - start_interval) - 1
  ) %>%
  # keep only rows where the bin is defined
  tidyr::drop_na(start_interval, bin_end, strand)

# (Optional) check you cleaned it
colSums(is.na(expr_with_end[c("start_interval", "bin_end", "strand")]))

# 2) Turn bins into GRanges
bins_gr <- GRanges(
  seqnames = Rle("chr3"),  # or expr_with_end$seqnames if you have that column
  ranges   = IRanges(
    start = expr_with_end$start_interval,
    end   = expr_with_end$bin_end
  ),
  strand   = Rle(expr_with_end$strand)
)

all_lps <- c(landingPad_8, landingPad_23)
names(all_lps) <- c("landingPad_8", "landingPad_23")

hits <- findOverlaps(all_lps, bins_gr)

bins_for_lps <- expr_with_end[subjectHits(hits), ] 

bins_for_lps =
bins_for_lps %>%
  mutate(landing_pad = case_when(interval_number == "72" ~ "landingPad_23",
                                 interval_number == "111" ~ "landingPad_8"))

bins_for_lps

bins_for_lps %>% 
  ggplot(aes(x = interval_number, y = expr_score_window, col = cell_line)) +
  geom_point()

bins_for_lps_ctrl = 
bins_for_lps %>% select(cell_line, landing_pad, expr_score_window_median_sel) %>% 
  filter(cell_line == "23_34A") %>%
  group_by(landing_pad) %>%
  mutate(expr_score_ctrl = mean(expr_score_window_median_sel)) %>%
  select(expr_score_ctrl)

tib_bins_for_lps_boots =
  bins_for_lps %>% 
  select(cell_line, landing_pad, expr_score_window_median_sel) %>% 
  left_join(bins_for_lps_ctrl, join_by(landing_pad)) %>% 
  #bind_cols(bins_for_lps_ctrl) %>%
  mutate(rel_mTurq = expr_score_window_median_sel /expr_score_ctrl)


tib_bins_for_lps_boots %>% 
  ggplot(aes(x = landing_pad, y = rel_mTurq, col = cell_line)) +
  geom_point(size = 2) +
  geom_hline(yintercept = 1, col = "black", linetype = "dotted") +
  theme_bw()


#combine bootstrapped with other table
tib_bins_for_lps_boots = tib_bins_for_lps_boots %>% ungroup() %>% select(landing_pad, rel_mTurq, cell_line) %>% 
  mutate(insert_name = case_when(cell_line == "23_34A" ~ "Sox2P",
                          cell_line == "23_MK01_C5" ~ "CBS3x_Fw_Sox2P",
                          cell_line == "23_MK02_C2" ~ "CBS3x_Rv_Sox2P",
                          cell_line == "23_MK03_C1" ~ "CBS1x_Fw_Sox2P"),
         exp_type = "expr_score_boots") %>% 
  select(landing_pad, insert_name, rel_mTurq, exp_type)


tib_comb_2 = rbind(tib_comb, tib_bins_for_lps_boots) %>% mutate(insert_name = factor(insert_name, levels = c("none", "WT", "Sox2P", "CBS1x_Fw_Sox2P", "CBS3x_Fw_Sox2P", "CBS3x_Rv_Sox2P")))

tib_comb_2 %>% drop_na() %>%
  ggplot(aes(x = exp_type, y = rel_mTurq, col = exp_type)) +
  facet_nested(~ landing_pad + insert_name) +
  geom_point(width = 0.2, size = 2) +
  # add mean line
  stat_summary(data = tib_comb_2 %>% filter(exp_type == "clones") %>% drop_na(),
    fun = "mean", geom = "crossbar", 
               width = 0.5, color = "black", 
               aes(y = rel_mTurq), fatten = 1) +
  geom_hline(yintercept = 1, col = "black", linetype = "dotted") + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# Plot for paper
pLP23 = 
tib_comb_2 %>% drop_na() %>%
  filter(exp_type != "expr_score" & landing_pad == "landingPad_23") %>%
  ggplot(aes(x = exp_type, y = rel_mTurq, col = exp_type)) +
  facet_nested(~ landing_pad + insert_name) +
  geom_point(width = 0.2, size = 3, alpha = 0.6) +
  # add mean line
  stat_summary(data = tib_comb_2 %>% filter(exp_type == "clones" & landing_pad == "landingPad_23") %>% drop_na(),
    fun = "mean", geom = "crossbar", 
               width = 0.5, color = "black", 
               aes(y = rel_mTurq), fatten = 1) +
  geom_hline(yintercept = 1, col = "black", linetype = "dotted") + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(legend.position = "none") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

pLP23

pLP8 = 
tib_comb_2 %>% drop_na() %>%
  filter(exp_type != "expr_score" & landing_pad == "landingPad_8") %>%
  ggplot(aes(x = exp_type, y = rel_mTurq, col = exp_type)) +
  facet_nested(~ landing_pad + insert_name) +
  geom_point(width = 0.2, size = 3, alpha = 0.6) +
  # add mean line
  stat_summary(data = tib_comb_2 %>% filter(exp_type == "clones" & landing_pad == "landingPad_8") %>% drop_na(),
    fun = "mean", geom = "crossbar", 
               width = 0.5, color = "black", 
               aes(y = rel_mTurq), fatten = 1) +
  geom_hline(yintercept = 1, col = "black", linetype = "dotted") + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(legend.position = "bottom")

pLP8

plot_grid(pLP23, pLP8, ncol = 1, align = "v", rel_heights = c(0.55, 1))
```


