---
title: "ME20260304_DelDE_FACS_plot"
author: "Christine Moene"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: 
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: inline
---

# Method
Here we want to import the xml FACS file from the sorting. This displays the gates that have been sorted and the populations in that gate. 

#Compatibility problems
The newer ggplot version is not compatible with the plotting in the ggcyto package, therefore i loaded an older version of ggplot in order to be compatible with the functions of ggcyto.
```{r}
# Load older ggplot version that is compatible with the GENOVA package
.libPaths(c("/DATA/usr/m.eder/R/ggplot2_3x", .libPaths()))

library(ggplot2)
packageVersion("ggplot2")  # should now show 3.5.1
```

# setup
## libraries
```{r load_libraries}

# Load dependencies
library(GenomicRanges)
library(rtracklayer)
library(tidyverse)
library(dtplyr)
library(tidyr)
library(ggbeeswarm)
library(ggplot2)
library(Biostrings)
library(stringr)
library(readr)
library(knitr)
library(RColorBrewer)
library(RcppRoll)
library(plotly)
library(readxl)
library(ggpubr)
library(ggpmisc)
library(ggridges)
library(dplyr)
library(import)
library(ggh4x)
# import::from(flowCore, .except=c("filter")) #I don't load the whole flowCore and ggcyto libraries, because they overwrite dplyr's filter function :0
import::from(ggcyto, 'fortify_fs')
library(flowDensity)
library(flowWorkspace)
#to install this you need to manually install RFOC first:
#install_url('https://cran.r-project.org/src/contrib/Archive/RFOC/RFOC_3.4-6.tar.gz')
library(caTools)
library(ggrastr)
library(cowplot)

split_string <- function(vect,sep,N1,N2=N1){
  library(stringr)
  sapply(vect, function(X){
    paste(str_split(X,sep)[[1]][N1:N2],collapse = sep)
  },USE.NAMES = F)
}
```
# Paths & parameters
```{r paths_parameters}
#datatag
datetag = paste0("ME",format(Sys.time(), '%Y%m%d'))

# Prepare output 
output_dir <- paste0("/DATA/projects/CTCF_Sox2/Scripts/for_figures/Figure_7/analysis_FACS_sorting_",datetag)
dir.create(output_dir, showWarnings = FALSE)

```


```{r knits setup}

library(knitr)
opts_chunk$set(fig.width = 20, 
               fig.height = 8, 
               cache = T,
               message = F, 
               warning = F,
               dev=c('png', 'pdf'), 
               dpi = 600,
               fig.path = file.path(output_dir, "figures/")) 
pdf.options(useDingbats = FALSE)
```


#load FACSDiva file
```{r}
library(openCyto) 
library(CytoML)
library(flowWorkspace)

diva_xml_file = "/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/E2270_hopping_mTurq_6gates_delDE/fcs_for_R_sorting/Mathias (BvS) 20230523_E2270_DelD_DelE.xml"

diva_ws <- open_diva_xml(diva_xml_file)
  

gs <- diva_to_gatingset(diva_ws,
                        name = 1, #the group to import
                        path = "/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/E2270_hopping_mTurq_6gates_delDE/fcs_for_R_sorting/raw_fcs",
                        swap_cols = F, #apparently we don't need to swap the FSC/SSC-H/W here (this is a known diva 'bug', if you import the xml into flowjo the columns are swapped but apparently not here) 
                        worksheet = "global",
                        verbose = F,
                        execute = T)

pdata_tib = pData(gs) %>%
  as_tibble() %>%
  mutate(name_short = str_remove(str_remove(name, "ME20230523_E2270_DelD_DelE_sort_"), "_([0-9]){3}.fcs"),
         cell_line = case_when(grepl("D2", name_short) ~ "B_deltaCBS-in-SCR",
                               grepl("E12", name_short) ~ "C_deltaCBS-after-SCR",
                               grepl("neg_ctrl", name_short) ~ "WT",
                               .default = "A_Sox2P"),
         treatment = case_when(grepl("PB", name_short) ~ "PB",
                               grepl("SB", name_short) ~ "SB",
                               .default = "untreated"),
         recording = case_when(grepl("_sort", name_short) ~ "sorting",
                               .default = "pre-recording")
         )

pdata_df = as.data.frame(pdata_tib)
rownames(pdata_df) = pdata_df$name

pData(gs) = pdata_df


```

ggcyto works very well for plotting (with the axis_x/y_inverse_trans we get the actual fluorescent values ont he axes again.)

```{r}
# 
# ggcyto::ggcyto(gs, aes(x = "GFP", y = "mTurqoise2"), subset = "mCh_pos") +
#   ggcyto::axis_x_inverse_trans() +
#   ggcyto::axis_y_inverse_trans() +
#   geom_hex(bins = 126, aes( fill = after_stat(ncount))) +
#   geom_gate(paste0("gate", 1:6, "-1"), col = 'black') +
#   theme_classic() +
#   ggcyto_par_set(limits = "data")
```
 
```{r, plot_prerecorded_only, height = 14, width = 7}
gs_oi = subset(gs, cell_line != "WT" & treatment != "untreated" & recording == "pre-recording") 
#plot the data from the recording, not the sorting, so we have a similar number of cells for PB and SB (SB we recorded many more during the sorting)

ggcyto::ggcyto(gs_oi, aes(x = "GFP", y = "mTurqoise2"), subset = "mCh_pos") +
  #facet by cell line / treatment
  facet_grid(cell_line ~ treatment) +
  
  #plot cells
  geom_hex(bins = 126, aes( fill = after_stat(ncount))) +
  scale_fill_distiller(palette = 'Spectral') + #manually adding a fill scale also forces the scale to be linear again.
  
  #plot_gates
  ggcyto::geom_gate(paste0("gate", 1:6, "-1"), col = 'black') +
  ggcyto::geom_stats(adjust = c(-0.3, 0.5), digits = 3, size = 3) +
    
  #layout
  theme_bw(base_size = 14) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        aspect.ratio = 1) +
  ggtitle(element_blank())+
  ggcyto::axis_x_inverse_trans() +
  ggcyto::axis_y_inverse_trans() +
  # scale_x_continuous(expand = expansion(mult = c(0, 0.2)))+
  scale_y_continuous(expand = expansion(mult = 0.1)) +
  ggcyto::labs_cyto(labels = "marker")



```
The digits command in geom_stats doesn't really work (it changes the number of digits, but not to the same amount for each gate).
Round down to max 3 digits in illustrator


```{r, plot_all_data, height = 21, width = 7}
gs_oi_all = subset(gs, cell_line != "WT" & treatment != "untreated") 
#plot the data from the recording, not the sorting, so we have a similar number of cells for PB and SB (SB we recorded many more during the sorting)

ggcyto::ggcyto(gs_oi_all, aes(x = "GFP", y = "mTurqoise2"), subset = "mCh_pos") +
  #facet by cell line / treatment
  facet_grid(cell_line ~ treatment) +
  
  #plot cells
  geom_hex(bins = 126, aes( fill = after_stat(ncount))) +
  scale_fill_distiller(palette = 'Spectral') +
  
  #plot_gates
  ggcyto::geom_gate(paste0("gate", 1:6, "-1"), col = 'black') +
  ggcyto::geom_stats(adjust = c(-0.3, 0.5), digits = 3, size = 3) +
  
  #layout
  theme_bw(base_size = 14) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        aspect.ratio = 1) +
  ggtitle(element_blank())+
  ggcyto::axis_x_inverse_trans() +
  ggcyto::axis_y_inverse_trans() +
  # scale_x_continuous(expand = expansion(mult = c(0, 0.2)))+
  scale_y_continuous(expand = expansion(mult = 0.1)) +
  ggcyto::labs_cyto(labels = "marker")



```

```{r}
gs_oi_all = subset(gs, cell_line != "WT" & treatment == "SB") 
#plot the data from the recording, not the sorting, so we have a similar number of cells for PB and SB (SB we recorded many more during the sorting)

ggcyto::ggcyto(gs_oi_all, aes(x = "GFP", y = "mTurqoise2"), subset = "mCh_pos") +
  #facet by cell line / treatment
  facet_grid(treatment ~ cell_line) +
  
  #plot cells
  geom_hex(bins = 126, aes( fill = after_stat(ncount))) +
  scale_fill_distiller(palette = 'Spectral') +
  
  #plot_gates
  ggcyto::geom_gate(paste0("gate", 1:6, "-1"), col = 'black') +
  ggcyto::geom_stats(adjust = c(-0.3, 0.5), digits = 3, size = 3) +
  
  #layout
  theme_bw(base_size = 14) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        aspect.ratio = 1) +
  ggtitle(element_blank())+
  ggcyto::axis_x_inverse_trans() +
  ggcyto::axis_y_inverse_trans() +
  # scale_x_continuous(expand = expansion(mult = c(0, 0.2)))+
  scale_y_continuous(expand = expansion(mult = 0.1)) +
  ggcyto::labs_cyto(labels = "marker")

```





