---
title: "ME20260304_DelD_E_clones_FACS"
date: "2024-04-05"
output: 
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---

# Method
We generated and genotyped different deltaCBS-in-SCR and deltaCBS-DS-SCR clones and measured their fluorescence by analytical FACS (Fortessa 1). 
Here I am loading the FACS data and plot the values. 

#Compatibility problems
The newer ggplot version is not compatible with the plotting in the ggcyto package, therefore i loaded an older version of ggplot in order to be compatible with the functions of ggcyto.
```{r}
# Load older ggplot version that is compatible with the GENOVA package
.libPaths(c("/DATA/usr/m.eder/R/ggplot2_3x", .libPaths()))

library(ggplot2)
packageVersion("ggplot2")  # should now show 3.5.1
```

# Set-up
Load the libraries and set the parameters.
```{r load_libraries}

# Load dependencies
library(GenomicRanges)
library(rtracklayer)
library(tidyverse)
library(dtplyr)
library(tidyr)
library(ggbeeswarm)
#library(ggplot2)
library(Biostrings)
library(stringr)
library(readr)
library(knitr)
library(RColorBrewer)
library(RcppRoll)
library(plotly)
library(readxl)
library(ggpubr)
library(ggpmisc)
library(ggridges)
library(dplyr)
library(import)
library(ggh4x)
library(cowplot)
import::from(flowCore, .except=c("filter")) #I don't load the whole flowCore and ggcyto libraries, because they overwrite dplyr's filter function :0
import::from(ggcyto, 'fortify_fs')
library(flowDensity)
#to install this you need to manually install RFOC first:
#install_url('https://cran.r-project.org/src/contrib/Archive/RFOC/RFOC_3.4-6.tar.gz')


split_string <- function(vect,sep,N1,N2=N1){
  library(stringr)
  sapply(vect, function(X){
    paste(str_split(X,sep)[[1]][N1:N2],collapse = sep)
  },USE.NAMES = F)
}
```

# Paths & parameters
```{r paths_parameters}
#datatag
datetag = paste0("ME",format(Sys.time(), '%Y%m%d'))

# Prepare output 
output_dir <- paste0("/DATA/projects/CTCF_Sox2/Scripts/for_figures/Figure_7/analysis_FACS_clones_",datetag)
dir.create(output_dir, showWarnings = FALSE)

```

```{r knits setup}

library(knitr)
opts_chunk$set(fig.width = 20, fig.height = 8, cache = T,
               message = F, warning = F,
               dev=c('png', 'pdf'), 
               fig.path = file.path(output_dir, "figures/")) 
pdf.options(useDingbats = FALSE)
```

```{r}
myCols_blue = c('#525252', "#080E5B",  "#212B71", "#3A4987", "#465893", "#5A719A", "#6C8AA0")
names(myCols_blue) = c("ctrl", paste0("P", 1:6))
colScale7_blue <- scale_colour_manual(name = "population", values = myCols_blue)
fillScale7_blue <- scale_fill_manual(name = "population", values = myCols_blue)

myCols_teal = c('#525252', "#004C4C", "#146B6B", "#2C8A8A", "#45AAAA", "#5FC9C9", "#78E8E8")
names(myCols_teal) = c("ctrl", paste0("P", 1:6))
colScale7_teal <- scale_colour_manual(name = "population", values = myCols_teal)
fillScale7_teal <- scale_fill_manual(name = "population", values = myCols_teal)

myCols_red = c('#525252', "#5B0000", "#751919", "#902F2F", "#AA4646", "#C45C5C", "#DE7373")
names(myCols_red) = c("ctrl", paste0("P", 1:6))
colScale7_red <- scale_colour_manual(name = "population", values = myCols_red)
fillScale7_red <- scale_fill_manual(name = "population", values = myCols_red)

# Combine all colors into a single vector
all_colors = c(myCols_blue, myCols_teal, myCols_red)

# Create the names dynamically for the combined vector
names(all_colors) = c(paste0("34A_P",1:6), 
                      paste0("D2_P",1:6), 
                      paste0("E12_P",1:6))

# Since ctrl is common across all, you may need to add it separately if it's used
all_colors["ctrl"] = '#525252'
colScaleAll <- scale_colour_manual(name = "col_name", values = all_colors)


# col scheme for cell line (one color per cell line)
myCols_23_34A = c("#3A4987")
myCols_23_34A_DelD2 = c('#2C8A8A')
myCols_23_34A_DelE12 = c("#902F2F")


cell_line_col = c(myCols_23_34A, myCols_23_34A_DelD2, myCols_23_34A_DelE12)

names(cell_line_col) = c("23_34A",
                         "Sox2P_DelD2",
                         "Sox2P_DelE12")
                      
colScaleCell_line <- scale_colour_manual(name = "cell_line", values = cell_line_col)
fillScaleCell_line <- scale_fill_manual(name = "cell_line", values = cell_line_col)
``` 

# Load data
## Load E2254
```{r load_E2254}

fcs_files_E2254 = dir('/DATA/projects/CTCF_Sox2/FACS_files/export_fcs_R_E2254_DelD_DelE/fcs_for_R_timegated',
                      full.names = T)

fcs_files_E2254_tib = tibble(file = fcs_files_E2254)%>%
  mutate(sample_name_tmp = split_string(str_remove(file, '.*/'), "_", 2, 9)) %>%
  mutate(cell_line_tmp = split_string(sample_name_tmp, "_", 2, 4),
         cell_line_tmp2 = case_when(grepl("DelA", cell_line_tmp) ~ "23_34_DelA",
                                    grepl("WT", cell_line_tmp) ~ "WT",
                                    grepl("8_34_8", cell_line_tmp) ~ "8_34",
                                    grepl("D12", cell_line_tmp) ~ "23_34_D12",
                                    grepl("D13", cell_line_tmp) ~ "23_34_D13",
                                    grepl("D2", cell_line_tmp) ~ "23_34_D2",
                                    grepl("E12", cell_line_tmp) ~ "23_34_E12",
                                    grepl("F121", cell_line_tmp) ~ "F121",
                                    grepl("23_34A", cell_line_tmp) ~ "23_34A",
                                    .default = cell_line_tmp),
         cell_line = case_when(cell_line_tmp == "23_34_E12" ~ "WT",
                               cell_line_tmp == "E12_D12_D12" ~ "23_34_E12",
                               .default = cell_line_tmp2),
         treatment = "untreated",
         
         sample_number = as.numeric(substr(str_extract(sample_name_tmp, "\\d{3}_Time"), 1, 3)),  #regex: 3 digits, followed by _Time, then take the 3 digits
         genotype = case_when(substr(split_string(sample_name_tmp, "_", 2),1,1) == "D" ~ "delD",
                              substr(split_string(sample_name_tmp, "_", 2),1,1) == "E" ~ "delE",
                              sample_number <= 27 ~ "delA",
                              .default = "WT"),
         
         sorted_phenotype = case_when(genotype == "delA" ~ "mCh-_mT+",
                                      .default = "none"),
         type = case_when(cell_line %in% c("F121", "WT") ~ "control",
                          genotype == "untreated" ~ "control",
                          .default = "clone"),
         wellID = case_when(type == "clone" & genotype == "delA" ~ split_string(sample_name_tmp, "_", 5, 5)),
         rownum = case_when(substr(wellID, 1, 1) == "A" ~ 1,
                            substr(wellID, 1, 1) == "B" ~ 2,
                            substr(wellID, 1, 1) == "C" ~ 3),
         colnum = as.numeric(substr(wellID, 2, 3)),
         clone_number = 12*(rownum - 1) + colnum,
         clone_name = case_when(
           genotype %in% c('delD', 'delE') ~ paste0("clone_", split_string(sample_name_tmp, "_", 2)),
           type == "clone" & genotype == "delA" ~ paste0("clone_A",clone_number),
           .default = cell_line)) %>%
  mutate(timepoint = "remeasurement",
         experiment = "E2254",
         flowcytometer = "analytical",
         date = split_string(sample_name_tmp, "_",1)) %>%
  group_by(cell_line, genotype, clone_name) %>%
  arrange(sample_number) %>%
  mutate(techrep = row_number(),
         sample_name = case_when(type == "clone" & genotype == "delA" ~  paste0(experiment, "_", cell_line, "_", genotype, "_", clone_name),
                                 type == "clone" & genotype %in% c("delD", "delE") ~  paste0(experiment, "_", cell_line, "_", genotype, "_", clone_name, "_", techrep),
                                 .default = paste0(experiment, "_", cell_line, "_", genotype, "_", techrep))) %>%
  select(-c("cell_line_tmp", "cell_line_tmp2"))

```

## Load E2270
```{r load_E2270}
fcs_files_E2270 = dir('/DATA/projects/CTCF_Sox2/FACS_files/export_fcs_R_E2270_DelD_DelE', 
                      full.names = T)

fcs_files_E2270_tib = tibble(file = fcs_files_E2270) %>%
  mutate(sample_name_tmp = split_string(str_remove(file, '.*/'), "_", 4, 10)) %>%
  mutate(cell_line = case_when(grepl("E12", sample_name_tmp) ~ "23_34_E12",
                               grepl("D2", sample_name_tmp) ~ "23_34_D2",
                               grepl("23_34A", sample_name_tmp) ~ "23_34A",
                               grepl("sort_neg_ctrl", sample_name_tmp) ~ "WT"),
         
         treatment = case_when(grepl("UT", sample_name_tmp) ~ "untreated",
                               grepl("PB", sample_name_tmp) ~ "PB",
                               grepl("SB", sample_name_tmp) ~ "SB",
                               grepl("sort_neg_ctrl", sample_name_tmp) ~ "untreated"),
         
         genotype = case_when(cell_line == "23_34A" ~ "WT",
                              cell_line == "23_34_D2" ~ "delD",
                              cell_line == "23_34_E12" ~ "delE",
                              .default = "WT"),
         
         clone_name = case_when(grepl("D2", cell_line) ~ "clone_D2",
                                grepl("E12", cell_line) ~ "clone_E12",
                                grepl("23_34A", cell_line) ~ "23_34A",
                                .default = genotype),
         fcs_filter = "mCh+",
         timepoint = "sort",
         experiment = "E2270",
         flowcytometer = "sorter",
         sample_name = paste0(experiment, "_", cell_line, "_", genotype, "_", treatment),
         date = "me20230523")

```

## Load E2339
```{r load_E2339}

fcs_files_E2339= dir("/DATA/projects/CTCF_Sox2/FACS_files/export_fcs_R_E2339_DelD_DelE", 
                      full.names = T)

fcs_files_E2339_tib = tibble(file = fcs_files_E2339) %>%
  mutate(sample_name_tmp = split_string(str_remove(file, '.*/'), "_", 2, 8)) %>%
  mutate(cell_line = case_when(grepl("23_34A_ctrl", sample_name_tmp) ~ "23_34A",
                               grepl("23_DelD", sample_name_tmp) ~ "23_34_D2",
                               grepl("23_DelE", sample_name_tmp) ~ "23_34_E12",
                               grepl("F121_ctrl", sample_name_tmp) ~ "F121",
                               grepl("WT", sample_name_tmp) ~ "WT"),
         
          treatment = case_when(grepl("UT", sample_name_tmp) ~ "untreated",
                               grepl("PB", sample_name_tmp) ~ "PB",
                               grepl("SB", sample_name_tmp) ~ "SB",
                               grepl("ctrl", sample_name_tmp) ~ "untreated",
                               grepl("WT", sample_name_tmp) ~ "untreated"),
         
         genotype = case_when(cell_line == "23_34A" ~ "WT",
                              cell_line == "23_34_D2" ~ "delD",
                              cell_line == "23_34_E12" ~ "delE",
                              .default = "WT"),
         
         clone_name = case_when(grepl("23_DelD", sample_name_tmp) ~ "clone_D2",
                                grepl("23_DelE", sample_name_tmp) ~ "clone_E12",
                                .default = cell_line),
         fcs_filter = "mCh+",
         timepoint = "sort",
         experiment = "E2339",
         flowcytometer = "sorter",
         sample_name = paste0(experiment, "_", cell_line, "_", genotype, "_", treatment),
         date = "me20231024"
         )
```

## Combine data
```{r}
fcs_files_combined = bind_rows(fcs_files_E2254_tib,
                 fcs_files_E2270_tib,
                 fcs_files_E2339_tib) %>%
  mutate(cell_line = factor(cell_line, c("23_34A", "23_34_DelA", "23_34_D2", "23_34_D12", "23_34_D13", "23_34_E12","8_34", "F121", "WT"))) %>%
  mutate(treatment = factor(treatment, c('untreated','PB','SB')))

# saveRDS(fcs_files_combined, '/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/Combined_for_paper/CM20240118_Sox2_CRISPR_fcs_files_tib.rds')

#not all channels are in all datasets, subset the ones we need (which are in all sets) so we can load all at once
flowset_all = read.flowSet(files = fcs_files_combined$file, alter.names = T, truncate_max_range = F, column.pattern = 'BL.B..530_30.A|YG.D..610_20.A|V.F..450_50.A|FSC.A|SSC.A')
```


```{r}
#add sample names and phenodata
sampleNames(flowset_all) = fcs_files_combined$sample_name
pData(flowset_all)$name = fcs_files_combined$sample_name
pData(flowset_all)$sample_name = fcs_files_combined$sample_name

pheno_data_tib = left_join(pData(flowset_all), fcs_files_combined, join_by(sample_name == sample_name) ) 
rownames(pheno_data_tib) = fcs_files_combined$sample_name

pData(flowset_all) = pheno_data_tib

#order the fluorophores and rename them
flowset_fluo = flowset_all[,c('BL.B..530_30.A', 'YG.D..610_20.A', 'V.F..450_50.A', 'FSC.A', 'SSC.A')]
colnames(flowset_fluo) = c("GFP", "mCherry", "mTurq", 'FSC', 'SSC')

N_cells = sapply(1:length(flowset_fluo), function(i){nrow(flowset_fluo[[i]])})
pData(flowset_fluo)$N_cells = N_cells
fcs_files_combined$N_cells = N_cells

rm(flowset_all)
```

# Setup colour scale
```{r}

# col scheme for cell line
myCols_23_34A = c("#3A4987")
myCols_8 = c("#3A4987")
myCols_23_34A_DelD2 = c('#45AAAA')
myCols_23_34A_DelD12 = c('#45AAAA')
myCols_23_34A_DelD13 = c('#78E8E8')
myCols_23_34A_DelE12 = c("#902F2F")
myCols_F121 = c("grey")
myCols_WT = c("darkgrey")

cell_line_col = c(myCols_23_34A, myCols_8, myCols_23_34A_DelD2, myCols_23_34A_DelD12, myCols_23_34A_DelD13, myCols_23_34A_DelE12,myCols_F121, myCols_WT)

names(cell_line_col) = c("23_34A",
                         "8_34",
                         "23_34_D2",
                         "23_34_D12", 
                         "23_34_D13",
                         "23_34_E12",
                         "F121",
                         "WT")
                      
colScaleCell_line <- scale_colour_manual(name = "cell_line", values = cell_line_col)
fillScaleCell_line <- scale_fill_manual(name = "cell_line", values = cell_line_col)
```

# statistics 
```{r}
# Calculate median expression
tib_median = as_tibble(rownames = "name", fsApply(flowset_fluo, each_col, median)) %>%
  rename(median_mCh = mCherry,
         median_GFP = GFP,
         median_mTurq = mTurq) %>%
  select(-c("FSC", "SSC"))

# Calculate mean expression
tib_mean = as_tibble(rownames = "name", fsApply(flowset_fluo, each_col, mean)) %>%
  rename(mean_mCh = mCherry,
         mean_GFP = GFP,
         mean_mTurq = mTurq) %>%
  select(-c("FSC", "SSC"))

# Calculate standard deviation
tib_std = as_tibble(rownames = "name", fsApply(flowset_fluo, each_col, sd)) %>%
  rename(std_mCh = mCherry,
         std_GFP = GFP,
         std_mTurq = mTurq) %>%
  select(-c("FSC", "SSC"))
  
tib_mean_median_std = tib_median %>% 
  left_join(tib_mean, by = "name") %>%
  left_join(tib_std, by = "name") %>%
  mutate(CV_mCh = std_mCh/mean_mCh, 
         CV_GFP = std_GFP/mean_GFP,
         CV_mTurq = std_mTurq/mean_mTurq)

tib_all = pheno_data_tib %>% left_join(tib_mean_median_std, by = "name")
```


# correct and normalize medians
```{r}
#subtract autofluorescence
autofluo_tib_comb = tib_all  %>%
  filter(cell_line == "WT") %>%
  group_by(date) %>%
  summarize(mTurq_WT = mean(median_mTurq),
           mCherry_WT = mean(median_mCh),
            GFP_WT = mean(median_GFP) )

#normalize to GFP
expression_tib_comb = tib_all %>%
  left_join(autofluo_tib_comb, join_by(date)) %>%
  mutate(median_mTurq_cor = median_mTurq - mTurq_WT,
         median_mCherry_cor = median_mCh - mCherry_WT,
         median_GFP_cor = median_GFP - GFP_WT) %>%
  mutate(mTurq_norm = median_mTurq_cor / median_GFP_cor,
         mCherry_norm = median_mCherry_cor / median_GFP_cor)

#normalize to 23_34A (to compare between measurement days and to compare them in general)
standard_tib = expression_tib_comb %>%
  group_by(date, experiment) %>%
  filter(cell_line == "23_34A" & treatment == "untreated") %>%
  summarize(mTurq_norm_standard = mean(mTurq_norm),
           mCherry_norm_standard = mean(mCherry_norm),
            GFP_standard = mean(median_GFP_cor) )

expression_tib_comb = expression_tib_comb %>%
  left_join(standard_tib, join_by(date, experiment)) %>%
  mutate(rel_mTurq = mTurq_norm / mTurq_norm_standard,
         rel_mCherry = mCherry_norm / mCherry_norm_standard,
         rel_GFP_cor = median_GFP_cor / GFP_standard)


expression_tib_comb %>% filter(genotype != "delA")  %>%
  group_by(clone_name) %>%
  summarize(mean(median_mTurq), mean(mTurq_norm), mean(rel_mTurq))
```

# For paper
```{r FACS plot paper}
#E2254 samples
E2254_samples = fcs_files_combined %>% 
  filter(experiment == "E2254" & treatment != "delA" & clone_name %in% c("F121",  "23_34A", "clone_E12", "clone_D2")) %>%
  pull(sample_name)

# Ridgeline plots D2, E12 with controls
pRidge_mTurq =
  ggplot(flowset_fluo[E2254_samples,], aes(x = mTurq, 
                                           y = factor(cell_line, levels = c("23_34_E12", "23_34_D2","23_34A","F121")), 
                                           fill = cell_line)) + #, col = as_factor(sort2))) + 
  geom_density_ridges(size =0.5, # rel_min_height  = 0.005,
                      # scale = 2,
                      # alpha = 0.75,
                      quantile_lines = TRUE, quantiles = 2) +
  ggcyto::scale_x_flowjo_biexp(widthBasis = -100, breaks = c(-10^3, 0, 10^3, 10^4, 10^5),
                               pos = 4.42) +
  
  theme_bw() +
  ggtitle("mTurq")+
  fillScaleCell_line  +
  theme(legend.position = "none",
        axis.title.y = element_blank())

pRidge_mTurq

pRidge_GFP =
  ggplot(flowset_fluo[E2254_samples,], aes(x = GFP,
                                           y = factor(cell_line, levels = c("23_34_E12", "23_34_D2","23_34A","F121")), 
                                           fill = cell_line)) + #, col = as_factor(sort2))) + 
  #geom_density_ridges() +
  geom_density_ridges(size =0.5, # rel_min_height  = 0.005,
                      # scale = 2,
                      # alpha = 0.75,
                      quantile_lines = TRUE, quantiles = 2) +
  ggcyto::scale_x_flowjo_biexp(widthBasis = -100, breaks = c(-10^3, 0, 10^3, 10^4, 10^5),
                               pos = 4.42) +
  theme_bw() +
  ggtitle("Sox2::GFP")+
  fillScaleCell_line  +
  theme(legend.position = "none",
        axis.text.y = element_blank(),
        axis.title.y = element_blank())
pRidge_GFP

pRidge_mCherry =
  ggplot(flowset_fluo[E2254_samples,], aes(x = mCherry, 
                                           y = factor(cell_line, levels = c("23_34_E12", "23_34_D2","23_34A","F121")), 
                                           fill = cell_line)) + #, col = as_factor(sort2))) + 
  geom_density_ridges(size =0.5, # rel_min_height  = 0.005,
                      # scale = 2,
                      # alpha = 0.75,
                      quantile_lines = TRUE, quantiles = 2) +
  ggcyto::scale_x_flowjo_biexp(widthBasis = -100, breaks = c(-10^3, 0, 10^3, 10^4, 10^5),
                               pos = 4.42) +
  theme_bw() +
  ggtitle("Sox2::mCherry")+
  fillScaleCell_line  +
  theme(legend.position = "none",
        axis.text.y = element_blank(),
        axis.title.y = element_blank())
pRidge_mCherry

plot_grid(pRidge_mTurq, pRidge_GFP, pRidge_mCherry, nrow = 1, axis = "lr", align = "v")
```

## How much of an reduction
```{r}
expression_tib_comb %>% filter(experiment == "E2254" & genotype != "delA") %>%
  group_by(cell_line) %>%
  summarise(mean_rel_mT = mean(rel_mTurq),
            mean_rel_mCh = mean(rel_mCherry))

expression_tib_comb %>% filter(experiment == "E2254" & genotype != "delA") %>%
  group_by(cell_line) %>%
  summarise(mean_rel_mT = mean(rel_mTurq),
            mean_rel_mCh = mean(rel_mCherry)) %>%
  filter(cell_line != "WT" & cell_line != "8_34") %>%
  ggplot(aes(x = factor(cell_line, levels = c("F121", "23_34A", "23_34_D2", "23_34_D12", "23_34_D13", "23_34_E12")), 
             y = mean_rel_mT, fill = cell_line)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.8) +
  geom_hline(yintercept = 1, col = "black", linetype = "dotted") +
  geom_jitter(data = expression_tib_comb %>% filter(experiment == "E2254" & genotype != "delA" & cell_line != "WT" & cell_line != "8_34"), 
             aes(x = cell_line, y = rel_mTurq, col = cell_line), size = 2) +
  theme_bw() +
  ggtitle("Relative reporter expression", subtitle = "autofluo cor & norm to 23_34A") +
  fillScaleCell_line  +
  colScaleCell_line + 
  theme(legend.position = "none")


expression_tib_comb %>% filter(experiment == "E2254" & genotype != "delA") %>%
  group_by(cell_line) %>%
  summarise(mean_rel_mT = mean(rel_mTurq),
            mean_rel_mCh = mean(rel_mCherry)) %>%
  filter(cell_line != "WT" & cell_line != "8_34") %>%
  ggplot(aes(x = factor(cell_line, levels = c("F121", "23_34A", "23_34_D2", "23_34_D12", "23_34_D13", "23_34_E12")), 
             y = mean_rel_mCh, fill = cell_line)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.8) +
  geom_hline(yintercept = 1, col = "black", linetype = "dotted") +
  geom_jitter(data = expression_tib_comb %>% filter(experiment == "E2254" & genotype != "delA" & cell_line != "WT" & cell_line != "8_34"), 
             aes(x = cell_line, y = rel_mCherry, col = cell_line), size = 2) +
  theme_bw() +
  ggtitle("Relative Sox2::mCherry expression", subtitle = "autofluo cor & norm to 23_34A")+
  fillScaleCell_line  +
  colScaleCell_line + 
  theme(legend.position = "none")
        
```


## Supplementary plot
```{r supplementary plot}

E2254_samples_clone = fcs_files_combined %>% 
  filter(experiment == "E2254" & treatment != "delA" & cell_line %in% c("F121",  "23_34A", "23_34_D2", "23_34_D12", "23_34_D13", "23_34_E12")) %>%
  pull(sample_name)


# Ridgeline plots D2, E12 with controls
pRidge_mTurq =
  ggplot(flowset_fluo[E2254_samples_clone,], aes(x = mTurq, 
                                                 y = factor(cell_line, levels = c("23_34_E12", "23_34_D13", "23_34_D12","23_34_D2","23_34A","F121")), 
                                                 fill = cell_line)) + 
  #geom_density_ridges() +
  stat_density_ridges(size =0.5, # rel_min_height  = 0.005,
                      quantile_lines = TRUE, quantiles = 2) +
  ggcyto::scale_x_flowjo_biexp(widthBasis = -100, breaks = c(-10^3, 0, 10^3, 10^4, 10^5),
                               pos = 4.42) +
  theme_bw() +
  ggtitle("mTurq")+
  fillScaleCell_line  +
  theme(legend.position = "none",
        axis.title.y = element_blank())

pRidge_GFP =
  ggplot(flowset_fluo[E2254_samples_clone,], aes(x = GFP,  
                                                 y = factor(cell_line, levels = c("23_34_E12", "23_34_D13", "23_34_D12","23_34_D2","23_34A","F121")), 
                                                 fill = cell_line)) + #, col = as_factor(sort2))) + 
  #geom_density_ridges() +
  stat_density_ridges(size =0.5, # rel_min_height  = 0.005,
                      quantile_lines = TRUE, quantiles = 2) +
  ggcyto::scale_x_flowjo_biexp(widthBasis = -100, breaks = c(-10^3, 0, 10^3, 10^4, 10^5),
                               pos = 4.42) +
  theme_bw() +
  ggtitle("Sox2::GFP")+
  fillScaleCell_line  +
  theme(legend.position = "none",
        axis.text.y = element_blank(),
        axis.title.y = element_blank())

pRidge_mCherry =
  ggplot(flowset_fluo[E2254_samples_clone,], aes(x = mCherry,  
                                                 y = factor(cell_line, levels = c("23_34_E12", "23_34_D13", "23_34_D12","23_34_D2","23_34A","F121")), 
                                                 fill = cell_line)) + #, col = as_factor(sort2))) + 
  #geom_density_ridges() +
  stat_density_ridges(size =0.5, # rel_min_height  = 0.005,
                      quantile_lines = TRUE, quantiles = 2) +
  ggcyto::scale_x_flowjo_biexp(widthBasis = -100, breaks = c(-10^3, 0, 10^3, 10^4, 10^5),
                               pos = 4.42) +
  theme_bw() +
  ggtitle("Sox2::mCherry")+
  fillScaleCell_line  +
  theme(legend.position = "none",
        axis.text.y = element_blank(),
        axis.title.y = element_blank())

plot_grid(pRidge_mTurq, pRidge_GFP, pRidge_mCherry, nrow = 1, axis = "lr", align = "v")
```



