---
title: "me20250724_DelD_DelE"
author: "Mathias Eder"
date: "2026-03-03"
output: 
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
    editor_options:
    chunk_output_type: console
---

# Aim 
In this script we analyse the Sox2P hopping insert in the WT background and the delta-CBS-in-SCR or delta-CBS-DS-SCR cell lines. We curate the dataset and plot the distribution of integrations. Further we calculate the enrichment score and compare them to the Sox2P control cell line. 

The Sox2P control integrations were published previoulsy here:
https://www.science.org/doi/10.1126/science.ads6552


# Set-up
Load the libraries and set the parameters.
```{r load_libraries}

# Load dependencies
library(GenomicRanges)
library(rtracklayer)
library(gtools)
library(tidyverse)
library(dtplyr)
library(dplyr)
library(tidyr)
library(ggbeeswarm)
library(ggplot2)
#library(broom)
library(Biostrings)
library(stringr)
library(readr)
#library(tagMeppr)
#library(withr)
library(knitr)
library(RColorBrewer)
#library(tagmapAnalyseR)
# library(UpSetR)
# library(ComplexHeatmap)
library(purrr)
# library(ggpubr)
# library(cowplot)
library(RcppRoll)
library(viridis)
library(plotly)
library(GGally)
library(cowplot)
library(ggh4x)
library(fuzzyjoin)
library(GENOVA)

split_string <- function(vect,sep,N1,N2=N1){
  library(stringr)
  sapply(vect, function(X){
    paste(str_split(X,sep)[[1]][N1:N2],collapse = sep)
  },USE.NAMES = F)
}
```

# Paths & parameters
```{r paths_parameters}
setwd("/DATA/usr/m.eder/GitHub/CBS_Sox2_manuscript/Figures/Figure_7/")

#datatag
datetag = paste0("ME",format(Sys.time(), '%Y%m%d'))

# Prepare output 
output_dir <- paste0("/DATA/projects/CTCF_Sox2/Scripts/for_figures/Figure_7/analysis_",datetag)
dir.create(output_dir, showWarnings = FALSE)

```

```{r knits setup}

opts_chunk$set(fig.width = 20, fig.height = 8, cache = T,
               message = F, 
               warning = F,
               dev=c('png', 'pdf'), 
               dpi = 600,
               fig.path = file.path(output_dir, "/figures_DelDE/")) 
pdf.options(useDingbats = FALSE)
```


# File paths
```{r}
#Sox2P integrations
path_tib_23_ints = "/DATA/projects/Sox2/GEO_collection/Tagmentation/still_to_upload/mapped_integrations_-116kb_Sox2P_pools.txt"

#delta CBS integrations
path_tib_DelDE_ints = "/DATA/projects/CTCF_Sox2/Integrations_files/mapped_integrations_-116kb_Sox2P_delCBS_pools.txt"


#Sorting stats
path_median_mTurq_relevant_cell_lines = "/DATA/projects/CTCF_Sox2/Info_files/me20250707_median_mTurq_mCherry_GFP_sorting_all.tsv"
path_sorting_stats = "/DATA/projects/CTCF_Sox2/Info_files/me20250707_stats_for_expression_score_mCherry.tsv"

#CTCF ChIP-seq data
path_CTCF_sites = "/DATA/usr/v.franceschini/Workspaces/2024_01_MATHIAS_CTCF_PEAKS/VF240812_calling_CTCF_motifs_again/02_OUTPUTS/CTCF_sites/6764_2_ME_E2_CGATGT_S2_R1_001_peaks_motifs.bed"
```

#Load functions
```{r load functions}
source("/DATA/projects/Sox2/General_functions/CM20240621_general_plotting_functions.R")
source("/DATA/projects/Sox2/General_functions/CM20250318_new_expression_score_functions.R")
source("/DATA/projects/CTCF_Sox2/functions/ME20250707_mCh_new_expression_score_functions.R")
source('/DATA/projects/Sox2/General_functions/CM20250228_expression_score_functions.R') #for expr score calculations of abitrary bins
```

# Relevant annotations
```{r element_coordinates}

# Location of enhancer / gene
enhancer <- GRanges(seqnames = "chr3", 
                    ranges = IRanges(start = 34753415,
                                     end = 34766401),
                    strand = "*")

# SCR sequence used by Luca Giorgetti study
enhancer_giorgetti <- GRanges(seqnames = "chr3", 
                    ranges = IRanges(start = 34757649,
                                     end = 34762535),
                    strand = "*")

#gene based on mm10, ucsc browser, RefSeq annotation
# Sox2
# chr3:34649995-34652461 (+)
# id = NM_011443.4

Sox2_gene <- GRanges(seqnames = "chr3", 
                    ranges = IRanges(start = 34649995,
                                     end = 34652461),
                    strand = "+")

#the old gene annotation, did not include the 5' UTR (so from start codon to end of 3' UTR)
gene_old <- GRanges(seqnames = "chr3", 
                    ranges = IRanges(start = 34650405,
                                     end = 34652461),
                    strand = "+")


landingPad_23 <- GRanges(seqnames = "chr3", 
                    ranges = IRanges(start = 34643960,
                                     end = 34643962),
                    strand = "+")

landingPad_8 <- GRanges(seqnames = "chr3", 
                    ranges = IRanges(start = 34721183,
                                     end = 34721184),
                    strand = "+")


LP_tib = tibble(launch_pad_location = c("-116 kb", "-39 kb"), 
                start = c( start(landingPad_23),start(landingPad_8)))

#CTCF sites based on our own mapping
# new analysis (based on mapping to mm10 only)
CTCF_mm10_new <- import.bed('/DATA/usr/v.franceschini/Workspaces/2024_01_MATHIAS_CTCF_PEAKS/VF240812_calling_CTCF_motifs_again/02_OUTPUTS/CTCF_sites/6764_2_ME_E2_CGATGT_S2_R1_001_peaks_motifs.bed')

#filtering:
CTCF_mm10.chr3 <- CTCF_mm10_new[seqnames(CTCF_mm10_new)== 'chr3']
# add the missing site after the SCR
CTCF_mm10.chr3_extra = sort(c(CTCF_mm10.chr3,
                         GRanges(seqnames = "chr3",
                                 IRanges(start = 34772210, end = 34772210),
                                 strand = "+")),
                         ignore.strand = T)


prange_zoom = c(34500000 , 35000000)
prange2 = c(34000000, 35500000)
prange_far = c(32000000, 37000000)
#zoom in on only the SCR +- 50kb
prange_LP = c(start(landingPad_23)-5E4, end(landingPad_23)+5E4)
prange_SCR = c(start(enhancer)-5E4, end(enhancer)+5E4)
prange_plot = c(32643960, 36643960) #+/- 2MB from landing pad
prange_loop = c(34631804, 34778268)
prange_zoom = c(34.5E6, 35E6)
prange_zoom_paper = c(34.59E6, 34.83E6)
prange_SCR = c(34.725E6, 34.785E6)
```

## Colour setup
```{r}
# Color annotations
## mTurq_hopping/ CTCF_mTurq_hopping
myCols_blue = c('#525252', "#080E5B",  "#212B71", "#3A4987", "#465893", "#5A719A", "#6C8AA0")

## SCR
brown = "#ffb000"
```

Specific colors for this figure
```{r}
myCols_blue = c('#525252', "#080E5B",  "#212B71", "#3A4987", "#465893", "#5A719A", "#6C8AA0")
names(myCols_blue) = c("ctrl", paste0("P", 1:6))
colScale7_blue <- scale_colour_manual(name = "population", values = myCols_blue)
fillScale7_blue <- scale_fill_manual(name = "population", values = myCols_blue)

myCols_teal = c('#525252', "#004C4C", "#146B6B", "#2C8A8A", "#45AAAA", "#5FC9C9", "#78E8E8")
names(myCols_teal) = c("ctrl", paste0("P", 1:6))
colScale7_teal <- scale_colour_manual(name = "population", values = myCols_teal)
fillScale7_teal <- scale_fill_manual(name = "population", values = myCols_teal)

myCols_red = c('#525252', "#5B0000", "#751919", "#902F2F", "#AA4646", "#C45C5C", "#DE7373")
names(myCols_red) = c("ctrl", paste0("P", 1:6))
colScale7_red <- scale_colour_manual(name = "population", values = myCols_red)
fillScale7_red <- scale_fill_manual(name = "population", values = myCols_red)

# Combine all colors into a single vector
all_colors = c(myCols_blue, myCols_teal, myCols_red)

# Create the names dynamically for the combined vector
names(all_colors) = c(paste0("34A_P",1:6), 
                      paste0("D2_P",1:6), 
                      paste0("E12_P",1:6))

# Since ctrl is common across all, you may need to add it separately if it's used
all_colors["ctrl"] = '#525252'
colScaleAll <- scale_colour_manual(name = "col_name", values = all_colors)


# col scheme for cell line (one color per cell line)
myCols_23_34A = c("#3A4987")
myCols_23_34A_DelD2 = c('#2C8A8A')
myCols_23_34A_DelE12 = c("#902F2F")


cell_line_col = c(myCols_23_34A, myCols_23_34A_DelD2, myCols_23_34A_DelE12)

names(cell_line_col) = c("23_34A",
                         "Sox2P_DelD2",
                         "Sox2P_DelE12")
                      
colScaleCell_line <- scale_colour_manual(name = "cell_line", values = cell_line_col)
fillScaleCell_line <- scale_fill_manual(name = "cell_line", values = cell_line_col)
``` 

# Load data
## Load 23_34A
```{r load 23_34A data}
#load the -116kb data (LP23)
tib_23_all_data = read_tsv(path_tib_23_ints) %>%
  filter(chr %in% paste0("chr", c(1:19, "X"))) %>%
  mutate(chr = factor(chr, levels = paste0("chr", c(1:19, "X")))) 

#the control pool in E2285 (rep3) was sequenced as 10 separate libraries, combine them into one sample
tib_23_E2285_ctrls = tib_23_all_data %>%
  filter(replicate == "rep3" & insert == "Sox2P" & sample_type == "control pool") %>%
  mutate(sample_name = "Sox2P-reporter -116 kb, unsorted control pool rep3")%>%
  group_by(chr, start, end, seq, strand, replicate, experiment, launch_pad_location, insert, genotype, sample_type, sorted_gate, pool_nr) %>%
  summarize(read_count = sum(read_count),
            read_count_1 = sum(read_count_1),
            read_count_2 = sum(read_count_2), .groups = "keep") 

tib_23 = tib_23_all_data %>%
  #select the rest of the data
  filter(!(replicate == "rep3" & insert == "Sox2P" & sample_type == "control pool")) %>%
  select(chr, start, end, seq, strand, replicate, experiment, launch_pad_location, insert, genotype, sample_type, sorted_gate, pool_nr, 
         read_count, read_count_1, read_count_2)%>%
  #add merged controls back in
  bind_rows(tib_23_E2285_ctrls) %>%
  #mark hopped integrations
  mutate(hopped = start != 34643961,
         cell_line = "23_34A") %>%
 filter(!(start >= 34721183 & start <= 34721192)) #filter contamination from LP8
```

## Load DelDE dataset
```{r}
#load the DelDE -116kb data (LP23)
tib_DelDE_data = read_tsv(path_tib_DelDE_ints) %>%
  filter(chr %in% paste0("chr", c(1:19, "X"))) %>%
  mutate(chr = factor(chr, levels = paste0("chr", c(1:19, "X"))),
         hopped = start != 34643961) %>%
  select(chr, start, end, seq, strand, replicate, experiment, launch_pad_location, insert, genotype, sample_type, sorted_gate, pool_nr, 
         read_count, read_count_1, read_count_2, hopped, cell_line)
```

## Combine Sox2P-WT and Sox2P-delCBS integrations
```{r}

filtered_tib_no_P1_dels = bind_rows(tib_DelDE_data, tib_23) %>%
  mutate(population = case_when(sorted_gate == "-" ~ "ctrl",
                                TRUE ~ sorted_gate),
         mapped_arms = case_when(read_count_1 >= 1 & read_count_2 >= 1 ~ "both_arms",
                                 read_count_1 == 0 ~ "rv_only",
                                 read_count_2 == 0 ~ "fw_only"),
         pool_nr = as.character(pool_nr)) %>%
  filter(read_count >= 2 | population == "ctrl") %>% #read cuttof per integration
  filter(strand %in% c("+", "-")) %>% #filter out unambigous strand
  filter(!(start >= 34721183 & start <= 34721192)) %>% #filter out contamination of LP8
  filter(!(population == "P1" & mapped_arms != "both_arms")) #filter out possible hopping induced deletions
  
all_exp = unique(filtered_tib_no_P1_dels$experiment)
```

# Beeswarms
## Plotting function
```{r beeswarm function}
plot_beeswarm_only = function(TIB, CL = "23_34A", EXP = all_exp, P_RANGE = prange_zoom,
                              PLOT_CTCF = F, PLOT_ANNOT = T,
                              TITLE = NULL, POINT_SIZE = 0.75, RANDOM_METHOD = "quasirandom",
                              NEXT_RANGE = NULL){
  tib_for_plot = filter(TIB, hopped & chr == "chr3" & cell_line %in% CL & experiment %in% EXP & start >= P_RANGE[1] & start <= P_RANGE[2])
  LP_tib_cl = select(tib_for_plot, c(cell_line, launch_pad_location)) %>% 
    distinct() %>% left_join(LP_tib)
  
  ggplot(filter(tib_for_plot, hopped & chr == "chr3" & cell_line %in% CL & experiment %in% EXP),
         aes(x = start, y = population, col = population)) +
    {if(length(CL)>1) facet_grid(cell_line ~ .)} +  #only facet when necessary

    {if (PLOT_ANNOT)
      list( #add elements in a list, you cannot use + inside an if statement
        # annotate enhancer
        geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown),
        annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown),
        # annotate gene
        annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf, alpha = .2, fill = 'red'),
        geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = 'red') )} +
    
     #annotate landing pad
    {if(PLOT_ANNOT & length(CL)== 1)
      geom_vline(xintercept = pull(
        filter(LP_tib, launch_pad_location %in% unique(tib_for_plot$launch_pad_location)), 
        start),
        col = 'darkgrey')} +
    {if(PLOT_ANNOT & length(CL) > 1)
       geom_vline(data = LP_tib_cl, aes(xintercept = start), col = 'darkgrey')} +

    # # annotate CTCFs
    {if(PLOT_CTCF)geom_vline(xintercept = start(CTCF_mm10.chr3_extra), col = 'darkgrey', lty="11")}+
    {if(!is.null(NEXT_RANGE)) geom_rug(data = tibble(range = NEXT_RANGE),
                                       aes(x = range),
                                       fill = 'grey',
                                       inherit.aes = F)}+
    
    
    #datapoints
    geom_quasirandom(alpha = 0.5, size = POINT_SIZE, varwidth = T, bandwidth = 0.8,
                     # shape = 16, #changes to point without border, so alpha applies to the whole point
                     method = RANDOM_METHOD,
                     stroke = NA) +
    geom_text(data = plyr::count(tib_for_plot, vars = c("cell_line", "population")), aes(label = paste("n = ", freq, sep = ""),
                                                                                         x = Inf),
              hjust = "inward", vjust = "top",
              col = 'black') +
    
    #layout:
    labs(x='Genomic coordinate (Mb)',y='sorted gate') +
    theme_classic(base_size = 16) +
    theme(legend.position = "none") +
    scale_x_continuous(breaks = scales::pretty_breaks(n = 6), labels = scales::unit_format(scale = 1E-6, accuracy = 0.01, unit=NULL),
                       limits=P_RANGE, expand=c(0,0)) +
    scale_y_discrete(limits = rev, drop=F) +
    ggtitle(TITLE) +
    ## CHANGED: choose the appropriate colour scale based on CL
    { if (length(CL)==1) {
        switch(CL,
          "23_34A"       = colScale7_blue,
          "Sox2P_DelD2"   = colScale7_teal,
          "Sox2P_DelE12"   = colScale7_red,
          colScale7_blue  # default fallback
        )
      } else {
        colScale7_blue  # when faceting multiple lines, fall back to blue (or choose your own default)
              }
    }
}

``` 

## Stranded plotting function
```{r stranded beeswarm}
plot_beeswarm_only_stranded = function(TIB, CL = "23_34A", EXP = all_exp, P_RANGE = prange_zoom,
                              PLOT_CTCF = FALSE, PLOT_ANNOT = TRUE,
                              SPLIT_STRAND = FALSE,
                              TITLE = NULL, POINT_SIZE = 0.75, RANDOM_METHOD = "quasirandom",
                              NEXT_RANGE = NULL){
    # 1) subset once
  tib_for_plot <- TIB %>%
    filter(hopped, chr == "chr3",
           cell_line %in% CL, experiment %in% EXP,
           between(start, P_RANGE[1], P_RANGE[2]))
  
  # 2) landing‐pad join (unchanged)
  LP_tib_cl <- tib_for_plot %>%
    distinct(cell_line, launch_pad_location) %>%
    left_join(LP_tib, by = "launch_pad_location")
  
  # 3) compute counts *with* strand if requested
  if (SPLIT_STRAND) {
    count_df <- tib_for_plot %>%
      count(cell_line, strand, population)
  } else {
    count_df <- tib_for_plot %>%
      count(cell_line, population)
  }
  # 4) now build the plot
  ggplot(tib_for_plot,
         aes(x = start, y = population, col = population)) +
    
    # faceting
    {
      if (SPLIT_STRAND) {
        if (length(CL) > 1) facet_grid(cell_line ~ strand)
        else                facet_wrap(~ strand)
      } else {
        if (length(CL) > 1) facet_grid(cell_line ~ .)
        else                NULL
      }
    } +
    
    # (all your annotations go here, unchanged) … +
    {if (PLOT_ANNOT)
      list( #add elements in a list, you cannot use + inside an if statement
        # annotate enhancer
        geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown),
        annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown),
        # annotate LP
        geom_vline(xintercept = start(landingPad_8), col = "black", linetype = "dotted"), 
        geom_vline(xintercept = start(landingPad_23), col = "black", linetype = "dotted"), 
        # annotate gene
        annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf, alpha = .2, fill = 'red'),
        geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = 'red') )} +
    
     # # annotate CTCFs
    {if(PLOT_CTCF)geom_vline(xintercept = start(CTCF_mm10.chr3_extra), col = 'darkgrey', lty="11")}+
    {if(!is.null(NEXT_RANGE)) geom_rug(data = tibble(range = NEXT_RANGE),
                                       aes(x = range),
                                       fill = 'grey',
                                       inherit.aes = F)}+
    
    geom_quasirandom(alpha = 0.5, size = POINT_SIZE, varwidth = TRUE,
                     bandwidth = 0.8, method = RANDOM_METHOD, stroke = NA) +
    
    # 5) the *only* geom_text, pointing at count_df
    geom_text(
      data = count_df,
      aes(label = paste0("n = ", n), x = Inf),
      hjust = "inward", vjust = "top",
      col = 'black'
    ) +
    
    labs(x = 'Genomic coordinate (Mb)', y = 'sorted gate') +
    theme_classic(base_size = 16) +
    theme(legend.position = "none") +
    scale_x_continuous(breaks = scales::pretty_breaks(6),
                       labels = scales::unit_format(1e-6, 0.01, unit = NULL),
                       limits = P_RANGE, expand = c(0,0)) +
    scale_y_discrete(limits = rev, drop = FALSE) +
    ggtitle(TITLE) +
    colScale7_blue
}

```

### Plot Beeswarms
```{r plot beeswarms}

plot_beeswarm_only(filtered_tib_no_P1_dels %>% 
                     filter(hopped & chr == "chr3") %>% 
                     droplevels(),
                   CL = "23_34A",
                   EXP = all_exp,
                   P_RANGE = prange_zoom_paper,
                   POINT_SIZE = 1,
                   TITLE = "23_34A")


plot_beeswarm_only(filtered_tib_no_P1_dels %>% filter(hopped & chr == "chr3"),
                   CL = "Sox2P_DelD2",
                   EXP = all_exp,
                   P_RANGE = prange_zoom_paper,
                   POINT_SIZE = 1,
                   TITLE = "Sox2P_DelD2")

plot_beeswarm_only(filtered_tib_no_P1_dels %>% 
                     filter(hopped & chr == "chr3") %>% 
                     droplevels(),
                   CL = "Sox2P_DelE12",
                   EXP = all_exp,
                   P_RANGE = prange_zoom_paper,
                   POINT_SIZE = 1,
                   TITLE = "Sox2P_DelE12")
```

#Expression data
```{r expression data}

#load median expression values (from sorting pdfs)
median_sorted_mTurq = read_tsv(path_median_mTurq_relevant_cell_lines) %>% 
  select(experiment, cell_line, mCh_status, population, replicate, construct, mCherry_selection, mTurq, mCherry, GFP)

#get a general mTurq level per gate, for the mCh+ hopping
general_mTurq_mChpos = median_sorted_mTurq %>% 
  filter(str_detect(population, "^P\\d+")) %>%
  filter(mCherry_selection == "mCh+") %>%
  group_by(cell_line, population, construct, mCherry_selection) %>%
  summarise(mTurq_mean = mean(mTurq),
            mCherry_mean = mean(mCherry),
            GFP_mean = mean(GFP)) %>%
  ungroup() %>%
  select(cell_line, population, mTurq_mean, mCherry_mean, GFP_mean)

#add mTurq values to the integration tibble
tib_filtered_FACS_vals = left_join(filtered_tib_no_P1_dels, general_mTurq_mChpos, by = join_by(cell_line, population)) %>%
  filter(population != "ctrl") %>%
   dplyr::rename(mTurq_expr_gate = mTurq_mean,
                 mCh_expr_gate = mCherry_mean) 

#check that all integrations have a fluorescence value:
tib_filtered_FACS_vals %>% filter(is.na(mTurq_expr_gate))
```

#Sorting statistics
```{r sorting statistics}
sorting_stats_tib_raw = read_tsv(path_sorting_stats) 

sorting_stats_tib = sorting_stats_tib_raw %>%
  mutate(recorded_fraction = (recorded_count + 1)/total_recorded) %>% #recorded fraction with psuedocount of 1
  mutate(weigh_factor = recorded_fraction / sorted_count, 
         weigh_factor_1M = weigh_factor * 1E6) # weigh factor 1M is just to make the factor more easily human readable
```

```{r}
# number or replications per population
sorting_stats_tib_repN = sorting_stats_tib %>%
  filter(cell_line %in% c("23_34A", "Sox2P_DelD2", "Sox2P_DelE12")) %>%
  group_by(cell_line, population) %>%
  summarize(N_reps = n()) %>% #number of reps per population per cell line
  group_by(cell_line) %>%
  mutate(min_reps_CL = min(N_reps) ) %>% #least replicated population
  mutate(rep_weigh_factor = min_reps_CL/N_reps) #scale the other populations to the least replicated pop

# adjust weigh factor by the replicate weigh factor
sorting_stats_tib_repcor = 
  sorting_stats_tib %>%
  left_join(sorting_stats_tib_repN) %>%
  mutate(weigh_factor_old = weigh_factor,
         weigh_factor = weigh_factor_old * rep_weigh_factor)
```

#Expression score
## 23_34A
```{r exp score 23_34A}
#all combined, 10kb window
set.seed(20250724)
min_ints_window = 3
N_bootstrap = 1000
BinSize = 1000

tib_filtered_FACS_vals_comb = tib_filtered_FACS_vals %>% left_join(sorting_stats_tib_repcor)

#calculate running mean on random sample of same size, 1000 times
# expr_score_tib_23_34A_20kb = create_bootstrap_tibble(FUN = function(TIB){expr_score_function_new_weigh(TIB,
#                                                                                               binsize = BinSize,
#                                                                                               N_bins_window = 20,
#                                                                                               calculation_window =  prange_zoom,
#                                                                                               CL = "23_34A", EXP = all_exp)},
#                                             TIB = tib_filtered_FACS_vals_comb,
#                                             N_resamples = N_bootstrap)

#saveRDS(expr_score_tib_23_34A_20kb, "/DATA/projects/CTCF_Sox2/analysis/DelD_DelE/ME20260107_new_expression_score_23_34A_20kb.rds")
expr_score_tib_23_34A_20kb = readRDS("/DATA/projects/CTCF_Sox2/analysis/DelD_DelE/ME20260107_new_expression_score_23_34A_20kb.rds")

expr_score_sum_tib_23_34A = bootstrap_summary_fun(expr_score_tib_23_34A_20kb,
                                           score_column = "expr_score_window",
                                           conf_int = 0.95)

#calculate actual expr score (for filtering # of ints per window)
expr_score_real_23_34A =
  expr_score_function_new_weigh(tib_filtered_FACS_vals_comb,
                                binsize = BinSize,
                                N_bins_window = 20,
                                calculation_window =  prange_zoom,
                                CL = "23_34A",
                                EXP = all_exp)
  
expr_score_sum_tib_filt_comb_23_34A = left_join(expr_score_sum_tib_23_34A,  expr_score_real_23_34A) %>%
  mutate(expr_score_window_median_sel = case_when(
    total_ints_window >= min_ints_window ~ expr_score_window_median,
    .default = NA),
    expr_score_window_lower_sel = case_when(
      total_ints_window >= min_ints_window ~ expr_score_window_lower,
      .default = NA),
    expr_score_window_upper_sel = case_when(
      total_ints_window >= min_ints_window ~ expr_score_window_upper,
      .default = NA))
```

```{r plot exp score 23_34A}
pExpr_23_34A = 
ggplot(expr_score_sum_tib_filt_comb_23_34A, aes(x = mid_interval)) +
  # annotate enhancer
  list(
    annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown),
    geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown),
    # annotate gene
    annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf,alpha = .2, fill = 'red'),
    geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = 'red'),
    geom_vline(xintercept = start(landingPad_23), col = 'darkgrey'),
    geom_vline(xintercept = start(CTCF_mm10.chr3_extra), col = 'darkgrey',linetype = 'dotted')
  ) +

  ## plot the confidence interval based on the resampled data
  geom_ribbon(aes(ymin = expr_score_window_lower_sel, ymax = expr_score_window_upper_sel), alpha = 0.2, fill = "#080E5B") +
  #geom_line( aes(y = expr_score_window_mean), col = "orange") +
  ## plot the median based on the resampled data
  geom_line(aes(y = expr_score_window_median_sel), col = "#080E5B") +
 
  #axes
  scale_x_continuous(breaks = scales::pretty_breaks(n = 6), labels = scales::unit_format(scale = 1E-6, accuracy = 0.01, unit=NULL),
                     limits=prange_zoom_paper, expand=c(0,0)) +
  scale_y_continuous(expand=expansion(mult = c(0, 0.05)),
                     breaks = scales::pretty_breaks(n=6),
                     limits = c(0, NA)) +
  labs(x='Genomic coordinate (Mb)',y='expression score') +
  theme_classic(base_size = 16) +
  theme(legend.position = 'none')

pExpr_23_34A
```

## DelD
```{r expr score DelD}
#all combined, 10kb window
set.seed(20250724)
min_ints_window = 3
N_bootstrap = 1000
BinSize = 1000
#all_exp = unique(tib_filtered_FACS_vals$experiment)

tib_filtered_FACS_vals_comb = tib_filtered_FACS_vals %>% left_join(sorting_stats_tib_repcor)

#calculate running mean on random sample of same size, 5000 times
# expr_score_tib_DelD_20kb = create_bootstrap_tibble(FUN = function(TIB){expr_score_function_new_weigh(TIB,
#                                                                                               binsize = BinSize,
#                                                                                               N_bins_window = 20,
#                                                                                               calculation_window =  prange_zoom,
#                                                                                               CL = "Sox2P_DelD2", EXP = all_exp)},
#                                             TIB = tib_filtered_FACS_vals_comb,
#                                             N_resamples = N_bootstrap)

#saveRDS(expr_score_tib_DelD_20kb, "/DATA/projects/CTCF_Sox2/analysis/DelD_DelE/ME20260107_new_expression_score_23_DelD_20kb.rds")
expr_score_tib_DelD_20kb  = readRDS("/DATA/projects/CTCF_Sox2/analysis/DelD_DelE/ME20260107_new_expression_score_23_DelD_20kb.rds")


expr_score_sum_tib_23_DelD = bootstrap_summary_fun(expr_score_tib_DelD_20kb,
                                           score_column = "expr_score_window",
                                           conf_int = 0.95)

#calculate actual expr score (for filtering # of ints per window)
expr_score_real_23_DelD =
  expr_score_function_new_weigh(tib_filtered_FACS_vals_comb,
                                binsize = BinSize,
                                N_bins_window = 20,
                                calculation_window =  prange_zoom,
                                CL = "Sox2P_DelD2",
                                EXP = all_exp)
  
expr_score_sum_tib_filt_comb_DelD = left_join(expr_score_sum_tib_23_DelD,  expr_score_real_23_DelD) %>%
  mutate(expr_score_window_median_sel = case_when(
    total_ints_window >= min_ints_window ~ expr_score_window_median,
    .default = NA),
    expr_score_window_lower_sel = case_when(
      total_ints_window >= min_ints_window ~ expr_score_window_lower,
      .default = NA),
    expr_score_window_upper_sel = case_when(
      total_ints_window >= min_ints_window ~ expr_score_window_upper,
      .default = NA))
```

```{r plot expr score DelD}
pExpr_23_DelD = 
ggplot(expr_score_sum_tib_filt_comb_DelD, aes(x = mid_interval)) +
  # annotate enhancer
  list(
    annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown),
    geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown),
    # annotate gene
    annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf,alpha = .2, fill = 'red'),
    geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = 'red'),
    geom_vline(xintercept = start(landingPad_23), col = 'darkgrey'),
    geom_vline(xintercept = start(CTCF_mm10.chr3_extra), col = 'darkgrey',linetype = 'dotted')
  ) +

  ## plot the confidence interval based on the resampled data
  geom_ribbon(aes(ymin = expr_score_window_lower_sel, ymax = expr_score_window_upper_sel), alpha = 0.2, fill = "#080E5B") +
  geom_line( aes(y = expr_score_window_mean), col = "orange") +
  ## plot the median based on the resampled data
  geom_line(aes(y = expr_score_window_median_sel), col = "#080E5B") +
 
  #axes
  scale_x_continuous(breaks = scales::pretty_breaks(n = 6), labels = scales::unit_format(scale = 1E-6, accuracy = 0.01, unit=NULL),
                     limits=prange_zoom_paper, expand=c(0,0)) +
  scale_y_continuous(expand=expansion(mult = c(0, 0.05)),
                     breaks = scales::pretty_breaks(n=6),
                     limits = c(0, NA)) +
  labs(x='Genomic coordinate (Mb)',y='expression score') +
  theme_classic(base_size = 16) +
  theme(legend.position = 'none')

pExpr_23_DelD
```

## DelE
```{r expr score DelE}
#all combined, 10kb window
set.seed(20250724)
min_ints_window = 3
N_bootstrap = 1000
BinSize = 1000
#all_exp = unique(tib_filtered_FACS_vals$experiment)

tib_filtered_FACS_vals_comb = tib_filtered_FACS_vals %>% left_join(sorting_stats_tib_repcor)

#calculate running mean on random sample of same size, 5000 times
# expr_score_tib_DelE_20kb = create_bootstrap_tibble(FUN = function(TIB){expr_score_function_new_weigh(TIB,
#                                                                                               binsize = BinSize,
#                                                                                               N_bins_window = 20,
#                                                                                               calculation_window =  prange_zoom,
#                                                                                               CL = "Sox2P_DelE12", EXP = all_exp)},
#                                             TIB = tib_filtered_FACS_vals_comb,
#                                             N_resamples = N_bootstrap)
# 
# saveRDS(expr_score_tib_DelE_20kb, "/DATA/projects/CTCF_Sox2/analysis/DelD_DelE/ME20260107_new_expression_score_23_DelE_20kb.rds")
expr_score_tib_DelE_20kb  = readRDS("/DATA/projects/CTCF_Sox2/analysis/DelD_DelE/ME20260107_new_expression_score_23_DelE_20kb.rds")

expr_score_sum_tib_23_DelE = bootstrap_summary_fun(expr_score_tib_DelE_20kb,
                                           score_column = "expr_score_window",
                                           conf_int = 0.95)

#calculate actual expr score (for filtering # of ints per window)
expr_score_real_23_DelE =
  expr_score_function_new_weigh(tib_filtered_FACS_vals_comb,
                                binsize = BinSize,
                                N_bins_window = 20,
                                calculation_window =  prange_zoom,
                                CL = "Sox2P_DelE12",
                                EXP = all_exp)
  
expr_score_sum_tib_filt_comb_DelE = left_join(expr_score_sum_tib_23_DelE,  expr_score_real_23_DelE) %>%
  mutate(expr_score_window_median_sel = case_when(
    total_ints_window >= min_ints_window ~ expr_score_window_median,
    .default = NA),
    expr_score_window_lower_sel = case_when(
      total_ints_window >= min_ints_window ~ expr_score_window_lower,
      .default = NA),
    expr_score_window_upper_sel = case_when(
      total_ints_window >= min_ints_window ~ expr_score_window_upper,
      .default = NA))
```

```{r plot expr score DelE}
pExpr_23_DelE = 
ggplot(expr_score_sum_tib_filt_comb_DelE, aes(x = mid_interval)) +
  # annotate enhancer
  list(
    annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown),
    geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown),
    # annotate gene
    annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf,alpha = .2, fill = 'red'),
    geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = 'red'),
    geom_vline(xintercept = start(landingPad_23), col = 'darkgrey'),
    geom_vline(xintercept = start(CTCF_mm10.chr3_extra), col = 'darkgrey',linetype = 'dotted')
  ) +

  ## plot the confidence interval based on the resampled data
  geom_ribbon(aes(ymin = expr_score_window_lower_sel, ymax = expr_score_window_upper_sel), alpha = 0.2, fill = "#080E5B") +
  geom_line( aes(y = expr_score_window_mean), col = "orange") +
  ## plot the median based on the resampled data
  geom_line(aes(y = expr_score_window_median_sel), col = "#080E5B") +
 
  #axes
  scale_x_continuous(breaks = scales::pretty_breaks(n = 6), labels = scales::unit_format(scale = 1E-6, accuracy = 0.01, unit=NULL),
                     limits=prange_zoom_paper, expand=c(0,0)) +
  scale_y_continuous(expand=expansion(mult = c(0, 0.05)),
                     breaks = scales::pretty_breaks(n=6),
                     limits = c(0, NA)) +
  labs(x='Genomic coordinate (Mb)',y='expression score') +
  theme_classic(base_size = 16) +
  theme(legend.position = 'none')

pExpr_23_DelE
```

# Plot combined
## facet by cell line
```{r plot expr score combined facet CL}

expr_score_sum_per_cl = bind_rows(`23_34A` = expr_score_sum_tib_filt_comb_23_34A, 
                                  `Sox2P_DelD2` = expr_score_sum_tib_filt_comb_DelD, 
                                  `Sox2P_DelE12` = expr_score_sum_tib_filt_comb_DelE,
                                  .id = 'cell_line') %>%
  mutate(cell_line = factor(cell_line, levels = c("23_34A", "Sox2P_DelD2", "Sox2P_DelE12")))


# Plot 
expr_score_sum_per_cl %>%
  ggplot(aes(x = mid_interval)) +
  facet_grid(cell_line ~ .) +
  # annotate enhancer
  list(
    annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown),
    geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown),
    # annotate gene
    annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf,alpha = .2, fill = 'red'),
    geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = 'red')
    #geom_vline(data = LP_tib_for_plot, aes(xintercept = start), col = 'darkgrey'),
    
  ) +
  
  ## plot the confidence interval based on the resampled data
  geom_ribbon(aes(ymin = expr_score_window_lower_sel, ymax = expr_score_window_upper_sel, fill = cell_line), alpha = 0.2) +
  # geom_line( aes(y = expr_score_window_mean), col = 'orange') +
  ## plot the median based on the resampled data
  geom_line(aes(y = expr_score_window_median_sel, col = cell_line), linewidth = 0.6) +
  
  #axes
  scale_x_continuous(breaks = scales::pretty_breaks(n = 6), labels = scales::unit_format(scale = 1E-6, accuracy = 0.01, unit=NULL),
                     limits=prange_zoom_paper, expand=c(0,0)) +
  scale_y_continuous(expand=expansion(mult = c(0, 0.05)),
                     breaks = scales::pretty_breaks(n=6),
                     limits = c(0, NA)) +
  labs(x='Genomic coordinate (Mb)',y='expression score') +
  theme_classic(base_size = 16) +
  theme(legend.position = 'bottom') +
  colScaleCell_line +
  fillScaleCell_line
```

## Combine DelD with Sox2P control for each plot
```{r}
# DelD and Sox2P
expr_score_sum_per_DelD_Sox2P = bind_rows(`23_34A` = expr_score_sum_tib_filt_comb_23_34A, 
                                  `Sox2P_DelD2` = expr_score_sum_tib_filt_comb_DelD,
                                  .id = 'cell_line') %>%
  mutate(cell_line = factor(cell_line, levels = c("23_34A", "Sox2P_DelD2")))

# Plot together
expr_score_sum_per_DelD_Sox2P %>%
  ggplot(aes(x = mid_interval)) +
  #facet_grid(. ~ cell_line) +
  # annotate enhancer
  list(
    annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown),
    geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown),
    # annotate gene
    annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf,alpha = .2, fill = 'red'),
    geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = 'red'),
    geom_vline(xintercept = start(landingPad_23), aes(xintercept = start), col = 'darkgrey'),
    geom_vline(xintercept = start(CTCF_mm10.chr3_extra), col = 'darkgrey',linetype = 'dotted')
  ) +
  
  ## plot the confidence interval based on the resampled data
  geom_ribbon(aes(ymin = expr_score_window_lower_sel, ymax = expr_score_window_upper_sel, fill = cell_line), alpha = 0.1) +
  # geom_line( aes(y = expr_score_window_mean), col = 'orange') +
  ## plot the median based on the resampled data
  geom_line(aes(y = expr_score_window_median_sel, col = cell_line), linewidth = 0.6) +
  
  #axes
  scale_x_continuous(breaks = scales::pretty_breaks(n = 6), labels = scales::unit_format(scale = 1E-6, accuracy = 0.01, unit=NULL),
                     limits=prange_zoom_paper, expand=c(0,0)) +
  scale_y_continuous(expand=expansion(mult = c(0, 0.05)),
                     breaks = scales::pretty_breaks(n=6),
                     limits = c(0, NA)) +
  labs(x='Genomic coordinate (Mb)',y='expression score') +
  theme_classic(base_size = 16) +
  theme(legend.position = 'bottom') +
  colScaleCell_line +
  fillScaleCell_line +
  ggtitle("20kb resolution")
```

## Combine DelE with Sox2P control for each plot
```{r}
# DelE and Sox2P
expr_score_sum_per_DelE_Sox2P = bind_rows(`23_34A` = expr_score_sum_tib_filt_comb_23_34A, 
                                  `Sox2P_DelE12` = expr_score_sum_tib_filt_comb_DelE,
                                  .id = 'cell_line') %>%
  mutate(cell_line = factor(cell_line, levels = c("23_34A", "Sox2P_DelE12")))


# Plot together
expr_score_sum_per_DelE_Sox2P %>%
  ggplot(aes(x = mid_interval)) +
  #facet_grid(. ~ cell_line) +
  # annotate enhancer
  list(
    annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown),
    geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown),
    # annotate gene
    annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf,alpha = .2, fill = 'red'),
    geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = 'red'),
    geom_vline(xintercept = start(landingPad_23), aes(xintercept = start), col = 'darkgrey'),
    geom_vline(xintercept = start(CTCF_mm10.chr3_extra), col = 'darkgrey',linetype = 'dotted')
  ) +
  
  ## plot the confidence interval based on the resampled data
  geom_ribbon(aes(ymin = expr_score_window_lower_sel, ymax = expr_score_window_upper_sel, fill = cell_line), alpha = 0.1) +
  # geom_line( aes(y = expr_score_window_mean), col = 'orange') +
  ## plot the median based on the resampled data
  geom_line(aes(y = expr_score_window_median_sel, col = cell_line), linewidth = 0.6) +
  
  #axes
  scale_x_continuous(breaks = scales::pretty_breaks(n = 6), labels = scales::unit_format(scale = 1E-6, accuracy = 0.01, unit=NULL),
                     limits=prange_zoom_paper, expand=c(0,0)) +
  scale_y_continuous(expand=expansion(mult = c(0, 0.05)),
                     breaks = scales::pretty_breaks(n=6),
                     limits = c(0, NA)) +
  labs(x='Genomic coordinate (Mb)',y='expression score') +
  theme_classic(base_size = 16) +
  theme(legend.position = 'bottom') +
  colScaleCell_line +
  fillScaleCell_line +
  ggtitle("20kb resolution")

```


#Supplementary figures
## Correlation of replicates
-correlate DelD, DelE and Sox2P-WT
### 23_34A
```{r Correlation of replicates for 23_34A}
bin_borders = seq(from = prange_zoom_paper[1], to = prange_zoom_paper[2], by = 5000)
bin_borders_tib = tibble(start_window = bin_borders) %>%
  mutate(interval_number = row_number())

#count the number of integrations per cell line, population and bin
N_ints_pop_tib = filtered_tib_no_P1_dels %>% 
  filter(cell_line == "23_34A" & hopped) %>%
  group_by(experiment, population) %>%
  summarize(N_ints_pop = n())

N_ints_bins_tmp = filtered_tib_no_P1_dels %>% 
  filter(chr == "chr3" & cell_line == "23_34A" & hopped) %>%
  mutate(interval_number = findInterval(start, bin_borders)) %>%
  group_by(experiment, population, interval_number) %>%
  summarize(N_ints_bin = n()) %>%
  left_join(N_ints_pop_tib) %>%
  mutate(fract_ints_bin = N_ints_bin/N_ints_pop)

frac_ints_bins_wide = N_ints_bins_tmp %>%
  pivot_wider(id_cols = c(population, interval_number),
              names_from = experiment,
              values_from = fract_ints_bin,
              values_fill = 0)

frac_ints_bins_wide_full = 
  #make all combinations of interval_number and population
  expand_grid(interval_number = c(0, bin_borders_tib$interval_number),
              population = unique(N_ints_bins_tmp$population)) %>%
  #add the fractions and replace all NAs with zeros
  left_join(frac_ints_bins_wide) %>%
  mutate(across(everything(), ~replace_na(.x, 0))) 

#plot pairs
filter(frac_ints_bins_wide_full, 
               population != 'ctrl' & 
               !interval_number %in% c(0, max(interval_number))) %>% #remove the outer bins, which are bigger than 5kb
  dplyr::rename(replicate_1 = E2138,
                replicate_2 = E2282,
                replicate_3 = E2285,
                replicate_4 = E2712) %>%
  mutate(replicate_4 = case_when(population %in% c("P1", "P2", "P6") ~ NA, #replicate 4 only contained P3-P5
                                 .default = replicate_4)) %>%
  ggpairs(aes(col = population),
        columns = paste0('replicate_',1:4),
        lower = list(continuous = "points", stroke = NA), #, mapping=aes(color=population
        upper = list(continuous = wrap("cor", method = "spearman"))) +
  theme_bw(base_size = 12) 
```

### DelD
```{r Correlation of replicates for DelD}

#count the number of integrations per cell line, population and bin
N_ints_pop_tib = filtered_tib_no_P1_dels %>% 
  filter(cell_line == "Sox2P_DelD2" & hopped) %>%
  group_by(experiment, population) %>%
  summarize(N_ints_pop = n())

N_ints_bins_tmp = filtered_tib_no_P1_dels %>% 
  filter(chr == "chr3" & cell_line == "Sox2P_DelD2" & hopped) %>%
  mutate(interval_number = findInterval(start, bin_borders)) %>%
  group_by(experiment, population, interval_number) %>%
  summarize(N_ints_bin = n()) %>%
  left_join(N_ints_pop_tib) %>%
  mutate(fract_ints_bin = N_ints_bin/N_ints_pop)

frac_ints_bins_wide = N_ints_bins_tmp %>%
  pivot_wider(id_cols = c(population, interval_number),
              names_from = experiment,
              values_from = fract_ints_bin,
              values_fill = 0)

frac_ints_bins_wide_full = 
  #make all combinations of interval_number and population
  expand_grid(interval_number = c(0, bin_borders_tib$interval_number),
              population = unique(N_ints_bins_tmp$population)) %>%
  #add the fractions and replace all NAs with zeros
  left_join(frac_ints_bins_wide) %>%
  mutate(across(everything(), ~replace_na(.x, 0))) 

#plot pairs
filter(frac_ints_bins_wide_full, 
               population != 'ctrl' & 
              !interval_number %in% c(0, max(interval_number))) %>% #remove the outer bins, which are bigger than 5kb
  dplyr::rename(replicate_1 = E2282,
                replicate_2 = E2347) %>%
  ggpairs(aes(col = population),
        columns = paste0('replicate_',1:2),
        lower = list(continuous = "points", stroke = NA), #, mapping=aes(color=population
        upper = list(continuous = wrap("cor", method = "spearman"))) +
  theme_bw(base_size = 12) 
```

### DelE
```{r Correlation of replicates for DelE}

#count the number of integrations per cell line, population and bin
N_ints_pop_tib = filtered_tib_no_P1_dels %>% 
  filter(cell_line == "Sox2P_DelE12" & hopped) %>%
  group_by(experiment, population) %>%
  summarize(N_ints_pop = n())

N_ints_bins_tmp = filtered_tib_no_P1_dels %>% 
  filter(chr == "chr3" & cell_line == "Sox2P_DelE12" & hopped) %>%
  mutate(interval_number = findInterval(start, bin_borders)) %>%
  group_by(experiment, population, interval_number) %>%
  summarize(N_ints_bin = n()) %>%
  left_join(N_ints_pop_tib) %>%
  mutate(fract_ints_bin = N_ints_bin/N_ints_pop)

frac_ints_bins_wide = N_ints_bins_tmp %>%
  pivot_wider(id_cols = c(population, interval_number),
              names_from = experiment,
              values_from = fract_ints_bin,
              values_fill = 0)

frac_ints_bins_wide_full = 
  #make all combinations of interval_number and population
  expand_grid(interval_number = c(0, bin_borders_tib$interval_number),
              population = unique(N_ints_bins_tmp$population)) %>%
  #add the fractions and replace all NAs with zeros
  left_join(frac_ints_bins_wide) %>%
  mutate(across(everything(), ~replace_na(.x, 0))) 

#plot pairs
filter(frac_ints_bins_wide_full, 
               population != 'ctrl' & 
              !interval_number %in% c(0, max(interval_number))) %>% #remove the outer bins, which are bigger than 5kb
  dplyr::rename(replicate_1 = E2282,
                replicate_2 = E2347) %>%
  ggpairs(aes(col = population),
        columns = paste0('replicate_',1:2),
        lower = list(continuous = "points", stroke = NA), #, mapping=aes(color=population
        upper = list(continuous = wrap("cor", method = "spearman"))) +
  theme_bw(base_size = 12) 
```

## Beeswarm of both replicates
```{r DelD replicate Beeswarm}
pBS_DelD_rep1 = 
plot_beeswarm_only(filtered_tib_no_P1_dels %>% 
                     filter(hopped & chr == "chr3") %>% 
                     droplevels(),
                   CL = "Sox2P_DelD2",
                   EXP = "E2282",
                   P_RANGE = prange_zoom_paper,
                   POINT_SIZE = 1,
                   TITLE = "Sox2P_DelD2_rep1")

pBS_DelD_rep2 = 
plot_beeswarm_only(filtered_tib_no_P1_dels %>% 
                     filter(hopped & chr == "chr3") %>% 
                     droplevels(),
                   CL = "Sox2P_DelD2",
                   EXP = "E2347",
                   P_RANGE = prange_zoom_paper,
                   POINT_SIZE = 1,
                   TITLE = "Sox2P_DelD2_rep2")

plot_grid(pBS_DelD_rep1, pBS_DelD_rep2, cols = 1, align = "hv")
```

## Beeswarm of both replicates
```{r beeswarm DelE combined}
pBS_DelE_rep1 = 
plot_beeswarm_only(filtered_tib_no_P1_dels %>% 
                     filter(hopped & chr == "chr3") %>% 
                     droplevels(),
                   CL = "Sox2P_DelE12",
                   EXP = "E2282",
                   P_RANGE = prange_zoom_paper,
                   POINT_SIZE = 1,
                   TITLE = "Sox2P_DelE12_rep1")

pBS_DelE_rep2 = 
plot_beeswarm_only(filtered_tib_no_P1_dels %>% 
                     filter(hopped & chr == "chr3") %>% 
                     droplevels(),
                   CL = "Sox2P_DelE12",
                   EXP = "E2347",
                   P_RANGE = prange_zoom_paper,
                   POINT_SIZE = 1,
                   TITLE = "Sox2P_DelE12_rep2")

plot_grid(pBS_DelE_rep1, pBS_DelE_rep2, cols = 1, align = "hv")
```

## Beeswarm of 23_34A
```{r}
pBS_23_34A_rep1 = 
plot_beeswarm_only(filtered_tib_no_P1_dels %>% 
                     filter(hopped & chr == "chr3") %>% 
                     droplevels(),
                   CL = "23_34A",
                   EXP = "E2138",
                   P_RANGE = prange_zoom_paper,
                   POINT_SIZE = 1,
                   TITLE = "23_34A_rep1")

pBS_23_34A_rep2 = 
plot_beeswarm_only(filtered_tib_no_P1_dels %>% 
                     filter(hopped & chr == "chr3") %>% 
                     droplevels(),
                   CL = "23_34A",
                   EXP = "E2282",
                   P_RANGE = prange_zoom_paper,
                   POINT_SIZE = 1,
                   TITLE = "23_34A_rep2")

pBS_23_34A_rep3 = 
plot_beeswarm_only(filtered_tib_no_P1_dels %>% 
                     filter(hopped & chr == "chr3") %>% 
                     droplevels(),
                   CL = "23_34A",
                   EXP = "E2285",
                   P_RANGE = prange_zoom_paper,
                   POINT_SIZE = 1,
                   TITLE = "23_34A_rep3")

pBS_23_34A_rep4 = 
plot_beeswarm_only(filtered_tib_no_P1_dels %>% 
                     filter(hopped & chr == "chr3"),
                   CL = "23_34A",
                   EXP = c("E2712", "E2716"),
                   P_RANGE = prange_zoom_paper,
                   POINT_SIZE = 1,
                   TITLE = "23_34A_rep4")


plot_grid(pBS_23_34A_rep1, pBS_23_34A_rep2, pBS_23_34A_rep3, pBS_23_34A_rep4, cols = 2, rows = 2, align = "hv")
```

