---
title: "CTCF_only_enrichment"
author: "Mathias Eder"
date: "2026-03-02"
output: 
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---

# Nomenclature
ctrl --> non sorted control population
mCh_mid --> sorted for a decrease in Sox2::mCherry expression ('mCh_decr')
mCh_high --> sorted for an increase in Sox2::mCherry expression ('mCh_incr')
mCh_low --> Sox2::mCherry expression is absent due to hopping induced deletions (off), not used for this study

# Analysis
Calculate enrichment of sorted populations over the non-sorted controls, bulk and motif orientation based. 
Plot correlations between replicates.
Plot distribution of integrations. 
Investigate if there are statistically significant differences between locations of integrations


# Aim
Here i want to plot the CTCF integrations. 
Specifically the enrichment over control and enrichment of Fw vs Rv strand. 

```{r load_libraries}

# Load dependencies
library(GenomicRanges)
library(rtracklayer)
library(tidyverse)
# library(gtools)
library(tidyverse)
# library(dtplyr)
library(dplyr)
# library(tidyr)
library(ggbeeswarm)
library(ggplot2)
library(Biostrings)
library(stringr)
library(readr)
library(knitr)
library(RColorBrewer)
library(cowplot)
library(RcppRoll)
library(viridis)
library(ggh4x)
library(ggrastr)
library(caTools)
library(readxl)
import::from(flowCore, .except=c("filter")) #I don't load the whole flowCore and ggcyto libraries, because they overwrite dplyr's filter function :0
import::from(ggcyto, 'fortify_fs')
library(GGally)
library(GENOVA)
library(ggpubr)
library(ggridges)
library(ggpmisc)
library(Signac)
library(corrplot)
library(ggnewscale)
library(ggsignif)

split_string <- function(vect,sep,N1,N2=N1){
  library(stringr)
  sapply(vect, function(X){
    paste(str_split(X,sep)[[1]][N1:N2],collapse = sep)
  },USE.NAMES = F)
}

```

# Paths & parameters
```{r paths_parameters}
setwd("/DATA/usr/m.eder/GitHub/CBS_Sox2_manuscript/Figures/Figure_5/")

#datatag
datetag = paste0("ME",format(Sys.time(), '%Y%m%d'))

# Prepare output 
output_dir <- paste0("/DATA/usr/m.eder/GitHub/CBS_Sox2_manuscript/Figures/Figure_5/",datetag)
dir.create(output_dir, showWarnings = FALSE)
```

```{r knits setup}

library(knitr)
opts_chunk$set(cache = T,
               message = F, 
               warning = F,
               dev=c('png', 'pdf'), 
               dpi = 600,
               fig.path = file.path(output_dir, "figures/"),
               cache.lazy = F) 
pdf.options(useDingbats = FALSE)
```

## File paths
```{r}
path_CTCF_sites = "/DATA/usr/v.franceschini/Workspaces/2024_01_MATHIAS_CTCF_PEAKS/VF240812_calling_CTCF_motifs_again/02_OUTPUTS/CTCF_sites/6764_2_ME_E2_CGATGT_S2_R1_001_peaks_motifs.bed"

# New path
path_CTCF_ints = "/DATA/usr/m.eder/GitHub/CBS_Sox2_manuscript/Tagmentation/Integration_tables/mapped_integrations_-116kb_CBS_pools.txt"

# Sorting stats
path_sorting_stats = "/DATA/projects/CTCF_Sox2/Info_files/me20250530_stats_for_expression_score_mCherry.xlsx"

# Expression score function
source("/DATA/projects/Sox2/General_functions/CM20250318_new_expression_score_functions.R")
```


# Load and filter CTCF hopping data
```{r load CTCF data}

# load CTCF data
tib_CTCF = read_tsv(path_CTCF_ints) %>%
  filter(chr %in% paste0("chr", c(1:19, "X"))) %>%
  mutate(chr = factor(chr, levels = paste0("chr", c(1:19, "X")))) %>%
  group_by(chr, start, end, seq, strand, replicate, experiment, launch_pad_location, insert, genotype, sample_type, sorted_gate, pool_nr, sample, cell_line, CBS_motif_orientation) %>%
  summarize(read_count = sum(read_count),
            read_count_1 = sum(read_count_1),
            read_count_2 = sum(read_count_2), .groups = "keep") %>%
  #mark hopped integrations
  mutate(hopped = start != 34643961)

#filter the integrations
tib_filtered_CTCF <- tib_CTCF %>%
  #add variables used in the plotting (old names of cell lines etc) 
  mutate(population = case_when(sample_type == "control pool" ~ "ctrl",
                                .default = sorted_gate),
         #annotate confidence of mapping (one or both ITRs)
         mapped_arms = case_when(read_count_1 == 0 ~ "rv_only",
                                 read_count_2 == 0 ~ "fw_only",
                                 read_count_1 > 0 & read_count_2 > 0 ~ "both_arms"),
         landing_pad = case_when(launch_pad_location == "-116 kb" ~ "23"),
         motif = case_when(CBS_motif_orientation == "FW" ~ "forward",
                           CBS_motif_orientation == "RV" ~ "reverse",
                           CBS_motif_orientation == "-" & strand == "+" ~ "forward",
                           CBS_motif_orientation == "-" & strand == "-" ~ "reverse",
                           .default = CBS_motif_orientation)) %>%
  #filter the data
  filter(population == "ctrl" | read_count >= 2) %>% #for ctrl distribution 1 mapping read is enough
  filter(!(start >= 34721183 & start <= 34721192)) %>% #remove contamination from LP8
  filter(strand %in% c("+", "-")) %>% #remove non annotated strand
  filter(!(population == "mCh_low" & mapped_arms != "both_arms")) %>% #only keep mCh_low integrations with 2 mapped arms
  filter(!(experiment == "E1665" & population == "mCh_mid" & mapped_arms != "both_arms")) %>% #mCh_close is the population of interest in E1665, mid is a bit lower and needs to be filtered for not having deletions
  mutate(population = case_when(experiment == "E1665" & population == "mCh_close" ~ "mCh_mid",
                                .default = population))

# quick check
table(tib_filtered_CTCF$cell_line, tib_filtered_CTCF$population)
table(tib_filtered_CTCF$population, tib_filtered_CTCF$chr)

# Combine the 1xCBS cell lines (CBS16A + 16A)
tib_CBS16_16A_comb = tib_filtered_CTCF %>% filter(population %in% c("ctrl", "mCh_mid", "mCh_high")) %>% 
  mutate(cell_line = case_when(cell_line == "CBS16" ~ "CBS16_16A",
                               cell_line == "16A" ~ "CBS16_16A",
                               .default = cell_line))
 
all_exp = unique(tib_filtered_CTCF$experiment)
```

## Setup
```{r setup}
path_CTCF_sites = "/DATA/usr/v.franceschini/Workspaces/2024_01_MATHIAS_CTCF_PEAKS/VF240812_calling_CTCF_motifs_again/02_OUTPUTS/CTCF_sites/6764_2_ME_E2_CGATGT_S2_R1_001_peaks_motifs.bed"


# Location of enhancer / gene
enhancer <- GRanges(seqnames = "chr3", 
                    ranges = IRanges(start = 34753415,
                                     end = 34766401),
                    strand = "*")
Sox2_gene <- GRanges(seqnames = "chr3", 
                    ranges = IRanges(start = 34649995,
                                     end = 34652461),
                    strand = "+")

landingPad_23 <- GRanges(seqnames = "chr3", 
                    ranges = IRanges(start = 34643960,
                                     end = 34643962),
                    strand = "+")


landingPad_8 <- GRanges(seqnames = "chr3", 
                    ranges = IRanges(start = 34721183,
                                     end = 34721184),
                    strand = "+")


landingPad_C9 <- GRanges(seqnames = "chr3", 
                    ranges = IRanges(start = 34598479,
                                     end = 34598480),
                    strand = "+")
landingPad_C10 <- GRanges(seqnames = "chr3", 
                    ranges = IRanges(start = 35478413,
                                     end = 35478414      ),
                    strand = "+")

LP_tib = tibble(landing_pad = c("23", "C9", "C10"), 
                start = c( start(landingPad_23),start(landingPad_C9), start(landingPad_C10)))


#CTCF sites based on our own mapping
# new analysis (based on mapping to mm10 only)
CTCF_mm10_new <- import.bed(path_CTCF_sites)

#filtering:
CTCF_mm10.chr3 <- CTCF_mm10_new[seqnames(CTCF_mm10_new)== 'chr3']
# add the missing site after the SCR
CTCF_mm10.chr3_extra = sort(c(CTCF_mm10.chr3,
                         GRanges(seqnames = "chr3",
                                 IRanges(start = 34772210, end = 34772210),
                                 strand = "+")),
                         ignore.strand = T)


prange_zoom = c(34.5E6, 35E6)
prange_zoom_paper = c(34.59E6, 34.83E6)
prange_SCR = c(34.725E6, 34.785E6)
prange_2Mb = c(33.643960E6, 35.643960E6)
prange_300kb = c(34549995, 34866401)

# Color annotations
## SCR 
brown = "#ffb000"


LP_tib = tibble(landing_pad = c("23", "C9", "C10"), 
                start = c( start(landingPad_23),start(landingPad_C9), start(landingPad_C10)))
```

## Filter out integrtations that are within the Sox2::mCherry gene
```{r}
# How many integrations are in the Sox2 gene?
tib_int_gene = tib_CBS16_16A_comb %>% 
  filter(start >= start(Sox2_gene) & start <= end(Sox2_gene)) %>% 
  group_by(cell_line, population) %>% 
  summarise(count = n())

tib_int_gene

#Filter out integrations in the gene
tib_filt = tib_CBS16_16A_comb %>% 
  filter(!(start >= start(Sox2_gene) & start <= end(Sox2_gene)))

#Add levels to cell lines for plotting order
tib_filt = tib_filt %>% 
  mutate(cell_line = factor(cell_line, levels = c("37A", "CRE6", "CBS16_16A", "3xCTCF")))

```

# Correlation of integration denisty distributions
- bin region into 5kb bins
- calculate fraction (bin/total)
- correlate cell lines with each other

## Correlation between replicates
### Figure S4D - CRE6
```{r Fig S4D - Correlation of replicates for CRE6}
bin_borders = seq(from = prange_2Mb[1], to = prange_2Mb[2], by = 20000)
bin_borders_tib = tibble(start_window = bin_borders) %>%
  mutate(interval_number = row_number())

#count the number of integrations per cell line, population and bin
N_ints_pop_tib = tib_filtered_CTCF  %>% 
  filter(cell_line == "CRE6" & hopped & population %in% c("mCh_mid")) %>%
  group_by(replicate, population) %>%
  summarize(N_ints_pop = n())

N_ints_bins_tmp = tib_filtered_CTCF  %>% 
  filter(chr == "chr3" & cell_line == "CRE6" & hopped & population %in% c("mCh_mid")) %>%
  mutate(interval_number = findInterval(start, bin_borders)) %>%
  group_by(replicate, population, interval_number) %>%
  summarize(N_ints_bin = n()) %>%
  left_join(N_ints_pop_tib) %>%
  mutate(fract_ints_bin = N_ints_bin/N_ints_pop)

frac_ints_bins_wide = N_ints_bins_tmp %>%
  pivot_wider(id_cols = c(population, interval_number),
              names_from = replicate,
              values_from = fract_ints_bin,
              values_fill = 0)

frac_ints_bins_wide_full = 
  #make all combinations of interval_number and population
  expand_grid(interval_number = c(0, bin_borders_tib$interval_number),
              population = unique(N_ints_bins_tmp$population)) %>%
  #add the fractions and replace all NAs with zeros
  left_join(frac_ints_bins_wide) %>%
  mutate(across(everything(), ~replace_na(.x, 0))) 

pCRE_cor_rep = 
frac_ints_bins_wide_full %>%
ggpairs(aes(col = population),
        columns = paste0('rep',1:3),
        lower = list(continuous = "points", stroke = NA), #, mapping=aes(color=population
        upper = list(continuous = wrap("cor", method = "spearman"))) +
  ggtitle("CRE6") +
  theme_bw(base_size = 12) 

pCRE_cor_rep
```

### Figure S4E - CBS16
```{r Fig S4E - Correlation of replicates for CBS16}
bin_borders = seq(from = prange_2Mb[1], to = prange_2Mb[2], by = 20000)
bin_borders_tib = tibble(start_window = bin_borders) %>%
  mutate(interval_number = row_number())

#count the number of integrations per cell line, population and bin
N_ints_pop_tib = tib_filtered_CTCF  %>% 
  filter(cell_line == "CBS16" & hopped & population %in% c("mCh_mid")) %>%
  group_by(replicate, population) %>%
  summarize(N_ints_pop = n())

N_ints_bins_tmp = tib_filtered_CTCF  %>% 
  filter(chr == "chr3" & cell_line == "CBS16" & hopped & population %in% c("mCh_mid")) %>%
  mutate(interval_number = findInterval(start, bin_borders)) %>%
  group_by(replicate, population, interval_number) %>%
  summarize(N_ints_bin = n()) %>%
  left_join(N_ints_pop_tib) %>%
  mutate(fract_ints_bin = N_ints_bin/N_ints_pop)

frac_ints_bins_wide = N_ints_bins_tmp %>%
  pivot_wider(id_cols = c(population, interval_number),
              names_from = replicate,
              values_from = fract_ints_bin,
              values_fill = 0)

frac_ints_bins_wide_full = 
  #make all combinations of interval_number and population
  expand_grid(interval_number = c(0, bin_borders_tib$interval_number),
              population = unique(N_ints_bins_tmp$population)) %>%
  #add the fractions and replace all NAs with zeros
  left_join(frac_ints_bins_wide) %>%
  mutate(across(everything(), ~replace_na(.x, 0))) 

pCBS16_cor_rep =
frac_ints_bins_wide_full %>%
ggpairs(aes(col = population),
        columns = paste0('rep',1:3),
        lower = list(continuous = "points", stroke = NA), #, mapping=aes(color=population
        upper = list(continuous = wrap("cor", method = "spearman"))) +
  ggtitle("CBS16") +
  theme_bw(base_size = 12) 

pCBS16_cor_rep
```

### Figure S4F - 16A
```{r Fig S4F -Correlation of replicates for 16A}
bin_borders = seq(from = prange_2Mb[1], to = prange_2Mb[2], by = 20000)
bin_borders_tib = tibble(start_window = bin_borders) %>%
  mutate(interval_number = row_number())

#count the number of integrations per cell line, population and bin
N_ints_pop_tib = tib_filtered_CTCF  %>% 
  filter(cell_line == "16A" & hopped & population %in% c("mCh_mid")) %>%
  group_by(replicate, population) %>%
  summarize(N_ints_pop = n())

N_ints_bins_tmp = tib_filtered_CTCF  %>% 
  filter(chr == "chr3" & cell_line == "16A" & hopped & population %in% c("mCh_mid")) %>%
  mutate(interval_number = findInterval(start, bin_borders)) %>%
  group_by(replicate, population, interval_number) %>%
  summarize(N_ints_bin = n()) %>%
  left_join(N_ints_pop_tib) %>%
  mutate(fract_ints_bin = N_ints_bin/N_ints_pop)

frac_ints_bins_wide = N_ints_bins_tmp %>%
  pivot_wider(id_cols = c(population, interval_number),
              names_from = replicate,
              values_from = fract_ints_bin,
              values_fill = 0)

frac_ints_bins_wide_full = 
  #make all combinations of interval_number and population
  expand_grid(interval_number = c(0, bin_borders_tib$interval_number),
              population = unique(N_ints_bins_tmp$population)) %>%
  #add the fractions and replace all NAs with zeros
  left_join(frac_ints_bins_wide) %>%
  mutate(across(everything(), ~replace_na(.x, 0))) 

p16A_cor_rep =
frac_ints_bins_wide_full %>%
ggpairs(aes(col = population),
        columns = paste0('rep',1:2),
        lower = list(continuous = "points", stroke = NA), #, mapping=aes(color=population
        upper = list(continuous = wrap("cor", method = "spearman"))) +
  ggtitle("16A") +
  theme_bw(base_size = 12) 

p16A_cor_rep
```

### Figure S4G - 3xCBS
```{r Fig S4G -Correlation of replicates for 3xCBS}
bin_borders = seq(from = prange_2Mb[1], to = prange_2Mb[2], by = 20000)
bin_borders_tib = tibble(start_window = bin_borders) %>%
  mutate(interval_number = row_number())

#count the number of integrations per cell line, population and bin
N_ints_pop_tib = tib_filtered_CTCF  %>% 
  filter(cell_line == "3xCTCF" & hopped & population %in% c("mCh_mid")) %>%
  group_by(replicate, population) %>%
  summarize(N_ints_pop = n())

N_ints_bins_tmp = tib_filtered_CTCF  %>% 
  filter(chr == "chr3" & cell_line == "3xCTCF" & hopped & population %in% c("mCh_mid")) %>%
  mutate(interval_number = findInterval(start, bin_borders)) %>%
  group_by(replicate, population, interval_number) %>%
  summarize(N_ints_bin = n()) %>%
  left_join(N_ints_pop_tib) %>%
  mutate(fract_ints_bin = N_ints_bin/N_ints_pop)

frac_ints_bins_wide = N_ints_bins_tmp %>%
  pivot_wider(id_cols = c(population, interval_number),
              names_from = replicate,
              values_from = fract_ints_bin,
              values_fill = 0)

frac_ints_bins_wide_full = 
  #make all combinations of interval_number and population
  expand_grid(interval_number = c(0, bin_borders_tib$interval_number),
              population = unique(N_ints_bins_tmp$population)) %>%
  #add the fractions and replace all NAs with zeros
  left_join(frac_ints_bins_wide) %>%
  mutate(across(everything(), ~replace_na(.x, 0))) 

p3xCTCF_cor_rep =
frac_ints_bins_wide_full %>%
ggpairs(aes(col = population),
        columns = paste0('rep',1:2),
        lower = list(continuous = "points", stroke = NA), #, mapping=aes(color=population
        upper = list(continuous = wrap("cor", method = "spearman"))) +
  ggtitle("3xCTCF") +
  theme_bw(base_size = 12) 

p3xCTCF_cor_rep
```

### Figure S4H - Correaltion of cell lines
```{r Fig S4H - correlation of int between cell lines 5kb bins}
# Change order of cell lines
tib_filtered_CTCF = tib_filtered_CTCF %>% mutate(cell_line = factor(cell_line, levels = c("CRE6", "CBS16", "16A", "3xCTCF")))

# Calculate total integrations per cell line and ctrl
total_int = tib_filtered_CTCF %>% 
  filter(hopped & chr == "chr3" & cell_line %in% c("CBS16", "16A", "3xCTCF", "CRE6") & population == "ctrl") %>% 
  group_by(cell_line, population) %>%
  summarise(total_int = n())

binsize = 5000

#determine boundaries
boundaries_all = seq(from = prange_2Mb[1], to = prange_2Mb[2], by = binsize)
boundaries_all_tib = tibble(interval_number = 1:length(boundaries_all), start_interval = boundaries_all) %>%
    arrange(start_interval) %>%
    mutate(mid_interval = (start_interval + lead(start_interval, 1))/2)

cut_by_intervals = tib_filtered_CTCF %>%
  filter(chr == "chr3" & hopped & start >= prange_2Mb[1] & start <= prange_2Mb[2]) %>%
  arrange(start) %>%
  #per integration: in which 1-ctrl interval is it located?
  mutate(interval_number = findInterval(start, boundaries_all)) %>%
  #per interval: how many integrations are there in each population
  group_by(population, cell_line, interval_number) %>%
  summarise(N_ints = n())
  
#make a wide table; first join with the boundaries_all_tib to make sure we have all intervals in the table (also those with no ctrl ints)
cut_by_intervals_wide = left_join(boundaries_all_tib, cut_by_intervals) %>%
  mutate(N_ints = replace_na(N_ints, 0)) %>%
  arrange(interval_number)

#calculate enrichment per window
enrich_ints_by_bin_ctrl  = cut_by_intervals_wide %>%
  #long format, calculate ints_in_region/total_ints
  filter(population == "ctrl") %>%
  left_join(total_int, by = c("population", "cell_line")) %>%
  mutate(fract_of_int = N_ints / total_int)

wide_data = 
enrich_ints_by_bin_ctrl %>% dplyr::select(c(interval_number, cell_line, fract_of_int)) %>%
  pivot_wider(names_from = cell_line, values_from = fract_of_int)

wide_data <- wide_data %>%
  rename(cell_C_16A = `16A`,
         cell_B_CBS16 = CBS16,
         cell_A_CRE6 = CRE6,
         cell_D_3xCTCF = `3xCTCF`)


#check for duplicates
enrich_ints_by_bin_ctrl %>%
  filter(cell_line %in% c("cell_C_16A", "cell_B_CBS16")) %>%
  count(interval_number, cell_line) %>%
  filter(n > 1)

# Plot correlation
wide_data %>% dplyr::select("cell_C_16A", "cell_B_CBS16") %>%
ggplot(aes(x = cell_C_16A, y = cell_B_CBS16)) +
  geom_point() +
  stat_cor(method = "pearson") +
  theme_bw()

ggscatter(wide_data,
          x = "cell_C_16A", y = "cell_B_CBS16",
          add = "reg.line",               # Add regression line
          conf.int = TRUE,               # Confidence interval around the line
          cor.coef = TRUE,               # Show Pearson correlation coefficient
          cor.method = "spearman",        # Can also be "spearman" or "kendall"
          xlab = "Fraction of Int (16A)",
          ylab = "Fraction of Int (CBS16)",
          title = "Correlation of fract_of_int between 16A and CBS16", subtitle = "spearman cor")


# Look for outliers
## Trim 1st and 99th percentiles
filtered <- wide_data %>%
  filter(
    between(cell_C_16A, quantile(cell_C_16A, 0.01, na.rm = TRUE), quantile(cell_C_16A, 0.99, na.rm = TRUE)) &
    between(cell_B_CBS16, quantile(cell_B_CBS16, 0.01, na.rm = TRUE), quantile(cell_B_CBS16, 0.99, na.rm = TRUE))
  )

cor(filtered$cell_C_16A, filtered$cell_B_CBS16, use = "complete.obs")

ggscatter(filtered,
          x = "cell_C_16A", y = "cell_B_CBS16",
          add = "reg.line",               # Add regression line
          conf.int = TRUE,               # Confidence interval around the line
          cor.coef = TRUE,               # Show Pearson correlation coefficient
          cor.method = "pearson",        # Can also be "spearman" or "kendall"
          xlab = "Fraction of Int (16A)",
          ylab = "Fraction of Int (CBS16)",
          alpha = 0.5,
          size = 1.8,
          title = "Correlation of fract_of_int between 16A and CBS16", subtitle = "Trim 1st and 99th percentiles - pearson")


# pairwise correlations
## Select only the columns you want to correlate
cor_data <- wide_data %>%
  dplyr::select(cell_A_CRE6,  cell_B_CBS16, cell_C_16A, cell_D_3xCTCF)

## Compute correlation matrix (Pearson by default)
cor_matrix <- cor(cor_data, use = "pairwise.complete.obs", method = "pearson")
print(cor_matrix)

library(corrplot)
corrplot(cor_matrix, method = "number", type = "upper", tl.col = "black")


# Generate scatterplot matrix with correlation
## Pearson correlation
ggpairs(cor_data,
        lower = list(continuous = wrap("points", alpha = 0.6, size = 1.5)),
        upper = list(continuous = wrap("cor", size = 4, method = "pearson")),
        diag = list(continuous = wrap("densityDiag")),
        title = "Pairwise Scatterplots of fract_of_int by Cell Line (pearson")

## Spearman correlation
ggpairs(cor_data,
        lower = list(continuous = wrap("points", alpha = 0.6, size = 1.5)),
        upper = list(continuous = wrap("cor", size = 4, method = "spearman")),
        diag = list(continuous = wrap("densityDiag")),
        title = "Pairwise Scatterplots of fract_of_int by Cell Line (Spearman")



# trimmed data
filtered_sel = filtered %>%
  dplyr::select(cell_A_CRE6, cell_B_CBS16, cell_C_16A, cell_D_3xCTCF)

## Pearson correlation
ggpairs(filtered_sel,
        lower = list(continuous = wrap("points", alpha = 0.6, size = 1.5)),
        upper = list(continuous = wrap("cor", size = 4, method = "pearson")),
        diag = list(continuous = wrap("densityDiag")),
        title = "Pairwise Scatterplots of fract_of_int by Cell Line (pearson") +
  theme_bw()

## Spearman correlation
ggpairs(filtered_sel,
        lower = list(continuous = wrap("points", alpha = 0.6, size = 1.5)),
        upper = list(continuous = wrap("cor", size = 4, method = "spearman")),
        diag = list(continuous = wrap("densityDiag")),
        title = "Pairwise Scatterplots of fract_of_int by Cell Line (Spearman")

```
**Conclusion:**
CBS16 and 16A control integrations binned on 5kb correlate very good 0.821. CRE6 also corrlates well with CBS16 and 16A. Correlation with 3xCTCF was not as good, indicating that the normal distribution of ctrl integrations is different. 


# Beeswarm fun - non stranded
```{r beeswarm_function}

plot_beeswarm_only = function(TIB, CL = "CBS16_16A", EXP = all_exp, P_RANGE = prange_zoom,
                              PLOT_CTCF = F, PLOT_ANNOT = T,
                              TITLE = NULL, POINT_SIZE = 0.75, RANDOM_METHOD = "quasirandom",
                              NEXT_RANGE = NULL){
  tib_for_plot = filter(TIB, hopped & chr == "chr3" & cell_line %in% CL & experiment %in% EXP & start >= P_RANGE[1] & start <= P_RANGE[2])
  LP_tib_cl = dplyr::select(tib_for_plot, c(cell_line, landing_pad)) %>% 
    distinct() %>% left_join(LP_tib)
  
  ggplot(filter(tib_for_plot, hopped & chr == "chr3" & cell_line %in% CL & experiment %in% EXP),
         aes(x = start, y = population, col = population)) +
    {if(length(CL)>1) facet_grid(cell_line ~ .)} +  #only facet when necessary

    {if (PLOT_ANNOT)
      list( #add elements in a list, you cannot use + inside an if statement
        # annotate enhancer
        geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown),
        annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown),
        # annotate LP23
        geom_vline(xintercept = c(start(landingPad_23)), col = "grey"),
        # annotate gene
        annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf, alpha = .2, fill = 'red'),
        geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = 'red') )} +
    
    # annotate landing pad
{ if (PLOT_ANNOT & length(CL) == 1)
    geom_vline(
      xintercept = pull(
        filter(LP_tib,
               landing_pad %in% unique(tib_for_plot$landing_pad)),
        start
      ),
      col = "darkgrey"
    )
} +
{ if (PLOT_ANNOT & length(CL) > 1)
    NULL  # <- don't draw landing pad lines in multi-cell-line comparisons
} +


    # # annotate CTCFs
    {if(PLOT_CTCF)geom_vline(xintercept = start(CTCF_mm10.chr3_extra), col = 'darkgrey', lty="11")}+
    {if(!is.null(NEXT_RANGE)) geom_rug(data = tibble(range = NEXT_RANGE),
                                       aes(x = range),
                                       fill = 'grey',
                                       inherit.aes = F)}+
    
    
    #datapoints
    geom_quasirandom(alpha = 0.5, size = POINT_SIZE, varwidth = T, bandwidth = 0.8,
                     # shape = 16, #changes to point without border, so alpha applies to the whole point
                     method = RANDOM_METHOD,
                     stroke = NA) +
    geom_text(data = plyr::count(tib_for_plot, vars = c("cell_line", "population")), aes(label = paste("n = ", freq, sep = ""),
                                                                                         x = Inf),
              hjust = "inward", vjust = "top",
              col = 'black') +
    
    #layout:
    labs(x='Genomic coordinate (Mb)',y='sorted gate') +
    theme_classic(base_size = 16) +
    theme(legend.position = "none") +
    scale_x_continuous(breaks = scales::pretty_breaks(n = 6), labels = scales::unit_format(scale = 1E-6, accuracy = 0.01, unit=NULL),
                       limits=P_RANGE, expand=c(0,0)) +
    scale_y_discrete(limits = rev, drop=F) +
    ggtitle(TITLE)
}
```

## Figure S5A - Plot Zoom-out of all sorted gates - non stranded
```{r Fig S5A - Plot Zoom-out non stranded 2Mb range}

plot_beeswarm_only(TIB = tib_filt %>% filter(hopped & chr == "chr3" & experiment %in% all_exp),
                   CL = c("CBS16_16A", "3xCTCF", "CRE6"),
                   PLOT_ANNOT = T,
                   PLOT_CTCF = F,
                   P_RANGE = prange_2Mb,
                   POINT_SIZE = 1)

all_exp = unique(tib_filt$experiment)
```

## Plot prange_zoom_paper of all sorted gates - non stranded
```{r Plot Zoom-out non stranded}

plot_beeswarm_only(TIB = tib_filt %>% filter(hopped & chr == "chr3" & experiment %in% all_exp),
                   CL = c("CBS16_16A", "3xCTCF", "CRE6"),
                   PLOT_ANNOT = T,
                   PLOT_CTCF = F,
                   P_RANGE = prange_zoom_paper,
                   POINT_SIZE = 1)
```

# Beeswarm fun - stranded
```{r stranded beeswarm function}
plot_beeswarm_only_stranded = function(TIB, CL = "23_34A", EXP = all_exp, P_RANGE = prange_zoom,
                              PLOT_CTCF = FALSE, PLOT_ANNOT = TRUE,
                              SPLIT_MOTIF = FALSE,
                              TITLE = NULL, POINT_SIZE = 0.75, RANDOM_METHOD = "quasirandom"
                              ){
  
  # 1) subset once
  tib_for_plot <- TIB %>%
    filter(hopped, chr == "chr3",
           cell_line %in% CL, experiment %in% EXP,
           between(start, P_RANGE[1], P_RANGE[2]))
  
  # 2) landing‐pad join (unchanged)
  LP_tib_cl <- tib_for_plot %>%
    distinct(cell_line, landing_pad) %>%
    left_join(LP_tib, by = "landing_pad")
  
  # 3) compute counts *with* motif if requested
  if (SPLIT_MOTIF) {
    count_df <- tib_for_plot %>%
      count(cell_line, motif, population)
  } else {
    count_df <- tib_for_plot %>%
      count(cell_line, population)
  }
  
 
  # 4) now build the plot
  ggplot(tib_for_plot,
         aes(x = start, y = population, col = population)) +
    
    # faceting
    {
      if (SPLIT_MOTIF) {
        if (length(CL) > 1) facet_grid(cell_line ~ motif)
        else                facet_wrap(~ motif)
      } else {
        if (length(CL) > 1) facet_grid(cell_line ~ .)
        else                NULL
      }
    } +
    
    # (all your annotations go here, unchanged) … +
    {if (PLOT_ANNOT)
      list( #add elements in a list, you cannot use + inside an if statement
        # annotate enhancer
        geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown),
        annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown),
        # annotate LP
        geom_vline(xintercept = start(landingPad_8), col = "black", linetype = "dotted"), 
        geom_vline(xintercept = start(landingPad_23), col = "black", linetype = "dotted"), 
        # annotate gene
        annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf, alpha = .2, fill = 'red'),
        geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = 'red') )} +
    
    # CTCF lines (with their own color scale)
{if(PLOT_CTCF)
  geom_vline(
    data = as_tibble(CTCF_mm10.chr3_extra) %>%
      mutate(mid = (start + end)/2, strand = as.character(strand)),
    aes(xintercept = mid, color = strand),
    lty = "11",
    inherit.aes = FALSE
  )
} +
    
    geom_quasirandom(alpha = 0.5, size = POINT_SIZE, varwidth = TRUE,
                     bandwidth = 0.8, method = RANDOM_METHOD, stroke = NA) +
    
    # 5) the *only* geom_text, pointing at count_df
    geom_text(data = plyr::count(tib_for_plot, vars = c("cell_line", "population", "motif")), aes(label = paste("n = ", freq, sep = ""),
                                                                                     x = Inf),
          hjust = "inward", vjust = "top",
          col = 'black') +
    
    # 6) scales / theme / title (unchanged) …
    labs(x = 'Genomic coordinate (Mb)', y = 'sorted gate') +
    theme_bw(base_size = 16) +
    theme(legend.position = "none") +
    scale_x_continuous(breaks = scales::pretty_breaks(6),
                       labels = scales::unit_format(scale = 1E-6, accuracy = 0.01, unit=NULL),
                       limits = P_RANGE, expand = c(0,0)) +
    scale_y_discrete(limits = rev, drop = FALSE) +
    ggtitle(TITLE)
}

```

## Figure S5B - Plot beeswarm stranded
```{r stranded beeswarm}
plot_beeswarm_only_stranded(TIB = filter(tib_filt, hopped & chr == "chr3" & experiment %in% all_exp),
                   CL = c("CBS16_16A", "3xCTCF", "CRE6"),
                   SPLIT_MOTIF = T, 
                   PLOT_ANNOT = T,
                   PLOT_CTCF = T,
                   P_RANGE = prange_zoom_paper,
                   POINT_SIZE = 1.5)
```

# Enrichment function (adapted from Christines function as follows)
How does the enrichment function work
1. makes bins, counts int per bin
2. calculates fraction per bin (fraction over total)
3. summes fractions per window per population
4. divides window fraction of interest over control 

NOTE: I added a section to count integrations of separate populations per window. This should allow us afterwards to filter for windows that have integrations present. 
```{r adapted enrichment function}
# Adapted from Christines function

####################
#Below is the function to calculate enrichment over control per N kb.
# Was built for 6 gate reporter hopping, but I think this one should directly work for CTCF hopping as well.

# It takes the integrations per binsize and number of bins, so 1000 *10 = 10kb window, slided by 1kb
#####################

enrichm_function_fixed_bins_CL <- function(TIB, CL = "CBS16", EXP = all_exp, 
                                           binsize = 1000, N_bins_window = 10, 
                                           calculation_window =  c(34091354, 35744137),
                                           MOTIF = c("forward", "reverse")){
  tib_for_analysis = dplyr::filter(TIB, hopped & cell_line %in% CL & experiment %in% EXP)

  # total integrations per population (no strand filter here)
  total_ints = tib_for_analysis %>% dplyr::group_by(population) %>% dplyr::summarise(total_ints = dplyr::n(), .groups = "drop")

  # bin boundaries
  boundaries_all = seq(from = calculation_window[1], to = calculation_window[2], by = binsize)
  boundaries_all_tib = tibble::tibble(interval_number = seq_along(boundaries_all), start_interval = boundaries_all) %>%
    dplyr::arrange(start_interval) %>%
    dplyr::mutate(mid_interval = (start_interval + dplyr::lead(start_interval, 1))/2)

  # counts per bin per population
  cut_by_intervals = tib_for_analysis %>% 
    dplyr::filter(motif %in% MOTIF) %>% 
    dplyr::filter(chr == "chr3", start >= calculation_window[1], start <= calculation_window[2]) %>%
    dplyr::arrange(start) %>%
    dplyr::mutate(interval_number = findInterval(start, boundaries_all)) %>%
    dplyr::group_by(population, interval_number) %>%
    dplyr::summarise(N_ints = dplyr::n(), .groups = "drop")

  # wide: counts per bin (include empty bins)
  cut_by_intervals_wide = dplyr::left_join(boundaries_all_tib, cut_by_intervals, by = "interval_number") %>%
    dplyr::mutate(N_ints = tidyr::replace_na(N_ints, 0)) %>%
    dplyr::arrange(interval_number) %>%
    tidyr::pivot_wider(id_cols = c('start_interval','interval_number','mid_interval'),
                       names_from = population, values_from = N_ints, values_fill = 0) %>%
    dplyr::select(-tidyselect::any_of('NA'))

  ## ---- NEW PART: rolling raw counts per window -----------------------------
  # compute rolling sum of raw counts per population
  counts_roll_wide <- cut_by_intervals_wide %>%
    dplyr::mutate(dplyr::across(
      -c(start_interval, interval_number, mid_interval),
      ~ RcppRoll::roll_sum(.x, n = N_bins_window, align = "center", fill = NA_real_),
      .names = "{.col}_counts"
    ))

  # long: counts_window per population
  counts_long <- counts_roll_wide %>%
    tidyr::pivot_longer(cols = tidyselect::ends_with("_counts"),
                        names_to = "population", values_to = "counts_window") %>%
    dplyr::mutate(population = sub("_counts$", "", population)) %>%
    dplyr::select(start_interval, interval_number, mid_interval, population, counts_window)

  # grab ctrl counts per window for later filtering
  ctrl_counts <- counts_long %>%
    dplyr::filter(population == "ctrl") %>%
    dplyr::select(interval_number, ctrl_counts_window = counts_window)
  ## -------------------------------------------------------------------------

  # enrichment as before
  enrich_ints_by_bin  = cut_by_intervals_wide %>%
    tidyr::pivot_longer(cols = -c('start_interval','interval_number','mid_interval'),
                        names_to = "population", values_to = 'N_in_window') %>%
    dplyr::left_join(total_ints, by = "population") %>%
    dplyr::mutate(fract_of_ints = N_in_window / total_ints) %>%
    dplyr::select(-c('N_in_window','total_ints')) %>%
    tidyr::pivot_wider(names_from = population, values_from = fract_of_ints) %>%
    dplyr::mutate(dplyr::across(
      -c('start_interval','interval_number','mid_interval'),
      ~ RcppRoll::roll_sum(.x, n = N_bins_window, align = "center", fill = NA_real_)
    )) %>%
    dplyr::mutate(dplyr::across(
      -c('start_interval','interval_number','mid_interval'),
      ~ .x / ctrl
    )) %>%
    tidyr::pivot_longer(-c('start_interval','interval_number','mid_interval'),
                        values_to = 'int_enrich', names_to = 'population')

  # add counts per window for current population + ctrl
  enrich_ints_by_bin %>%
    dplyr::left_join(counts_long, by = c("start_interval","interval_number","mid_interval","population")) %>%
    dplyr::left_join(ctrl_counts, by = "interval_number")
}


enrichm_function_fixed_bins = function(TIB, CL = "CBS16", EXP = all_exp, binsize = 1000, N_bins_window = 10,  calculation_window =  c(34091354, 35744137), MOTIF = c("foward","reverse")){
  enrich_ls = lapply(CL, function(cl){
    enrich_cl = enrichm_function_fixed_bins_CL(TIB, CL = cl, EXP, binsize = binsize, N_bins_window = N_bins_window, calculation_window = calculation_window,  MOTIF)
    enrich_cl
  })
  names(enrich_ls) = CL
  bind_rows(enrich_ls, .id = 'cell_line')
}

enrichm_function_fixed_bins_stranded = function(TIB, CL = "CBS16", EXP = all_exp, binsize = 1000, N_bins_window = 10,  calculation_window =  c(34091354, 35744137)){
  bind_rows("forward" = enrichm_function_fixed_bins(TIB, CL, EXP, binsize = binsize, N_bins_window = N_bins_window, calculation_window = calculation_window, MOTIF = "forward"),
            "reverse" = enrichm_function_fixed_bins(TIB, CL, EXP, binsize = binsize, N_bins_window = N_bins_window, calculation_window = calculation_window, MOTIF = "reverse"), .id = 'motif')}

```

## Adapt boostrapping for non stranded application
- tib_filt has integrations in the Sox2 gene fitlered out
- tib_CBS16_16A_comb has integrations in the Sox2 gene still present
```{r bootstrap enrichment, results = FALSE}

tib_boot = tib_filt %>% 
  filter(population %in% c("ctrl", "mCh_mid", "mCh_high")) %>% 
  ungroup()

# 1) Do bootstrap replicates, computing enrichment each time
set.seed(456)
n_boot <- 1000

boot_enrich_list <- map(seq_len(n_boot), function(bi) {
  cat("▶ Bootstrap", bi, "\n")
  
  # 1a) resample with replacement
  boot_dat <- tib_boot %>% slice_sample(prop = 1, replace = TRUE)
  
  # 1b) quick check
  cat("    rows:", nrow(boot_dat),
      " uniq:", n_distinct(boot_dat), "\n")
  
  # 1c) compute enrichment
  enrichm_function_fixed_bins(boot_dat,
                              binsize = 2000,
                              N_bins_window = 10,
                              calculation_window = prange_zoom_paper,
                              CL = c("CBS16_16A", "CRE6", "3xCTCF"),
                              EXP = all_exp,
                              MOTIF = c("forward", "reverse")) %>%
    mutate(bootstrap = bi)
})

# Save object 
saveRDS(boot_enrich_list, "/DATA/projects/CTCF_Sox2/analysis/Enrichment/ME20260303_enrichment_CBS16_16A_3xCBS_CRE_1000boot_2000_10.rds")
boot_enrich_list = readRDS("/DATA/projects/CTCF_Sox2/analysis/Enrichment/ME20260303_enrichment_CBS16_16A_3xCBS_CRE_1000boot_2000_10.rds")

# 2) bind into one data-frame
boot_enrich_df <- bind_rows(boot_enrich_list)

# 3) verify we have the same number of enrichment-rows for each replicate
boot_enrich_df %>% count(bootstrap)

# 4) Compute 95% CI at each mid_interval × population × cell_line × motif
ci_df <- boot_enrich_df %>%
  dplyr::filter(population != "ctrl") %>%
  dplyr::filter(ctrl_counts_window >= 3) %>%
  dplyr::filter(is.finite(int_enrich)) %>%
  group_by(cell_line, mid_interval, population) %>%
  summarize(
    lower  = quantile(int_enrich, 0.025, na.rm = TRUE),
    median = quantile(int_enrich, 0.5,   na.rm = TRUE),
    upper  = quantile(int_enrich, 0.975, na.rm = TRUE),
    .groups = "drop"
  )

# 5) Spot-check one position
ci_df %>% filter(population == "mCh_mid") %>% head()
```

### Calculate actual enrichment
```{r}
enrich_fixed = enrichm_function_fixed_bins(tib_filt,
                                           binsize = 2000,
                                           N_bins_window = 10,
                                           calculation_window = prange_zoom_paper,
                                           CL = c("CBS16_16A", "CRE6", "3xCTCF"),
                                           EXP = all_exp,
                                           MOTIF = c("forward", "reverse"))
```

### Figure 5D - Combine boostrapped and real enrich and plot
```{r fig 5D}
tib_boot_real_enrich = ci_df %>% left_join(enrich_fixed, join_by(cell_line, mid_interval, population))
# Change oder for plotting
tib_boot_real_enrich = tib_boot_real_enrich %>% mutate(cell_line = factor(cell_line, levels = c("CRE6", "CBS16_16A", "3xCTCF")))


# Plotted in linear space (median)
tib_boot_real_enrich %>% 
  filter(population == "mCh_mid" & ctrl_counts_window >= 1) %>%
  ggplot(aes(x = mid_interval, y = median)) +
  geom_line(aes(col = cell_line)) +
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = cell_line), alpha = 0.2) +
  #geom_line(aes(y = int_enrich), col = 'orange') + # enrich of original data
  facet_grid(cell_line ~ ., scales = "free_y") +
  labs(
    x     = "Window midpoint (bp)",
    y     = "Enrichment (sorted / ctrl)",
    title = "Bootstrap 95% CI of enrichment"
  ) +
    list(
    geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown),
    annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown),
    geom_vline(xintercept = start(landingPad_23), col = "darkgrey"),
    geom_hline(yintercept = 1, col = "black", linetype = "dotted"),
    #annotate("rect", xmin = start(SRR2_element), xmax = end(SRR2_element), ymin = -Inf, ymax = Inf, alpha = .2, fill = 'purple'),
    # annotate gene
    annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf, alpha = .2, fill = 'red'),
    geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = 'red')) +
  
  scale_x_continuous(breaks = scales::pretty_breaks(n = 6), labels = scales::unit_format(scale = 1E-6, accuracy = 0.01, unit=NULL),
                     limits= prange_zoom_paper, expand=c(0,0)) +
  theme_bw() +
  theme(legend.position = "bottom")



# Plotted in log2 (median)
tib_boot_real_enrich %>% 
  filter(population == "mCh_mid") %>%
  ggplot(aes(x = mid_interval, y = log2(median))) +
  geom_line(aes(color = cell_line)) +
  geom_ribbon(aes(ymin = log2(lower), ymax = log2(upper),  fill = cell_line), alpha = 0.2) +
  #geom_line(aes(y = log2(int_enrich), col = 'orange')) + # enrich of original data
  facet_grid(cell_line ~ .) +
  labs(
    x     = "Window midpoint (bp)",
    y     = "Enrichment (sorted / ctrl)",
    title = "Bootstrap 95% CI of enrichment"
  ) +
    list(
    geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown),
    annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown),
    geom_vline(xintercept = start(landingPad_23), col = "darkgrey"),
    geom_hline(yintercept = 0, col = "black", linetype = "dotted"),
    #annotate("rect", xmin = start(SRR2_element), xmax = end(SRR2_element), ymin = -Inf, ymax = Inf, alpha = .2, fill = 'purple'),
    # annotate gene
    annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf, alpha = .2, fill = 'red'),
    geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = 'red')) +
  
  scale_x_continuous(breaks = scales::pretty_breaks(n = 6), labels = scales::unit_format(scale = 1E-6, accuracy = 0.01, unit=NULL),
                     limits= prange_zoom_paper, expand=c(0,0)) +
  theme_bw() +
  theme(legend.position = "bottom")
```

### Plot CRE with CBS cell lines together
```{r}
# Plot CBS16_16A and CRE together
tib_boot_real_enrich %>% 
  filter(population == "mCh_mid" & cell_line %in% c("CRE6", "CBS16_16A")) %>%
  ggplot(aes(x = mid_interval, y = median)) +
  geom_line(aes( color = cell_line)) +
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = cell_line), alpha = 0.2) +
  #geom_line(aes(y = int_enrich), col = 'orange') + # enrich of original data
  #facet_grid(cell_line ~ .) +
  labs(
    x     = "Window midpoint (bp)",
    y     = "Enrichment (sorted / ctrl)",
    title = "Bootstrap 95% CI of enrichment"
  ) +
    list(
    geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown),
    annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown),
    geom_vline(xintercept = start(landingPad_23), col = "darkgrey"),
    geom_hline(yintercept = 1, col = "black", linetype = "dotted"),
    #annotate("rect", xmin = start(SRR2_element), xmax = end(SRR2_element), ymin = -Inf, ymax = Inf, alpha = .2, fill = 'purple'),
    # annotate gene
    annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf, alpha = .2, fill = 'red'),
    geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = 'red')) +
  
  scale_x_continuous(breaks = scales::pretty_breaks(n = 6), labels = scales::unit_format(scale = 1E-6, accuracy = 0.01, unit=NULL),
                     limits= prange_zoom_paper, expand=c(0,0)) +
  theme_classic() +
  theme(legend.position = "bottom")



# Plot CRE and 3xCBS together
tib_boot_real_enrich %>% 
  filter(population == "mCh_mid" & cell_line %in% c("CRE6", "3xCTCF")) %>%
  ggplot(aes(x = mid_interval, y = median)) +
  geom_line(aes(color = cell_line)) +
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = cell_line), alpha = 0.2) +
  #geom_line(aes(y = int_enrich), col = 'orange') + # enrich of original data
  #facet_grid(cell_line ~ .) +
  labs(
    x     = "Window midpoint (bp)",
    y     = "Enrichment (sorted / ctrl)",
    title = "Bootstrap 95% CI of enrichment"
  ) +
    list(
    geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown),
    annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown),
    geom_vline(xintercept = start(landingPad_23), col = "darkgrey"),
    geom_hline(yintercept = 1, col = "black", linetype = "dotted"),
    #annotate("rect", xmin = start(SRR2_element), xmax = end(SRR2_element), ymin = -Inf, ymax = Inf, alpha = .2, fill = 'purple'),
    # annotate gene
    annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf, alpha = .2, fill = 'red'),
    geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = 'red')) +
  
  scale_x_continuous(breaks = scales::pretty_breaks(n = 6), labels = scales::unit_format(scale = 1E-6, accuracy = 0.01, unit=NULL),
                     limits= prange_zoom_paper, expand=c(0,0)) +
  theme_classic() +
  theme(legend.position = "bottom")


# Plot all three together
tib_boot_real_enrich %>% 
  filter(population == "mCh_mid") %>%
  ggplot(aes(x = mid_interval, y = median)) +
  geom_line(aes(color = cell_line)) +
  geom_ribbon(aes(ymin = lower, ymax = upper,  fill = cell_line), alpha = 0.2) +
  #geom_line(aes(y = int_enrich), col = 'orange') + # enrich of original data
  #facet_grid(cell_line ~ .) +
  labs(
    x     = "Window midpoint (bp)",
    y     = "Enrichment (sorted / ctrl)",
    title = "Bootstrap 95% CI of enrichment"
  ) +
    list(
    geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown),
    annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown),
    geom_vline(xintercept = start(landingPad_23), col = "darkgrey"),
    geom_hline(yintercept = 1, col = "black", linetype = "dotted"),
    #annotate("rect", xmin = start(SRR2_element), xmax = end(SRR2_element), ymin = -Inf, ymax = Inf, alpha = .2, fill = 'purple'),
    # annotate gene
    annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf, alpha = .2, fill = 'red'),
    geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = 'red')) +
  
  scale_x_continuous(breaks = scales::pretty_breaks(n = 6), labels = scales::unit_format(scale = 1E-6, accuracy = 0.01, unit=NULL),
                     limits= prange_zoom_paper, expand=c(0,0)) +
  theme_classic() +
  theme(legend.position = "bottom")
```

### Fig 5D numbers - Extracting min/max values
Exract bootstraped and real enrichment between Sox2-SCR interval for the different cell lines.
```{r Fig 5D numbers - extracting min/max values}
#Extracting min/max values of bootstrapped mean
sum_CL_boot = tib_boot_real_enrich %>% 
  select(cell_line, median, population) %>% 
  group_by(cell_line, population) %>% 
  summarise(mean = mean(median),
            max = max(median),
            min = min(median))

sum_CL_boot_Sox2_SCR = tib_boot_real_enrich %>% 
  filter(start_interval >= end(Sox2_gene) & start_interval <= start(enhancer)) %>% 
  select(cell_line, population, median) %>% 
  group_by(cell_line, population) %>% 
  summarise(mean = mean(median),
            max = max(median),
            min = min(median))


#Extracting min/max values of bootstrapped mean
sum_CL_real = tib_boot_real_enrich %>% 
  filter(start_interval >= end(Sox2_gene) & start_interval <= start(enhancer)) %>% 
  select(cell_line, int_enrich, population) %>% 
  group_by(cell_line, population) %>% 
  summarise(mean = mean(int_enrich),
            max = max(int_enrich),
            min = min(int_enrich))
```
**Conclusion:** Boostrapping 1000x leads to almost same min, max, mean enrichment compared to real measured data. 


# Calculate fraction of cells that are between Sox2-SCR interval or outside
```{r Figure 5E - calculate fraction of int for region Sox_SCR and outside}
# outside = 1Mb upstream from the beginning of Sox2 gene till the start of the Sox2 gene and 1Mb downstream of the end+1 of the SCR
# inside = end of Sox2 gene till start of SCR

tib_ctrl_fract_tot = tib_filt %>% 
  filter(population %in% c("ctrl", "mCh_mid") & chr == "chr3" & hopped & start >= (start(Sox2_gene)-1e6) & start <= end(enhancer)+1e6) %>%
  group_by(cell_line, population) %>% 
  summarise(total_n = n())
  
# coordinate of the launch pad (numeric)
outside_upstream_coord <- start(landingPad_23)

# count upstream and downstream separately
tib_ctrl_fract_split <- tib_filt %>%
  filter(population %in% c("ctrl", "mCh_mid"),
         chr == "chr3",
         hopped,
         start >= (start(Sox2_gene)-1e6),
         start <= (end(enhancer)+1e6)) %>%
  mutate(region = case_when(
    start < start(Sox2_gene) ~ "outside",
    start > start(enhancer)+1 ~ "outside",
    TRUE ~ "Sox2_SCR_interval"
  )) %>%
  group_by(cell_line, population, region) %>%
  summarise(n = n(), .groups = "drop") %>%
  # join with total counts
  left_join(tib_ctrl_fract_tot, by = c("cell_line", "population")) %>%
  mutate(frac = n / total_n)

tib_ctrl_fract_split
```

## Function to calculate p-value fisher test
```{r compute Fisher exact test for enrichment of integrations in between vs outside per cell line}
# your tibble is called integration_table
integration_table = tib_ctrl_fract_split %>% filter(cell_line != "37A")

# Make sure the order is right for the fisher test, we group each cell line and transform to a 2x2 matrix, rows = population, col = region

integration_table <- integration_table %>%
  mutate(
    population = factor(population, levels = c("mCh_mid", "ctrl")), # this order is important, since odds ratio of >1 means enrichment of ints in mid over ctrl between Sox2-SCR interval compared to outside
    region = factor(region, levels = c("Sox2_SCR_interval", "outside"))
  ) %>%
  arrange(cell_line, population, region)


# run fisher test per cell line
fisher_results <- integration_table %>%
  group_by(cell_line) %>%
  summarise(
    {mat <- matrix(n, nrow = 2, byrow = TRUE)
      ft  <- fisher.test(mat)

      tibble(
        p_value    = ft$p.value,
        odds_ratio = unname(ft$estimate),
        conf_low   = ft$conf.int[1],
        conf_high  = ft$conf.int[2])
    }
  )

fisher_results
```
**Conclusion:** All 3 cell lines show significantly more integrations between Sox2-SCR in the mCh-mid sorted population compared to control. CRE only boderline significant and lowest odds ratio, CBS containg ones much more convincing

CRE6 has ~1.26x higher odds of having integrations in the Sox2-SCR interval (vs outside) compared to control. Modest enrichment but real.
CBS16_16A = mCh_mid has almost ~2x the odds of integrating in the Sox2-SCR vs outside, relative to ctrl
3xCTCF = ~3.5x more likely to have int. in teh Sox2-SCR interval rather than outside, compared to ctrl. 

## Plot results
```{r Plot Fract of ctrl vs mCh_mid}
ggplot(tib_ctrl_fract_split %>% filter(cell_line != "37A"),
       aes(x = population, y = frac, fill = region)) +
  geom_col(position = "fill") +
  facet_wrap(~ cell_line) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = c("Sox2_SCR_interval" = "skyblue", "outside" = "orange")) +
  ylab("Fraction of integrations") +
  xlab("Cell line") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Figure 5E - Build in labels to plot
```{r Figure 5E -  Plot Fract of ctrl vs mCh_mid with sign and OR labels}

# 1. Prepare plotting data (drop 37A, fix x order)
plot_data <- tib_ctrl_fract_split %>%
  filter(cell_line != "37A") %>%
  mutate(population = factor(population, levels = c("ctrl", "mCh_mid")))

# 2. Prepare annotation data
label_df <- fisher_results %>%
  mutate(
    label = paste0(
      "OR = ", round(odds_ratio, 2),
      " [", round(conf_low, 2), ", ", round(conf_high, 2), "]",
      "\nP = ", signif(p_value, 2)
    ),
    population = factor("mCh_mid", levels = c("ctrl", "mCh_mid")),
    y = 1
  )

# 3. Plot with annotations
ggplot(plot_data,
       aes(x = population, y = frac, fill = region)) +
  geom_col(position = "fill") +
  facet_grid(cell_line ~ .) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = c(
    "Sox2_SCR_interval" = "skyblue",
    "outside" = "orange"
  )) +
  ylab("Fraction of integrations") +
  xlab("Cell line") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_text(
    data = label_df,
    aes(x = population, y = y, label = label),
    inherit.aes = FALSE,
    vjust = 1.1,
    size = 3
  )

#Alternative is forest plot style
ggplot(fisher_results, aes(x = cell_line, y = odds_ratio)) +
  geom_pointrange(aes(ymin = conf_low, ymax = conf_high)) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_y_log10() +
  ylab("Odds ratio (mCh_mid vs ctrl)") +
  xlab("Cell line") +
  theme_bw()


# Figure 5E
plot_data %>%
  filter(cell_line != "37A") %>%
  group_by(cell_line, population) %>%
  mutate(
    perc  = frac / sum(frac) * 100,        # convert to %
    ypos  = cumsum(frac) - 0.5 * frac      # vertical midpoint for label
  ) %>%
  ggplot(aes(x = population, y = frac, fill = region)) +
  geom_col(position = "fill") +
  geom_text(aes(y = ypos, label = sprintf("%.1f%%", perc)),
            size = 3, color = "black") +
  facet_grid(cell_line ~ .) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = c(
    "Sox2_SCR_interval" = "skyblue",
    "outside" = "orange"
  )) +
  labs(
    x = "Population",
    y = "Fraction of integrations",
    fill = "Region"
  ) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(legend.position = "bottom") +
  geom_text(
    data = label_df,
    aes(x = population, y = y, label = label),
    inherit.aes = FALSE,
    vjust = 1.1,
    size = 3
  )
```


# Stranded enrichment
## Stranded Boostrapping
```{r stranded bootstrap enrichment, results = FALSE}

tib_boot = tib_filt %>% 
  filter(population %in% c("ctrl", "mCh_mid", "mCh_high")) %>% 
  ungroup()

# 1) Do bootstrap replicates, computing enrichment each time
set.seed(456)
n_boot <- 1000

boot_enrich_list <- map(seq_len(n_boot), function(bi) {
  cat("▶ Bootstrap", bi, "\n")
  
  # 1a) resample with replacement
  boot_dat <- tib_boot %>% slice_sample(prop = 1, replace = TRUE)
  
  # 1b) quick check
  cat("    rows:", nrow(boot_dat),
      " uniq:", n_distinct(boot_dat), "\n")
  
  # 1c) compute stranded enrichment
  enrichm_function_fixed_bins_stranded(
    TIB                = boot_dat,
    CL                 = c("CBS16_16A", "3xCTCF", "CRE6"),
    EXP                = all_exp,
    binsize            = 2000,
    N_bins_window      = 10,
    calculation_window = prange_zoom_paper
  ) %>%
    mutate(bootstrap = bi)
})

# Save object 
saveRDS(boot_enrich_list, "/DATA/projects/CTCF_Sox2/analysis/Enrichment/ME20260303_enrichment_STRANDED_CBS16_16A_3xCBS_CRE_1000boot_2000_10.rds")
boot_enrich_list = readRDS("/DATA/projects/CTCF_Sox2/analysis/Enrichment/ME20260303_enrichment_STRANDED_CBS16_16A_3xCBS_CRE_1000boot_2000_10.rds")


# 2) bind into one data-frame
boot_enrich_df <- bind_rows(boot_enrich_list)

# 3) verify we have the same number of enrichment-rows for each replicate
boot_enrich_df %>% count(bootstrap)

# 4) Compute 95% CI at each mid_interval × population × cell_line × motif
ci_df <- boot_enrich_df %>%
  dplyr::filter(population != "ctrl") %>%
  dplyr::filter(ctrl_counts_window >= 2) %>%
  dplyr::filter(is.finite(int_enrich)) %>%
  group_by(cell_line, motif, mid_interval, population) %>%
  summarize(
    lower  = quantile(int_enrich, 0.025, na.rm = TRUE),
    median = quantile(int_enrich, 0.5,   na.rm = TRUE),
    upper  = quantile(int_enrich, 0.975, na.rm = TRUE),
    .groups = "drop"
  )

# 5) Spot-check one position
ci_df %>% filter(population == "mCh_mid") %>% head()
```

## Calculate real stranded enrichment
```{r}
enrich_fixed_stranded = enrichm_function_fixed_bins_stranded(tib_filt,
                                           binsize = 2000,
                                           N_bins_window = 10,
                                           calculation_window = prange_zoom_paper,
                                           CL = c("CBS16_16A", "CRE6", "3xCTCF"),
                                           EXP = all_exp)
```

### Figure 5F - Combine boostrapped and real enrich and plot
```{r Figure 5F}
tib_boot_real_enrich_strand = ci_df %>% left_join(enrich_fixed_stranded, join_by(cell_line, motif, mid_interval, population))
# Change oder for plotting
tib_boot_real_enrich_strand = tib_boot_real_enrich_strand %>% mutate(cell_line = factor(cell_line, levels = c("CRE6", "CBS16_16A", "3xCTCF")))

#facet by motif
CTCF_df <- as.data.frame(CTCF_mm10.chr3_extra) %>%
  transmute(
    mid = (start + end) / 2,
    strand = factor(as.character(strand))
  )

tib_boot_real_enrich_strand %>% 
  filter(population == "mCh_mid") %>%
  ggplot(aes(x = mid_interval, y = log2(median))) +
  geom_line(aes(color = cell_line)) +
  geom_ribbon(aes(ymin = log2(lower), ymax = log2(upper),fill = cell_line), alpha = 0.2) +
  #geom_line(aes(y = log2(int_enrich)), col = 'orange') + # enrich of original data
  facet_grid(motif ~ .) +
  labs(
    x     = "Window midpoint (bp)",
    y     = "Enrichment log2 (sorted / ctrl)",
    title = "Bootstrap 95% CI of enrichment"
  ) +
  list(
    geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown),
    annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown),
    geom_vline(xintercept = start(landingPad_23), col = "darkgrey"),
    geom_hline(yintercept = 0, col = "black", linetype = "dotted"),
    #annotate("rect", xmin = start(SRR2_element), xmax = end(SRR2_element), ymin = -Inf, ymax = Inf, alpha = .2, fill = 'purple'),
    # annotate gene
    annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf, alpha = .2, fill = 'red'),
    geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = 'red')) +
  
  # NEW: start a second colour scale for the vlines
  ggnewscale::new_scale_colour() +
  geom_vline(
    data = CTCF_df,
    aes(xintercept = mid, colour = strand),
    linetype = "dotted",
    linewidth = 0.3,
    inherit.aes = FALSE
  ) +
  scale_colour_manual(values = c("+" = "grey30", "-" = "lightgrey"), name = "CTCF") +
  
  scale_x_continuous(breaks = scales::pretty_breaks(n = 6), labels = scales::unit_format(scale = 1E-6, accuracy = 0.01, unit=NULL),
                     limits= prange_zoom_paper, expand=c(0,0)) +
  theme_classic() +
  theme(legend.position = "bottom")
```

### Extracting min/max values - numbers for Fig. 5F
```{r extracting min/max values of stranded}
#Extracting min/max values of bootstrapped mean
sum_CL_boot_strand = tib_boot_real_enrich_strand %>% 
  select(cell_line, median, population, motif) %>% 
  group_by(cell_line, population, motif) %>% 
  summarise(mean = mean(median),
            max = max(median),
            min = min(median))

sum_CL_boot_Sox2_SCR_strand = tib_boot_real_enrich_strand %>% 
  filter(start_interval >= end(Sox2_gene) & start_interval <= start(enhancer)) %>% 
  select(cell_line, median, population, motif) %>% 
  group_by(cell_line, population, motif) %>% 
  summarise(mean = mean(median),
            max = max(median),
            min = min(median))


#Extracting min/max values of bootstrapped mean
sum_CL_real_strand = tib_boot_real_enrich_strand %>% 
  select(cell_line, int_enrich, population, motif) %>% 
  group_by(cell_line, population, motif) %>% 
  summarise(mean = mean(int_enrich),
            max = max(int_enrich),
            min = min(int_enrich))


sum_CL_real_strand_Sox2_SCR = tib_boot_real_enrich_strand %>% 
  filter(start_interval >= start(Sox2_gene) & start_interval <= start(enhancer)) %>% 
  select(cell_line, int_enrich, motif) %>% 
  group_by(cell_line, motif) %>% 
  summarise(mean = mean(int_enrich),
            max = max(int_enrich),
            min = min(int_enrich))
```
**Conclusion:** Boostrapping 1000x leads to almost same min, max, mean enrichment compared to real measured data. 

# Stranded Calculate fraction of cells that are between Sox2-SCR interval or outside
```{r stranded - calculate fraction of int for region Sox_SCR and outside}
# outside = 1Mb upstream from the beginning of Sox2 gene till the start of the Sox2 gene and 1Mb downstream of the end+1 of the SCR
# inside = end of Sox2 gene till end of SCR

tib_ctrl_fract_tot_motif = tib_filt %>% 
  filter(population %in% c("ctrl", "mCh_mid") & chr == "chr3" & hopped & start >= (end(Sox2_gene)-1e6) & start <= start(enhancer)+1e6) %>%
  group_by(cell_line, population, motif) %>% 
  summarise(total_n = n())
  

# count upstream and downstream separately
tib_ctrl_fract_split <- tib_filt %>%
  filter(population %in% c("ctrl", "mCh_mid"),
         chr == "chr3",
         hopped,
         start >= (start(Sox2_gene)-1e6),
         start <= (end(enhancer)+1e6)) %>%
  mutate(region = case_when(
    start < start(Sox2_gene) ~ "outside",
    start > start(enhancer)+1 ~ "outside",
    TRUE ~ "Sox2_SCR_interval"
  )) %>%
  group_by(cell_line, population, motif, region) %>%
  summarise(n = n(), .groups = "drop") %>%
  # join with total counts
  left_join(tib_ctrl_fract_tot_motif, by = c("cell_line", "population", "motif")) %>%
  mutate(frac = n / total_n)

tib_ctrl_fract_split
```

## Function to calculate p-value fisher test
## Fisher test - is there a difference between inside and outside of ctrl and mCh_mid sorted pools per motif
```{r Fisher test for testing significant difference between outside vs inside per motif}
# Make sure region + population have consistent levels everywhere
integration_table2 <- tib_ctrl_fract_split %>%
  filter(cell_line != "37A",
         population %in% c("ctrl", "mCh_mid"),
         motif %in% c("forward", "reverse"),
         region %in% c("Sox2_SCR_interval", "outside")) %>%
  mutate(
    population = factor(population, levels = c("mCh_mid", "ctrl")),
    region     = factor(region, levels = c("Sox2_SCR_interval", "outside")),
    motif      = factor(motif, levels = c("forward", "reverse"))
  )

fisher_results_region_by_motif <- integration_table2 %>%
  group_by(cell_line, motif) %>%
  summarise(
    mat = list({
      # 2x2 in a defined order:
      # rows: mCh_mid, ctrl
      # cols: Sox2_SCR_interval, outside
      xt <- xtabs(n ~ population + region, data = pick(n, population, region))
      xt[c("mCh_mid","ctrl"), c("Sox2_SCR_interval","outside")]
    }),
    .groups = "drop"
  ) %>%
  mutate(
    ft = map(mat, ~ fisher.test(.x)),
    p_value    = map_dbl(ft, "p.value"),
    odds_ratio = map_dbl(ft, ~ unname(.x$estimate)),
    conf_low   = map_dbl(ft, ~ .x$conf.int[1]),
    conf_high  = map_dbl(ft, ~ .x$conf.int[2])
  ) %>%
  select(cell_line, motif, p_value, odds_ratio, conf_low, conf_high)

fisher_results_region_by_motif
```
### Figure S5C - Plot results
```{r Figure S5C}
# 1) Prepare plotting data (drop 37A, fix x order)
plot_data <- tib_ctrl_fract_split %>%
  filter(cell_line != "37A") %>%
  mutate(
    population = factor(population, levels = c("ctrl", "mCh_mid")),
    region     = factor(region, levels = c("outside", "Sox2_SCR_interval")),
    motif      = factor(motif, levels = c("forward", "reverse"))
  )

# 2) Prepare annotation data (one label per facet: cell_line x motif)
label_df <- fisher_results_region_by_motif %>%
  mutate(
    motif = factor(motif, levels = c("forward", "reverse")),
    label = paste0(
      "OR = ", round(odds_ratio, 2),
      " [", round(conf_low, 2), ", ", round(conf_high, 2), "]",
      "\nP = ", signif(p_value, 2)
    ),
    population = factor("mCh_mid", levels = c("ctrl", "mCh_mid")),
    y = 1
  )

# 3) Plot with annotations
ggplot(plot_data, aes(x = population, y = frac, fill = region)) +
  geom_col(position = "fill") +
  facet_grid(cell_line ~ motif) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = c(
    "Sox2_SCR_interval" = "skyblue",
    "outside" = "orange"
  )) +
  ylab("Fraction of integrations") +
  xlab("Population") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_text(
    data = label_df,
    aes(x = population, y = y, label = label),
    inherit.aes = FALSE,
    vjust = 1.1,
    size = 3
  )

# Plot with annotated % as well
plot_data %>%
  filter(cell_line != "37A") %>%
  mutate(
    population = factor(population, levels = c("ctrl", "mCh_mid")),
    motif      = factor(motif, levels = c("forward", "reverse"))
  ) %>%
  group_by(cell_line, motif, population) %>%
  mutate(
    perc = frac / sum(frac) * 100,          # % within each stacked bar
    ypos = cumsum(frac) - 0.5 * frac        # midpoint for label
  ) %>%
  ungroup() %>%
  ggplot(aes(x = population, y = frac, fill = region)) +
  geom_col(position = "fill") +
  geom_text(aes(y = ypos, label = sprintf("%.1f%%", perc)),
            size = 3, color = "black") +
  facet_grid(cell_line ~ motif) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = c(
    "Sox2_SCR_interval" = "skyblue",
    "outside" = "orange"
  )) +
  labs(
    x = "Population",
    y = "Fraction of integrations",
    fill = "Region"
  ) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(legend.position = "bottom") +
  geom_text(
    data = label_df %>% mutate(motif = factor(motif, levels = c("forward","reverse"))),
    aes(x = population, y = y, label = label),
    inherit.aes = FALSE,
    vjust = 1.1,
    size = 3
  )

```


# Explore high and mid integrations
## Not stranded - Enrich & boostrapping high int
```{r bootstrap enrichment Sox2 mCh high, results = FALSE}

tib_boot = tib_filt %>% 
  filter(population %in% c("ctrl", "mCh_mid", "mCh_high")) %>% 
  ungroup()

# 1) Do 5 bootstrap replicates, computing enrichment each time
set.seed(456)
n_boot <- 1000

boot_enrich_list <- map(seq_len(n_boot), function(bi) {
  cat("▶ Bootstrap", bi, "\n")
  
  # 1a) resample with replacement
  boot_dat <- tib_boot %>% slice_sample(prop = 1, replace = TRUE)
  
  # 1b) quick check
  cat("    rows:", nrow(boot_dat),
      " uniq:", n_distinct(boot_dat), "\n")
  
  # 1c) compute enrichment
  enrichm_function_fixed_bins(boot_dat,
                              binsize = 5000,
                              N_bins_window = 5,
                              calculation_window = prange_zoom_paper,
                              CL = c("CBS16_16A", "CRE6", "3xCTCF"),
                              EXP = all_exp,
                              MOTIF = c("forward", "reverse")) %>%
    mutate(bootstrap = bi)
})

# Save object 
saveRDS(boot_enrich_list, "/DATA/projects/CTCF_Sox2/analysis/Enrichment/ME20260303_enrichment_CBS16_16A_3xCBS_CRE_1000boot_5000_5.rds")
boot_enrich_list = readRDS("/DATA/projects/CTCF_Sox2/analysis/Enrichment/ME20260303_enrichment_CBS16_16A_3xCBS_CRE_1000boot_5000_5.rds")


# 2) bind into one data-frame
boot_enrich_df <- bind_rows(boot_enrich_list)

# 3) verify we have the same number of enrichment-rows for each replicate
boot_enrich_df %>% count(bootstrap)

# 4) Compute 95% CI at each mid_interval × population × cell_line × motif
ci_df <- boot_enrich_df %>%
  dplyr::filter(population != "ctrl") %>%
  dplyr::filter(ctrl_counts_window >= 1) %>%
  dplyr::filter(is.finite(int_enrich)) %>%
  group_by(cell_line, mid_interval, population) %>%
  summarize(
    lower  = quantile(int_enrich, 0.025, na.rm = TRUE),
    median = quantile(int_enrich, 0.5,   na.rm = TRUE),
    upper  = quantile(int_enrich, 0.975, na.rm = TRUE),
    .groups = "drop"
  )

# 5) Spot-check one position
ci_df %>% filter(population == "mCh_high") %>% head()
```

### Calculate actual enrichment
```{r enrichment Sox2 mCh_high}
enrich_fixed_high = enrichm_function_fixed_bins(tib_filt,
                                           binsize = 5000,
                                           N_bins_window = 5,
                                           calculation_window = prange_zoom_paper,
                                           CL = c("CBS16_16A", "CRE6", "3xCTCF"),
                                           EXP = all_exp,
                                           MOTIF = c("forward", "reverse"))
```

### Combine boostrapped and real enrich and plot
```{r}
tib_boot_real_enrich = ci_df %>% left_join(enrich_fixed_high, join_by(cell_line, mid_interval, population))
# Change oder for plotting
tib_boot_real_enrich = tib_boot_real_enrich %>% mutate(cell_line = factor(cell_line, levels = c("CRE6", "CBS16_16A", "3xCTCF")))


# Plot non stranded
tib_boot_real_enrich %>% 
  filter(ctrl_counts_window >= 1 & counts_window >= 1) %>%
  ggplot(aes(x = mid_interval, y = median, color = population, fill = population)) +
  #geom_point() +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2) +
  #geom_line(aes(y = int_enrich), col = 'orange') + # enrich of original data
  facet_grid(cell_line ~ ., scales = "free_y") +
  labs(
    x     = "Window midpoint (bp)",
    y     = "Enrichment (sorted / ctrl)",
    title = "Bootstrap 95% CI of enrichment"
  ) +
    list(
    geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown),
    annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown),
    geom_vline(xintercept = start(landingPad_23), col = "darkgrey"),
    geom_hline(yintercept = 1, col = "black", linetype = "dotted"),
    #annotate("rect", xmin = start(SRR2_element), xmax = end(SRR2_element), ymin = -Inf, ymax = Inf, alpha = .2, fill = 'purple'),
    # annotate gene
    annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf, alpha = .2, fill = 'red'),
    geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = 'red')) +
  
  scale_x_continuous(breaks = scales::pretty_breaks(n = 6), labels = scales::unit_format(scale = 1E-6, accuracy = 0.1, unit=NULL),
                     limits= prange_zoom_paper, expand=c(0,0)) +
  theme_bw() +
  theme(legend.position = "bottom")
```

```{r}
#add CRE line to the plot
# break at any binsize bigger than 5kb
bin <- 5000

pEnrich_high_1xCBS_CRE_25kb =
tib_boot_real_enrich %>% 
  filter(ctrl_counts_window >= 1 & counts_window >= 1 & cell_line %in% c("CRE6", "CBS16_16A")) %>%
  mutate(CL_pop = case_when(cell_line == "CRE6" & population == "mCh_mid" ~ "CRE6_mCh_mid",
                            cell_line == "CBS16_16A" & population == "mCh_mid" ~ "CBS16_16A_mCh_mid",
                            cell_line == "CBS16_16A" & population == "mCh_high" ~ "CBS16_16A_mCh_high",
                            cell_line == "3xCTCF" & population == "mCh_mid" ~ "3xCTCF_mCh_mid",
                            cell_line == "3xCTCF" & population == "mCh_high" ~ "3xCTCF_mCh_high",
                            .default = "other")) %>%
  filter(CL_pop != "other") %>%
  arrange(CL_pop, mid_interval) %>%
  group_by(CL_pop) %>%
  mutate(
    gap = mid_interval - lag(mid_interval),
    seg = cumsum(replace_na(gap > bin, FALSE))
  ) %>%
  ungroup() %>%
  ggplot(aes(
    x = mid_interval, y = median,
    group = interaction(CL_pop, seg)   # <-- THIS is the key bit
  )) +
  geom_line(aes(color = CL_pop)) +
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = CL_pop), alpha = 0.2) +
  #geom_line(aes(y = int_enrich), col = 'orange') + # enrich of original data
  #facet_grid(cell_line ~ ., scales = "free_y") +
  labs(
    x     = "Window midpoint (bp)",
    y     = "Enrichment (sorted / ctrl)",
    title = "Bootstrap 95% CI of enrichment"
  ) +
    list(
    geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown),
    annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown),
    geom_vline(xintercept = start(landingPad_23), col = "darkgrey"),
    geom_hline(yintercept = 1, col = "black", linetype = "dotted"),
    #annotate("rect", xmin = start(SRR2_element), xmax = end(SRR2_element), ymin = -Inf, ymax = Inf, alpha = .2, fill = 'purple'),
    # annotate gene
    annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf, alpha = .2, fill = 'red'),
    geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = 'red')) +
  
  scale_x_continuous(breaks = scales::pretty_breaks(n = 6), labels = scales::unit_format(scale = 1E-6, accuracy = 0.1, unit=NULL),
                     limits= prange_zoom_paper, expand=c(0,0)) +
  theme_bw() +
  theme(legend.position = "bottom")


pEnrich_high_3xCBS_CRE_25kb =
  tib_boot_real_enrich %>% 
  filter(ctrl_counts_window >= 1 & counts_window >= 1 & cell_line %in% c("CRE6", "3xCTCF")) %>%
  mutate(CL_pop = case_when(
    cell_line == "CRE6"   & population == "mCh_mid"  ~ "CRE6_mCh_mid",
    cell_line == "3xCTCF" & population == "mCh_mid"  ~ "3xCTCF_mCh_mid",
    cell_line == "3xCTCF" & population == "mCh_high" ~ "3xCTCF_mCh_high",
    .default = "other"
  )) %>%
  filter(CL_pop != "other") %>%
  arrange(CL_pop, mid_interval) %>%
  group_by(CL_pop) %>%
  mutate(
    gap = mid_interval - lag(mid_interval),
    seg = cumsum(replace_na(gap > bin, FALSE))
  ) %>%
  ungroup() %>%
  ggplot(aes(
    x = mid_interval, y = median,
    group = interaction(CL_pop, seg)   
  )) +
  geom_line(aes(color = CL_pop)) +
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = CL_pop), alpha = 0.2) +
  labs(
    x     = "Window midpoint (bp)",
    y     = "Enrichment (sorted / ctrl)",
    title = "Bootstrap 95% CI of enrichment"
  ) +
  list(
    geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown),
    annotate("rect", xmin = start(enhancer), xmax = end(enhancer),
             ymin = -Inf, ymax = Inf, alpha = .2, fill = brown),
    geom_vline(xintercept = start(landingPad_23), col = "darkgrey"),
    geom_hline(yintercept = 1, col = "black", linetype = "dotted"),
    annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene),
             ymin = -Inf, ymax = Inf, alpha = .2, fill = "red"),
    geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = "red")
  ) +
  scale_x_continuous(
    breaks = scales::pretty_breaks(n = 6),
    labels = scales::unit_format(scale = 1E-6, accuracy = 0.1, unit = NULL),
    limits = prange_zoom_paper, expand = c(0,0)
  ) +
  theme_bw() +
  theme(legend.position = "bottom")

plot_grid(pEnrich_high_1xCBS_CRE_25kb, pEnrich_high_3xCBS_CRE_25kb, ncol = 1, axis = "lr", align = "v")
```

### Do not display the missing values
```{r}
plot_df <- tib_boot_real_enrich %>% 
  filter(ctrl_counts_window >= 1 & counts_window >= 1) %>%
  arrange(cell_line, population, mid_interval) %>%
  group_by(cell_line, population) %>%
  mutate(
    step = median(diff(mid_interval), na.rm = TRUE),          # typical spacing in this series
    new_seg = replace_na((mid_interval - lag(mid_interval)) > 1.1 * step, FALSE),
    seg = cumsum(new_seg)                                     # segment id
  ) %>%
  ungroup()

ggplot(plot_df, aes(
  x = mid_interval, y = median,
  color = population, fill = population,
  group = interaction(population, seg)   # <- breaks lines/ribbons at gaps
)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2) +
  facet_grid(cell_line ~ ., scales = "free_y") +
  labs(
    x     = "Window midpoint (bp)",
    y     = "Enrichment (sorted / ctrl)",
    title = "Bootstrap 95% CI of enrichment"
  ) +
  scale_x_continuous(
    breaks = scales::pretty_breaks(n = 6),
    labels = scales::unit_format(scale = 1E-6, accuracy = 0.1, unit = NULL),
    limits = prange_2Mb, expand = c(0,0)
  ) +
  theme_bw() +
  theme(legend.position = "bottom")


# break at any binsize bigger than 5kb
bin <- 5000

plot_df <- tib_boot_real_enrich %>%
  filter(ctrl_counts_window >= 1 & counts_window >= 1) %>%
  arrange(cell_line, population, mid_interval) %>%
  group_by(cell_line, population) %>%
  mutate(
    gap = mid_interval - lag(mid_interval),
    seg = cumsum(replace_na(gap > bin, FALSE))
  ) %>%
  ungroup()

ggplot(
  plot_df,
  aes(
    x = mid_interval, y = median,
    color = population, fill = population,
    group = interaction(population, seg)
  )
) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2) +
  facet_grid(cell_line ~ ., scales = "free_y") +
  theme_bw()
```


## Stranded Boostrapping
```{r mCh_high stranded bootstrap enrichment, results = FALSE}

tib_boot = tib_filt %>% 
  filter(population %in% c("ctrl", "mCh_mid", "mCh_high")) %>% 
  ungroup()

# 1) Do 5 bootstrap replicates, computing enrichment each time
set.seed(456)
n_boot <- 1000

boot_enrich_list <- map(seq_len(n_boot), function(bi) {
  cat("▶ Bootstrap", bi, "\n")
  
  # 1a) resample with replacement
  boot_dat <- tib_boot %>% slice_sample(prop = 1, replace = TRUE)
  
  # 1b) quick check
  cat("    rows:", nrow(boot_dat),
      " uniq:", n_distinct(boot_dat), "\n")
  
  # 1c) compute stranded enrichment
  enrichm_function_fixed_bins_stranded(
    TIB                = boot_dat,
    CL                 = c("CBS16_16A", "3xCTCF", "CRE6"),
    EXP                = all_exp,
    binsize            = 5000,
    N_bins_window      = 10,
    calculation_window = prange_zoom_paper
  ) %>%
    mutate(bootstrap = bi)
})

# Save object 
saveRDS(boot_enrich_list, "/DATA/projects/CTCF_Sox2/analysis/Enrichment/ME20260303_enrichment_STRANDED_CBS16_16A_3xCBS_CRE_1000boot_5000_10.rds")
boot_enrich_list = readRDS("/DATA/projects/CTCF_Sox2/analysis/Enrichment/ME20260303_enrichment_STRANDED_CBS16_16A_3xCBS_CRE_1000boot_5000_10.rds")


# 2) bind into one data-frame
boot_enrich_df <- bind_rows(boot_enrich_list)

# 3) verify we have the same number of enrichment-rows for each replicate
boot_enrich_df %>% count(bootstrap)

# 4) Compute 95% CI at each mid_interval × population × cell_line × motif
ci_df <- boot_enrich_df %>%
  dplyr::filter(population != "ctrl") %>%
  dplyr::filter(ctrl_counts_window >= 1) %>%
  dplyr::filter(is.finite(int_enrich)) %>%
  group_by(cell_line, motif, mid_interval, population) %>%
  summarize(
    lower  = quantile(int_enrich, 0.025, na.rm = TRUE),
    median = quantile(int_enrich, 0.5,   na.rm = TRUE),
    upper  = quantile(int_enrich, 0.975, na.rm = TRUE),
    .groups = "drop"
  )

# 5) Spot-check one position
ci_df %>% filter(population == "mCh_high") %>% head()
```

### Calculate stranded enrich
```{r Calculate stranded enrich - mCh_high}
enrich_fixed_stranded_mCh_high = enrichm_function_fixed_bins_stranded(tib_filt,
                                           binsize = 5000,
                                           N_bins_window = 10,
                                           calculation_window = prange_zoom_paper,
                                           CL = c("CBS16_16A", "CRE6", "3xCTCF"),
                                           EXP = all_exp)
```

### Combine boostrapped and real enrich and plot mCh_high
```{r Combine boostrapped and real enrich and plot mCh_high}

tib_boot_real_enrich_strand = ci_df %>% left_join(enrich_fixed_stranded_mCh_high, join_by(cell_line, motif, mid_interval, population))
# Change oder for plotting
tib_boot_real_enrich_strand = tib_boot_real_enrich_strand %>% mutate(cell_line = factor(cell_line, levels = c("CRE6", "CBS16_16A", "3xCTCF")))

```

# Plots for paper
##Stranded enrichment plots
```{r}
# Plotted in linear space (median)
tib_boot_real_enrich_strand %>% 
  ggplot(aes(x = mid_interval, y = median, color = population, fill = population)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2) +
  geom_line(aes(y = int_enrich), color = 'orange') + # enrich of original data
  facet_grid(cell_line ~ motif, scales = "free_y") +
  labs(
    x     = "Window midpoint (bp)",
    y     = "Enrichment (sorted / ctrl)",
    title = "Bootstrap 95% CI of enrichment"
  ) +
    list(
    geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown),
    annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown),
    geom_vline(xintercept = start(landingPad_23), col = "darkgrey"),
    geom_hline(yintercept = 1, col = "black", linetype = "dotted"),
    #annotate("rect", xmin = start(SRR2_element), xmax = end(SRR2_element), ymin = -Inf, ymax = Inf, alpha = .2, fill = 'purple'),
    # annotate gene
    annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf, alpha = .2, fill = 'red'),
    geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = 'red')) +
  
  scale_x_continuous(breaks = scales::pretty_breaks(n = 6), labels = scales::unit_format(scale = 1E-6, accuracy = 0.1, unit=NULL),
                     limits= prange_300kb, expand=c(0,0)) +
  theme_bw() +
  theme(legend.position = "bottom")


#Add CRE to it
tib_boot_real_enrich_strand %>%
filter(ctrl_counts_window >= 1 & counts_window >= 1 & cell_line %in% c("CRE6", "CBS16_16A")) %>%
  mutate(CL_pop = case_when(cell_line == "CRE6" & population == "mCh_mid" ~ "CRE6_mCh_mid",
                            cell_line == "CBS16_16A" & population == "mCh_mid" ~ "CBS16_16A_mCh_mid",
                            cell_line == "CBS16_16A" & population == "mCh_high" ~ "CBS16_16A_mCh_high",
                            cell_line == "3xCTCF" & population == "mCh_mid" ~ "3xCTCF_mCh_mid",
                            cell_line == "3xCTCF" & population == "mCh_high" ~ "3xCTCF_mCh_high",
                            .default = "other")) %>%
  ggplot(aes(x = mid_interval, y = median, color = CL_pop, fill = CL_pop)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2) +
  #geom_line(aes(y = int_enrich), color = 'orange') + # enrich of original data
  facet_grid(motif ~ ., scales = "free_y") +
  labs(
    x     = "Window midpoint (bp)",
    y     = "Enrichment (sorted / ctrl)",
    title = "Bootstrap 95% CI of enrichment"
  ) +
    list(
    geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown),
    annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown),
    geom_vline(xintercept = start(landingPad_23), col = "darkgrey"),
    geom_hline(yintercept = 1, col = "black", linetype = "dotted"),
    #annotate("rect", xmin = start(SRR2_element), xmax = end(SRR2_element), ymin = -Inf, ymax = Inf, alpha = .2, fill = 'purple'),
    # annotate gene
    annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf, alpha = .2, fill = 'red'),
    geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = 'red')) +
  
  scale_x_continuous(breaks = scales::pretty_breaks(n = 6), labels = scales::unit_format(scale = 1E-6, accuracy = 0.1, unit=NULL),
                     limits= prange_2Mb, expand=c(0,0)) +
  theme_bw() +
  theme(legend.position = "bottom")
```
  
### Do not display the missing values
```{r}
# Plot CRE with 1xCBS
bin <- 5000

pEnrich_STRANDED_high_1xCBS_CRE_25kb =
tib_boot_real_enrich_strand %>%
  filter(ctrl_counts_window >= 1 & counts_window >= 1 & cell_line %in% c("CRE6", "CBS16_16A")) %>%
  mutate(
    CL_pop = case_when(
      cell_line == "CRE6"      & population == "mCh_mid"  ~ "CRE6_mCh_mid",
      cell_line == "CBS16_16A" & population == "mCh_mid"  ~ "CBS16_16A_mCh_mid",
      cell_line == "CBS16_16A" & population == "mCh_high" ~ "CBS16_16A_mCh_high",
      cell_line == "3xCTCF"    & population == "mCh_mid"  ~ "3xCTCF_mCh_mid",
      cell_line == "3xCTCF"    & population == "mCh_high" ~ "3xCTCF_mCh_high",
      .default = "other"
    )
  ) %>%
  filter(CL_pop != "other") %>%                      # optional, but avoids drawing an "other" group
  arrange(motif, CL_pop, mid_interval) %>%
  group_by(motif, CL_pop) %>%                        # <- define what a continuous line is
  mutate(
    gap = mid_interval - lag(mid_interval),
    seg = cumsum(replace_na(gap > bin, FALSE))
  ) %>%
  ungroup() %>%
  ggplot(aes(
    x = mid_interval, y = log2(median),
    color = CL_pop, fill = CL_pop,
    group = interaction(CL_pop, seg)                 # <- breaks the line/ribbon at gaps
  )) +
  geom_line() +
  geom_ribbon(aes(ymin = log2(lower), ymax = log2(upper)), alpha = 0.2) +
  facet_grid(motif ~ .) +
  labs(
    x     = "Window midpoint (bp)",
    y     = "Enrichment (sorted / ctrl)",
    title = "Bootstrap 95% CI of enrichment"
  ) +
  list(
    geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown),
    annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown),
    geom_vline(xintercept = start(landingPad_23), col = "darkgrey"),
    geom_hline(yintercept = 1, col = "black", linetype = "dotted"),
    annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf, alpha = .2, fill = "red"),
    geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = "red")
  ) +
  scale_x_continuous(
    breaks = scales::pretty_breaks(n = 6),
    labels = scales::unit_format(scale = 1E-6, accuracy = 0.1, unit = NULL),
    limits = prange_zoom_paper, expand = c(0, 0)
  ) +
  theme_bw() +
  theme(legend.position = "bottom")


# Plot CRE with 3xCBS
bin <- 5000

pEnrich_STRANDED_high_3xCBS_CRE_25kb =
tib_boot_real_enrich_strand %>%
  filter(ctrl_counts_window >= 1 & counts_window >= 1 & cell_line %in% c("CRE6", "3xCTCF")) %>%
  mutate(
    CL_pop = case_when(
      cell_line == "CRE6"      & population == "mCh_mid"  ~ "CRE6_mCh_mid",
      cell_line == "CBS16_16A" & population == "mCh_mid"  ~ "CBS16_16A_mCh_mid",
      cell_line == "CBS16_16A" & population == "mCh_high" ~ "CBS16_16A_mCh_high",
      cell_line == "3xCTCF"    & population == "mCh_mid"  ~ "3xCTCF_mCh_mid",
      cell_line == "3xCTCF"    & population == "mCh_high" ~ "3xCTCF_mCh_high",
      .default = "other"
    )
  ) %>%
  filter(CL_pop != "other") %>%                      # optional, but avoids drawing an "other" group
  arrange(motif, CL_pop, mid_interval) %>%
  group_by(motif, CL_pop) %>%                        # <- define what a continuous line is
  mutate(
    gap = mid_interval - lag(mid_interval),
    seg = cumsum(replace_na(gap > bin, FALSE))
  ) %>%
  ungroup() %>%
  ggplot(aes(
    x = mid_interval, y = median,
    color = CL_pop, fill = CL_pop,
    group = interaction(CL_pop, seg)                 # <- breaks the line/ribbon at gaps
  )) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2) +
  facet_grid(motif ~ .) +
  labs(
    x     = "Window midpoint (bp)",
    y     = "Enrichment (sorted / ctrl)",
    title = "Bootstrap 95% CI of enrichment"
  ) +
  list(
    geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown),
    annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown),
    geom_vline(xintercept = start(landingPad_23), col = "darkgrey"),
    geom_hline(yintercept = 1, col = "black", linetype = "dotted"),
    annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf, alpha = .2, fill = "red"),
    geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = "red")
  ) +
  scale_x_continuous(
    breaks = scales::pretty_breaks(n = 6),
    labels = scales::unit_format(scale = 1E-6, accuracy = 0.1, unit = NULL),
    limits = prange_zoom_paper, expand = c(0, 0)
  ) +
  theme_bw() +
  theme(legend.position = "bottom")


plot_grid(pEnrich_STRANDED_high_1xCBS_CRE_25kb, pEnrich_STRANDED_high_3xCBS_CRE_25kb, ncol = 1, axis = "lr", align = "v")
```






