---
title:  "CM20230406_E2227_genotype_clones"
author: "Christine Moene"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: 
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: inline
---
# Goal
I sequenced amplicons around the CTCF sites for Del D and E clones.
I used a little snakemake pipeline to extract and count the sequences inbetween the primers. For now I count unique sequences (and not grouped by starcode), because the real difference between the two alleles might only be one or two bp.

Here I will try to determine which sequences are wildtype and which are (useable) mutations.


## Set-up
Load the libraries and set the parameters.

```{r load_libraries}

# Load dependencies
library(GenomicRanges)
library(rtracklayer)
library(tidyverse)
library(dtplyr)
library(dplyr)
library(tidyr)
library(ggbeeswarm)
library(ggplot2)
library(Biostrings)
library(stringr)
library(readr)
library(knitr)
library(RColorBrewer)
library(RcppRoll)
library(plotly)
library(readxl)
library(ggpubr)

split_string <- function(vect,sep,N1,N2=N1){
  library(stringr)
  sapply(vect, function(X){
    paste(str_split(X,sep)[[1]][N1:N2],collapse = sep)
  },USE.NAMES = F)
}
```

```{r paths_parameters}
#datatag
datetag = paste0("CM",format(Sys.time(), '%Y%m%d'))

# Prepare output 
output_dir <- paste0("/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/E2227_genotyping_CTCF_deletion_clones/analysis_E2227_",datetag)
dir.create(output_dir, showWarnings = FALSE)

```


```{r knits setup}

library(knitr)
opts_chunk$set(fig.width = 20, fig.height = 8, cache = T,
               message = F, warning = F,
               dev=c('png', 'pdf'), 
               fig.path = file.path(output_dir, "figures/")) 
pdf.options(useDingbats = FALSE)
```

# Load the data
```{r}
StarcodeDataDir <- "/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/E2227_genotyping_CTCF_deletion_clones/snakemake_output/starcode/"

# List all files
metadata_starcode <- tibble(file = dir(StarcodeDataDir, 
                   recursive = T, full.names = T)) %>%
  # filter(grepl("insertions_ref", file)) %>%
  mutate(sample = str_remove(basename(file), "\\..*")) %>%
  filter(grepl("starcode0_sequences", file))

# Load all files
starcode_tib <- bind_rows(lapply(1:nrow(metadata_starcode),
                                 function(i) {
                                   tmp <- read_tsv(metadata_starcode$file[i],
                                                   col_names = c("sequence","count","alternative_seq"),
                                                   col_types = cols(
                                                     .default = col_character(),
                                                     sequence = col_character(),
                                                     alternative_seq = col_character(),
                                                     count = col_double())) %>%
                                     mutate(sample = metadata_starcode$sample[i],
                                       clone_type = split_string(sample,"_",1))
                                   # mutate(pool_number = as.numeric(substr(pool, 2, length(pool))))
                                 }))
# saveRDS(starcode_tib, "CM20230406_E2227_all_sequences_starcode0.tib")


```

Immediately filter for the top 5 sequences per sample, and remove the 'alternative_seq' column (is identical to the 'sequence' column), to make the tibble smaller (more manageable)
```{r}
starcode_tib2 = starcode_tib %>%
  select(sequence, count, sample, clone_type) %>%
  group_by(sample) %>%
  slice_max(order_by = count, n = 5, with_ties = F)

top_sequences_D_tib = starcode_tib2 %>%
  filter(clone_type == "DelD") %>%
  slice_max(order_by = count, n = 3, with_ties = F) %>%
  mutate(fasta_name = paste0(sample,"_",count,"reads"))
  
top_sequences_E_tib = starcode_tib2 %>%
  filter(clone_type == "DelE") %>%
  slice_max(order_by = count, n = 3, with_ties = F) %>%
  mutate(fasta_name = paste0(sample,"_",count,"reads"))
 
top_sequences_D = DNAStringSet(top_sequences_D_tib$sequence)
names(top_sequences_D) = top_sequences_D_tib$fasta_name
top_sequences_E = DNAStringSet(top_sequences_E_tib$sequence)
names(top_sequences_E) = top_sequences_E_tib$fasta_name

writeXStringSet(top_sequences_D, "CM20230406_E2227_top_sequences_D.fasta")
writeXStringSet(top_sequences_E, "CM20230406_E2227_top_sequences_E.fasta")

```